\section*{Example: The Game Begins}

\begin{multicols}{3}

\noindent
The table has two players because one is running late.
Such is life.


\begin{description}\sf
  \item[Player 1:]
  Shall we start character creation now?
  \item[\Glsentrytext{gm}:]
  Yes!
  There's your characters, already made.
  Who are they?
  I haven't looked at any of them.
  \item[Player 1:]
  Says here I'm a thief called `\composeHumanName'.
  \item[Player 2:]
  I got `\composeHumanName'; also a thief.
  Hopefully the next one's a fighter.
  \item[\Glsentrytext{gm}:]
  All the characters are thieves, cutthroats, dick-heads, and bandits, and most of them fighters as well.
  You were sentenced in the \gls{court} to repay your debt to society by joining the \gls{templeOfBeasts}.
  \item[Player 1:]
  It says on the sheet that the debt's `100~\glsentrylongpl{sp}'.
  \item[\Gls{gm}:]
  \Glspl{guard} may not set foot in towns, \glspl{village}, or other civilized places until they repay the debt.
  You live in the between-places; \glspl{bothy} and \glspl{broch}, keeping others safe by hunting \glspl{monster}.
  \item[Player 1:]
  And we pay back the debt with our wages from this \gls{temple}?
  \item[Player 2:]
  What's \pgls{bothy}?
  \item[\Gls{gm}:]
  No, you pay the \gls{temple} half of what you receive from selling monster bodies, and do missions.

  \Pgls{bothy}'s a half-way house, where people sleep just to stay safe from the \glspl{monster}.
  You're coming up to Stoatfen \Gls{bothy} now, where you'll receive your first mission.

  \item[Player 1:]
  What's a `Stoatfen'?
  \item[\Gls{gm}:]
  Each \gls{bothy} starts as a cairn --- a pile of rocks --- to remember someone who died on the road.
  If enough people bring enough rocks, then someone builds \pgls{bothy}.
  So something ate someone called `Stoatfen', just up ahead, and now there's \pgls{bothy}.
  \item[Player 2:]
  Why are the names weird?
  \item[\Gls{gm}:]
  Local custom.
  People say the forest eats the best things in life, so naming your kid `Sunbeam' is just asking for something to come out the forest and eat it.
  \item[Player 2:]
  Are we there yet?
  \item[\Glsentrytext{gm}:]
  Yes.
  The road widens, and the shadows of the tall trees disperse.
  Stoatfen~\Gls{bothy} has the standard clearing of 50~\glspl{step} in all directions.
  Trees felled, and bushes burnt, to push back the forest, leaving a circle of
  \ifnum\value{temperature}=0%
    white%
  \else%
    yellowed grass%
  \fi.%
  Where most \glspl{bothy} have only one room, Stoatfen has many.
  But it performs the same function; it lets people sleep safely.
  \Glspl{monster} are wandering.
  \item[Player 1:]
  So people here are aware of random encounters?
  \item[\Glsentrytext{gm}:]
  Very.
  Once or twice a week on average, something comes from the forest to feed.
  \item[Player 2:]
  So this place is \gls{monster}-proof?
  \item[\Glsentrytext{gm}:]
  Mostly, but they still try.
  Sometimes one feels along the walls and roof, searching for an opening during the night, while people inside debate what kind of creature is trying to come in.
  \item[Player 1:]
  So they only come out at night.
  \item[\Glsentrytext{gm}:]
  No, they often wait until morning.
  That's where you come in!
  In \pgls{bothy}, people expect the \gls{guard} to open the door in the morning, maybe wave \pgls{dawnDolly} around, and give the all-clear.
  \item[Player 2:]
  A what?
  \item[\Glsentrytext{gm}:]
  Like a scarecrow on a stick.
  You waggle it outside the door in the morning, and if something reaches down from the roof and grabs it, then you know there's something on the roof.
  \item[Player 2:]
  How the hell does anyone live in this place?
  \item[\Glsentrytext{gm}:]
  How would you live?
  How would you farm?
  \item[Player 3:]
  Hey guys!
  Did I miss the start?
  Have you done character creation?
  \item[\Glsentrytext{gm}:]
  Here's your character.
  Now, how do you think you'd farm in a world of wandering monsters?
  \item[Player 3:]
  Why's my name `\composeHumanName'?
\end{description}

\bigLine
\vspace{2em}

\noindent
The `wandering monsters' question makes a nice opener, as people can have their own ideas about it.
It also explains most of what the \glspl{pc} do, and why society needs them.

\bigLine

\begin{description}\sf
  \item[\Glsentrytext{gm}:]
  Stoatfen looks dead.
  Someone's left the doors and windows open.
  And somewhere upwind, past the turn in the road, the smell of badly rotten meat.

  What do you do?
  \item[Player 1:]
  Clearly the rotten meat-smell.
  We'll go towards that.
  \item[Player 2:]
  That can't be something we can help with.
  Let's check the \gls{bothy}'s safe.
  \item[\Glsentrytext{gm}:]
  It is, but let's do a roll, just to get a feel for the how actions work here.
  You roll two dice, then add your wits and a skill, and try to get over the \glsfmtlong{tn}.
  The \glsfmtlong{tn} is eight.
  What's your wits and vigilance?
  \item[Player 2:]
  \twoDice{4}
  Minus one, and one, so none.
  So four in total.

  So if there were \pgls{monster}, it would have eaten me?
  \item[\Gls{gm}:]
  Ambush first, eat second, but yea --- that's broadly it.
  \item[Player 1:]
  \twoDice{10}
  I got it!
  \item[\Gls{gm}:]
  Just one roll for everyone I'm afraid.
  Everyone adds their own traits, but the dice stay put.

  Also, what's your wits?
  \item[Player 1:]
  How is my wits that bad?
  \item[\Gls{gm}:]
  Maybe the character likes to consider all options before doing anything?
\end{description}

\bigLine
\vspace{2em}

The roll was to investigate Stoatfen~\Gls{bothy} for clues, not checking if anything's inside.
But it's the same roll in either case, so now the game can proceed without players shifting focus to what they missed.

\Glspl{gm} should never lie, but misleading works a treat.
\end{multicols}

\bigLine

\begin{multicols}{3}
\begin{description}\sf
  \item[\Gls{gm}:]
  A wagon approaches from the road with the rotten smell.
  It comes right up to the corner, then stops.
  \item[Player 1:]
  They must be investigating.
  Why can't we investigate?
  \item[Player 3:]
  I'll go investigate, \glspl{monster} be damned.
  \item[\Gls{gm}:]
  Round the corner, two wagons and some large rocks block the road.
  Two men lie dead and rotting, next to bits of broken wood, and old mushrooms.
  Behind them all, a tubby man in black, with greasy hair and a wax record-tablet.
  And behind him, his wagon, and half a dozen more \glspl{guard}.

  He shouts `\textit{have you checked inside?
  Are you the new \glspl{fodder}?}'.
  \item[Player 3:]
  The what?
  \item[Player 1:]
  No manners.
  What happened to `hello?'.
  \item[Player 1:]
  Yes, we checked.
  \item[Player 3:]
  Yea, we have the rank of `\glspl{fodder}'.
  It's on the sheet.
  \item[Player 3:]
  Nevertheless.
  \item[\Gls{gm}:]
  The \glspl{guard} around him are rolling their eyes, clearly used to his behaviour.
  Then he asks you to clear the wreckage from the road as he gives his speech.

  \item[Player 1:]
  I guess we'll clear then.
  Just one roll again?
  \item[Player 2:]
  Wait, we can't work together?
  \item[\Gls{gm}:]
  Sure you can.
  Roll and add strength and crafts at \glsfmtlong{tn} ten, then the second person adds half their total.

  The large man introduces himself as `\pgls{susjot}', and says `\textit{welcome to the \gls{templeOfBeasts}!}
  \item[Player 1:]
  \twoDice{5}
  I have plus nothing.
  \item[Player 2:]
  Just plus one.
  \item[Player 3:]
  Same, so we get one and a half?
  \item[\Gls{gm}:]
  Always round up on halves.
  It's a plus two, so the total's seven.
  That means it takes all morning, as carriages get stuck in the ground while turning.
  \item[Player 1:]
  Why didn't \pgls{monster} eat the bodies?
  \item[\Glsentrytext{gm}:]
  \Gls{susjot} replies that most like fresh meat and %
  \ifnum\thetemperature>0%
    luckily, there's no \glspl{basilisk} about.
  \else%
    the \glspl{basilisk} hibernate when it snows.
  \fi%

  So after a morning of hard labour\ldots
  \item[Player 1:]
  And apparently no help from anyone, just speeches.
  \item[\Gls{gm}:]
  \ldots a carriage rounds the bend\ldots
\end{description}

\end{multicols}

\bigLine
\vspace{2em}

\begin{multicols}{2}
\noindent
The slow start here gives players a chance to get used to rolling dice to resolve \glspl{action}, and in particular how BIND handles things.
For example, in that last roll, the dice are rolled first, and the character's reaction comes from interpreting the roll.

Nearly all of BIND's systems are just reinterpretations of the same basic roll: $2D6$ plus \gls{attribute} plus \gls{skill}.
Sometimes \gls{equipment} adds a Bonus, like \pgls{weapon} during combat, or using a map for \gls{navigation}.

\begin{enumerate}
  \item
  The \gls{pc} wants some prize (e.g. `check building', or `notice someone stayed the night alone, without a fire').
  \item
  The \gls{pc} wants to avoid a danger (`not jumped by \pgls{monster}', or `avoid being tricked').
  \item
  The \gls{gm} decides on the most appropriate \gls{attribute} and \gls{skill} to use, then the \glsentrylong{tn}.
\end{enumerate}

\begin{tabularx}{\linewidth}{ccccccL}
\hline
$2D6$ & + & \gls{attribute} & + & \gls{skill} & vs & \Glsfmttext{tn} \\
\hline
$2D6$ & + & Intelligence & + & Vigilance & vs & 7 \\
$2D6$ & + & Intelligence -1 & + & Vigilance 2 & vs & 7 \\
$2D6$ & + & -1 & + & 2 & vs & 7 \\
\twoDice{6} & + & -1 & + & 2 & vs & 7 \\
\hline
6 & + & 1 & = & 7 & vs & 7 \\
\hline
\end{tabularx}

\medskip

\noindent
If the roll on the \gls{tn} (i.e. `a tie'), both results occur, or neither, depending on what makes sense.
If both make sense, the player usually chooses.

Interpretation is key, as each roll should be explained.
A skilled hunter missing tracks should be explained by the \gls{gm} with bad conditions, while a successful roll to steal a coin-purse, or a failed roll to convince \pgls{jotter} to grant another \gls{torch} can be fleshed out by the player.

Once players feel comfortable with the dice, it's time to throw rapid decisions at them.

\end{multicols}

\bigLine

\begin{multicols}{3}

\begin{description}\sf
  \item[\Gls{gm}:]
  A couple of miles up the road, a caravan approaches.
  \Pgls{susjot} shouts out, ``\textit{there she is -- \gls{woetide}, the only one to escape the goblin attack}''.

  Two wagons arrive, flanked by two more \glspl{guard}.
  \Gls{woetide} tips her wide hat at you, but she just looks worried.

  She points through the trees, and informs everyone that she's just seen \glspl{guard}, exiting the old mine, carrying a chest of gold.

  \item[Player 3:]
  How does she know it was a chest of gold?
  \item[\Gls{gm}:]
  \Gls{susjot} starts shouting that's it's obviously the same chest that the goblins stole.
  Then he shouts at the \glspl{soldier} to hunt them down, quickly.
  \item[Player 1:]
  Should we help?
  \item[\Gls{gm}:]
  ``\textit{No, we have a smaller job for the \glspl{fodder}.
  Get up that mountain and into the old mine.
  Anyone who brings back a goblin head get promoted to \gls{gDigger}.}''
  \item[Player 1:]
  Okay, looks like we're goblin-hunting.
  \item[\Gls{gm}:]
  \Gls{woetide} wishes you well, and goes to leave saying \textit{``travel straight, never curious''}.
  \item[Player 2:]
  `Never curious'?
  What kind of farewell is that?
  \item[Player 1:]
  Curiosity's dangerous.
  \item[Player 2:]
  Can the other \glspl{guard} with \gls{woetide} help us?
  \item[\Gls{gm}:]
  Yes.
  Maybe?
  Do you want to spend \pgls{storypoint}?
\end{description}

\end{multicols}
