\newcommand\hibernationSegment{
  \segment[\glsentrysymbol{sable}]{lonelyRoad}% AREA
  {Don't Wake the Baby}% NAME
  {Three \glsfmtplural{cutter} crunch through the \gls{snow} with a hibernating \glsfmttext{hulk} in their cart}% SUMMARY

  \begin{exampletext}
    Three \glspl{cutter} found \pgls{hulk} hibernating in a cave.
    They trundled a cart into the cavern, loaded the \gls{hulk} onto it, began moving towards \gls{ftown} to sell it to the \gls{court} so people can watch it fight.
    The \gls{hulk} should fetch 3~\glspl{gp} at least!
    But first, they must find \pgls{witch}, or anyone who knows enough \gls{alchemy} to ensure the beast remains in hibernation.
    Because it's starting to wake up\ldots
  \end{exampletext}

  \ifnum\thetemperature>0
    This \gls{segment} only works during \gls{cTwo}, so if the \gls{snow} has not fallen, skip it, or just introduce the \glspl{cutter} talking about the \gls{hulk} they killed last \gls{cTwo}.
  \fi

  \begin{boxtext}
    A cart appears ahead, coming slowly, and silently.
    \Pgls{guard} pulls it, while two push from behind, then one stops, and seems to be indicating something to you with his hands.
    He looks serious, but doesn't say anything.
  \end{boxtext}

  \paragraph{If the \glspl{pc} help,}
  they have the \glspl{cutter}' thanks.
  They're tired, and don't have the energy to run to a nearby \gls{village}.

  \humansoldier[\npc{\glsentrysymbol{sylf}\T[3]\F\M\Hu}{\composeHumanName, \composeHumanName, \& \composeHumanName}]

  \paragraph{Any loud noises}
  could wake the \gls{hulk}; roll $1D6$ and on the roll of \dicef{1}, the \gls{hulk} wakes and attacks after 3~\glspl{round} (\roll{Wits}{Survival} at \tn[12] to notice it waking).

  \umberhulk
}

\newcommand\fridgeMap[1][t]{
  \renewcommand\csComments{
    \mapFeature{15,2}{\Huge \glssymbol{abderian}}
    \mapFeature{16,9}{\Huge \glssymbol{wrecan}}
    \mapFeature{3,6}{\Huge \glssymbol{sylf}}
    \mapFeature{0.3,1}{\Huge \glssymbol{sable}}
    \mapFeature{13,6}{\Huge \glssymbol{sable}}
    \mapFeature{12,3.6}{\Huge \glssymbol{paik}}
    \mapFeature{8,12}{\Huge \glssymbol{yonder}}
    \colorlet{background}{white!10!black!90}%
    \callout{17,5.7}{0,0}{\ref{fStairs}:~\nameref{fStairs}}{}
    \callout[7]{17.5,4}{-0.5,0}{\ref{fHallStore}:~\nameref{fHallStore}}{}
    \callout[8]{13,4.8}{0,0}{\ref{fLift}:~\nameref{fLift}}{}
    \callout[9]{17,1}{-0.5,0.5}{\ref{fWheat}:~\nameref{fWheat}}{}
    \callout[9]{13.2,1}{0.5,0.3}{\ref{fIngStore}:~\nameref{fIngStore}}{}
    \callout[10]{17,10.5}{-0.5,-0.5}{\ref{fOffice}:~\nameref{fOffice}}{Everything entered must be registered}
    \callout[12]{15,12}{-0.7,-3}{\ref{fDip}:~\nameref{fDip}}{with oil barrels}
    \callout{11,9.8}{0.8,0}{\ref{fSide}:~\nameref{fSide}}{}
    \colorlet{background}{white!10!black!80}%
    \callout[9]{7,5.8}{0.8,-0.3}{\ref{fBoat}:~\nameref{fBoat}}{}
    \callout[6]{0,6.5}{0.6,-1}{\ref{fWeaponStore}:~\nameref{fWeaponStore}}{}
    \callout[6]{0,8}{0.3,0}{\ref{fCrossbow}:~\nameref{fCrossbow}}{}
    \callout[8]{1,11}{0,-1}{\ref{fHibernation}:~\nameref{fHibernation}}{}
    \colorlet{background}{white!10!black!70}%
    \callout[10]{4.5,2.8}{-1.1,-0.3}{\ref{fCheeseStorage}:~\nameref{fCheeseStorage}}{}
    \callout[8]{3.4,1}{-0,0.4}{\ref{fFabrics}:~\nameref{fFabrics}}{}
    \callout[10]{2,0}{-0.4,0.5}{\ref{fGrain}:~\nameref{fGrain}}{}
    \callout[6]{0,3.6}{0.2,-0.3}{\ref{fSpirits}:~\nameref{fSpirits}}{}
    \colorlet{background}{white!10!black!60}%
    \callout[10]{2,14}{4,-1}{\ref{fLibrarians}:~\nameref{fLibrarians}}{}
    \callout[8]{9.8,10.8}{0,0.7}{\ref{fGiftschrank}:~\nameref{fGiftschrank}}{}
    \callout[9]{9.9,13.8}{0,-0.8}{\ref{fBrack}:~\nameref{fBrack}}{}
    \callout[12]{14.3,13.5}{-2,-0.9}{\ref{fLibrary}:~\nameref{fLibrary}}{}
    \colorlet{background}{white!10!black!50}%
    \callout[7]{6,0}{1,1.2}{\ref{fDoor}:~\nameref{fDoor}}{}
    \callout[8]{10,0}{-0.6,1.3}{\ref{fWaiting}:~\nameref{fWaiting}}{}
    \callout[9]{12,2.2}{0,0}{\ref{fCourt}:~\nameref{fCourt}}{}
  }

  \widePic[t]{Dyson_Logos/bowels}


}
