\section{The House of Sable}


% Light candles to keep track of time.
% Enter AP tracking soon after.

\begin{exampletext}
  Renchkolt woke at Sundown with dread.
  \ifcase\value{temperature}%
    After promotion to \gls{jotter} via clerical error, he had wanted someone to teach him how to read books, and maybe earn some extra coins by writing books.
    He put on his boots and went to count \glspl{monster}.
  \or%
    Today would be just like the rest.
    Down underground, to freeze his bollocks off while counting \glspl{monster}.

  \else%
    He dragged his overcoat over his work clothes.
    People stare with confusion when a man wears his big overcoat on a \showTemperature\ day, but they stare with contempt when they see the \gls{guard} uniform.
  \fi%

  First he turned into Nettleway --- while still trying to pleat his scraggly beard --- to pick up the rest of his clothes at the \gls{weaversGuild}.
  One of the \glspl{weaver} told him to sit down as she put on a tea, another asked him to pass on a message to his mum, and the next scolded him for leaving his colleague alone in the darkness.

  Renchkolt left with three candles and a telling-off.
  And on tea.
  The stairs down are always open to the main street, but nobody goes down without a good reason, so nobody locks it.
  The town fridge has no door.
  The street-rats usually can't afford the light, and the serious criminals know better than to mess with all of the \glspl{guild} at once.

  At the first bend, the real darkness begins.
  Even over \gls{cFive}, you walk with a candle or you walk blind.
  By the second bend, the air goes still and you stop holding the candle protectively.
  His longsword tapped against the wall here and there.
  Renchkolt often wanted to leave it at home, or sell it, but it's part of the uniform.
  The \gls{templeOfBeasts} never bothered to make it official, because everyone knows what the \gls{guard} look like.

  The \gls{templeOfPoison}'s barrels of their new beer made strange noises in darkness to the left.
  After years of coming down here, he'd never been to the left.
  Survival instincts like that were probably how he received his promotions, though he never asked, just smiled and accepted new titles with an enthusiastic handshake.

  \begin{speechtext}
    Is.That. YOU, Renchkolt!?
  \end{speechtext}

  \Gls{proctor} Cainpelt liked to remind people that he had exceptional hearing, so he couldn't be all that old, despite the grey hair poking out from under his hat.
  He would not acknowledge that anyone might guess who was coming, given that most went to their work at the same time each day.
  The \gls{proctor} didn't seem to have any actual work to do, as far as anyone could see, so that probably helped his guessing games.
  Maybe he just sat in the dark shouting names until someone came by?

  \begin{speechtext}
    `So that's one SWORD\ldots

    \ldots

    \ldots and one CANDLE.
    \ldots
    and
    \ldots OVER-COAT.
  \end{speechtext}

  \Pgls{proctor} usually resolves disagreements.
  This one stopped disagreements by recording every potential fact before it entered the frigid hall.

  Down the hall, down the stairs, passing more barrels.
  This hall was colder than the last, and the barrels quieter.

  Up the stairs, and through the narrow arch, the air grew colder, and more barrels lay, nearly silent.

  Past the tall columns of the mirror-lake, four \glspl{temple} hold four vaults.

  \begin{itemize}
    \item
    The \gls{templeOfFrost} holds preservatives, wool, and food, including a hallway lined entirely with cheese wheels.
    \item
    The \gls{templeOfCuriosity} holds their `Secret Library' (meaning the contents \emph{should} remain secret, not that nobody knows about the library, because everyone --- even illiterate beggars --- know about the Secret Library).
    \item
    The \gls{templeOfPoison} holds a massive network of caves, all stuffed with dry, barrelled, immortal wheat.
    Or so they say.
    The doors are over to the left of the great hall, but Renchkolt has never wandered off to the left to see what's in the new darkness.
    \item
    The \gls{templeOfBeasts} has Cuspid, who always asks if we can go for a drink after work, while he drinks at work.
  \end{itemize}

The two \glspl{guard} climbed to the Vault of the \gls{templeOfBeasts} and unlocked the door.
They checked the giant crossbow was in working order (mostly by saying `\textit{hmm, yes, okay}' while poking at it, but never actually using it).
Then they went to count the \glspl{monster} eyes.



This time the count is twenty-\emph{two}.
Maybe twenty-three?



\end{exampletext}


%%%%%%%%

% 
% Alcoholic Guard
% Wheel of cheese walls
% Temple of Hate office.
% 

