\section{The Fridge}
\label{theFridge}

\pacingExample

\begin{multicols}{2}

\noindent
The \gls{templeOfFrost} holds a massive, underground fridge, used by the entire city to store food, \glspl{ingredient}, and hibernating \glspl{monster}.

Just about every \gls{temple} in the city holds something down there, and none of them trust the others enough to go down and fetch anything, for fear they'll place the items in their own area.

\subsection{Turns' Relentless March}

The session should track time from start to finish, because the fridge is melting.
Just measure broad turns with \glspl{ap} as usual, \emph{but} characters spend \pgls{ap} to:

\begin{itemize}
  \item
  Move anywhere in line of site.
  \item
  Get into a boat.
  \item
  Engage in a conversation.
  \item
  Open the heavy doors to room \vref{fDoor}.
  \item
  Engage in combat (even spending 4 \glspl{round} just counts as one turn).
\end{itemize}

Whenever the troupe moves, every \gls{pc} must spend \pgls{ap} or they lag behind.
However, their pace depends on decisions more than the \glspl{pc}' Speed Bonuses (note the example, \vpageref{pacingExample}).

Whenever a new turn begins, chech which events occur.

\subsubsection[Upon Entry]{1: Upon Entry}
the water has begun to melt.
The dark centre of the map has waist-high water to a human, and the grey sections of the map (\vpageref{Dyson_Logos/bowels}) have no water yet.

\subsubsection[Water Rises]{4: Water Rises}
through the grey areas, but stops to fill up the \nameref{fDip} in room~\vref{fDip}.

\subsubsection[Oil Spills]{5: Oil Spills}
from the barrels in room \ref{fDip}, and water moves through the rest of the grey areas.

\subsubsection[An Escaped Prisoner]{7: An Escaped Prisoner}
jumps from the \gls{court} and out the doors of room \vref{fDoor}, then heads towards the nearest light (e.g. any \gls{torch} \pgls{pc} carries).
\Pgls{sunGuard} bolts after him a moment later with his own \gls{torch}, and the prisoner has to swim to safety.

The \glspl{pc} might not want to waste their time on this ruckus, but the \gls{sunGuard} wants their boat (he's not getting into that filthy water with his nice, white, vestige).

\humanthief[\NPC{\M\Hu}{\composeHumanName}%
  {one eye, bald}% DESCRIPTION
  {spinning thumbs}% MANNERISM
  {find valuables, then bolt}% WANTS
  \npcQuote{Wouldn't want to join you lads, no offence yea?}]% QUOTE

\sunguard[\NPC{\glsentrysymbol{paik}\M\Hu}{\Glsfmttext{sunGuard} \composeHumanName}%
  {muscular frame, sticky-out ears}% DESCRIPTION
  {stands uncomfortably close while yelling}% MANNERISM
  {to catch the criminal}% WANTS
  \npcQuote{Where is the cretin?
  Give me that boat!}]% QUOTE

\subsubsection[Waist Height Water]{9: Waist Height Water}
means everyone has to spend 2~\glspl{ap} to move when not in a boat.

\subsubsection[Water Moves up the Stairs]{11: Water Moves up the Stairs}
entering the rooms \vpagerefrange{fDoor}{fCourt}, and moves towards the \nameref{fCourt} \vpagerefrange{fGuardDoor}{fHibernation}.

\subsubsection[The \Glsfmtplural{monster} Wake]{12: The \Glsfmtplural{monster} Wake}
and the first moves to the door at room \vref{fGuardDoor}.
The \glspl{monster} also spend \glspl{ap} to move.
\fridgeMap

\subsection{Changing Space}

\mapentry[fStairs]{Stairs}

Descending these long stairs turns the noisy streets above into silence, and replaces the
\ifnum\value{temperature}>2%
  warm air with cold%
\else%
  \showTemperature\ winds with stagnant air%
\fi.

\mapentry[fHallStore]{Barrels in the Hallway}
\begin{exampletext}
  \Gls{chef} Ginshank has been experimenting with brewing larger, but has moved the caskets into the warm hallway too early.
\end{exampletext}

\begin{boxtext}
  Creaks are coming from the left, where an alcove holds two dozen wooden containers: caskets, barrels, and crates.
\end{boxtext}

\paragraph{Prodding the boxes}
is dangerous.

\Glsfmtname{lagerCasket}
\glsentrydesc{lagerCasket}.

\Glsfmtname{flour}
\glsentrydesc{flour}.

\mapentry[fLift]{Crank Shaft}

\begin{exampletext}
  The \gls{weaversGuild} lies directly above this room, so the \gls{templeOfFrost} commissioned a lift here.
  They don't need a lift, but the \gls{templeOfPoison} \emph{really} need a lift for their barrels, so they can avoid the stairs.
\end{exampletext}

\begin{boxtext}
  This circular room has nothing inside but a tapestry depicting a naked man with a peaceful smile and a black nose, sleeping in the snow.
  It stretches from the ceiling down to a wheel on the wall.
  The wheel has four metal rods sticking out, and one wedged into a hole in the ground.
\end{boxtext}

\paragraph{If the \glspl{pc} look behind the depiction of \gls{sable},}
they find two chains reaching from the crank through the ceiling.

\paragraph{Removing the rod and turning the crank}
lowers a circular segment of the ceiling, and exposes the \gls{weaversGuild} above.

\begin{boxtext}
  The wheel spins with a rusty squeal, and light floods down from behind a circular segment of the ceiling.
  Four chains lower the wooden disk gently.
  A fire crackles above as six women look down, with stitches in their hand.
  A dozen voices from above ask `who's that?', and `where are you going?', `is it cold down there?', `do you need a jumper?'.
\end{boxtext}

\paragraph{If the \glspl{pc} stop to chat,}
they can gain useful information and free \gls{equipment} but lose time chatting.
People who work in the \gls{weaversGuild} take in facts like a river drinks.
Every day, everyone in \gls{ftown} needs their clothes washed, mended, adjusted or more clothes created.
The \glspl{weaver} now think it's normal to recite recipes, while memorizing a weave pattern, by listening to an argument about how to start it.

\begin{speechtext}
  \rotatebox{-3}{Do you need \pgls{torch}?
  We've got one here somewhere!}

  Do you know about the barrels?
  Very important, they said.
  Dwarvish `special ale', they said.
  Should be good.

  {\small
    Is that you, `\composeHumanName'?
    I don't see so well now.
  }

  \vspace{-.5em}
  {\large
    Mind the dragon!
    We don't want you to get \emph{too} roasty-toasty.
  }

  That dragon's not real!
  Did you lot hunt any \gls{griffin} recently?
  It's just that we could use some parts soon.
  I got this new preservation technique\ldots

  \vspace{-.3em}
  \rotatebox{5}{Mind ye shut that hatch!}
\end{speechtext}

The \glspl{weaver} happily gift the \glspl{pc} anything they might reasonably have to-hand.
Each item costs $1D6$~\glspl{ap} while someone locates it and passes it down to the troupe.

\paragraph{Before the \glspl{pc} leave}
the \glspl{weaver} \emph{demand} that they crank the floor back up so they don't have to worry about any \glspl{monster} coming up.

\begin{speechtext}
  We're not light the fire to listen to it, okay?

  \vspace{-.4em}
  \rotatebox{4}{Is that 'Shank cunt with the silly hat down there?}

  \vspace{-.4em}
  \rotatebox{5}{\large I hope there is a dragon, so it eats him.}

  \vspace{-.3em}
  \rotatebox{-5}{\small That the one that called your one a `hussy'?}

  \rotatebox{3}{\slshape I put some sugar in his caskets, that'll teach him\ldots}

  Too right!
  He puts his aprons in here again and I know the stitches I'll fray.

  \vspace{-.3em}
  \rotatebox{3}{\slshape Sootmill!  That's dirty.}

  \vspace{-.3em}
  \rotatebox{-2}{\sl Tell Cuspid I kept his tea warm!}

\end{speechtext}

\mapentry[fWheat]{Flour Drying Room}

\Glspl{server} scatter dusty chalk in this room so it can absorb the moisture, and let the flour dry.
The flour rests in wide-open trays, stacked on shelves.

The room has no brooms,%
\footnote{The brooms are all in room \vref{fSide}.}
so if \pgls{pc} enters, they cannot help their chalky footprints.

\mapentry[fIngStore]{\Glsfmtplural{ingredient} Storage}

The \gls{templeOfFrost} keeps \glspl{ingredient} for preservation \gls{alchemy} in here,
\footnote{Technically, this uses the `Death' \glsfmttext{sphere} (composed of Fate and Air) to stop rot.}
along with steel cauldrons, glass tubes, books on \gls{alchemy} and other rare materials.
Each one has \pgls{weight} of~1 and could fetch around 5~\glspl{sp}.

\paragraph{Opening the door}
requires an \roll{Intelligence}{Crafts} check.

\begin{boxtable}
  \Gls{tn}~12 & The \gls{pc} slides the inner bolt across, cracks the door, then gently pulls it upward to avoid any squeak as it opens. \\
  \Gls{tn}~10 & They open the door, and the rusting hinges squeal so loud that the \glspl{weaver} above \vref{fLift} hear the intrusion, and decide that the \glspl{pc} are the `wrong sort'. \\
  \Gls{tn}~9 & They break the lock-mechanism while slamming the door open, which breaks a mirror and some half-prepared Fate~\glspl{ingredient}. \\
\end{boxtable}

\mapentry[fOffice]{Office}

The \gls{templeOfHate} have secured rights to record everything which passes through here to stop the other \glspl{temple} fighting.
Stacks of catalogues hold information on every item within the fridge, including the books in the giftschrank (room \vref{fGiftschrank}).
A ledger notes everyone who has passed in or out over the last five years.

\paragraph{Finding specific information in the \nameref{fOffice}}
means sorting through the \gls{temple}'s peculiar style of record-keeping, and their own brand of short-hand notes.
This isn't intentional encryption, which somehow makes it feel even worse.
The roll is \roll{Wits}{Academics} at \tn[14].
Rolling a tie means the characters finds something random; roll $3D6$ and tell the player the contents of the room with that number.

\mapentry[fDip]{Chasm Barrels}

\begin{exampletext}
  These little runways, down to a natural cavity and back up again, make the worst part of anyone's day if they have to transport barrels through here.
  Unfortunately, the \gls{wheatGuild} refuse to pay for a small bridge.
\end{exampletext}

The cavity has two-dozen barrels of cooking oil.
If the oil ever spills and receives 8~Damage from fire, it combusts, inflicting $\dmg{-2}$~Damage.
The fire continues to inflict the same Damage rolled, minus one, each \gls{round}.
So `3~Damage' continues to burn for another 2~Damage on the second \gls{round}, then 1~Damage before dying down.
\Gls{dr} only counts for the first \gls{round} as the oil seeps into the cracks while heating the armour.
\iftoggle{intro}{
  \Glspl{fp} which mitigate the Damage reduce the Damage on subsequent \glspl{round}.
}{}

\mapentry[fSide]{Side Store}

Pickaxes, chalk, brooms, ropes, shortswords, and barrels of salt.

%%%%%%%%%%%%

\mapentry[fBoat]{Boat by the Shore}

Any \gls{pc} with at least 2 points in \roll{Wits}{Vigilance} spots a little boat, tethered to the shore.
It can take a total \gls{weight} of~100.%
\iftoggle{intro}{%
  \footnote{Characters have \pgls{weight} equal to their max. $\glspl{hp} \times 3$ plus the \gls{weight} of their items.}
}{}

\begin{boxtext}
  The \gls{torch}-light shines on a canopy of stone pillars, standing in an underground lake.
  The shore has no tide.
  The water's surface almost looks like a massive mirror.

  Along the shore, far to the left, caskets and crates.
\end{boxtext}

Close to the boat, three crates of \gls{flour} stand ready to go into long-term storage.%
\footnote{\Gls{chef} Ginshank hasn't gotten around to it yet, but says he will soon.}
Five \glspl{lagerCasket} stand next to them, heating far too rapidly, and on the edge of exploding.


%%%%%%%%%%%%

\mapentry[fGuardDoor]{Guard Post}

In the distance, Renchkolt flails his arms, hoping to get the \glspl{pc}' attention.
He knows the \glspl{monster} will awaken soon, and wants someone to kill them.

\humansoldier[\NPC{\glsentrysymbol{sylf}\M\Hu}{Renchkolt}%
  {droopy eyes, unkempt hair}% DESCRIPTION
  {wheels about, clanking sword}% MANNERISM
  {to get out alive}% WANTS
  \npcQuote{Can you go and poke something for me?  Don't worry, the dragon's just a legend}]% QUOTE

\paragraph{If someone locks the door,}
then the \glspl{monster} cannot escape\ldots for a while.
However, a collection of hungry and terrified creatures, with a Strength Bonus of~+5 or more, will not stay contained for long.

\mapentry[fCrossbow]{Siege Crossbow}

It's a fucking massive crossbow, pointed right at the \nameref{fWeaponStore} (\vpageref[next]{fWeaponStore} room).
It takes an age to reload, but it'll deal \dmg{11}~Damage to anything it hits.

To one side sits a barrel of \glspl{torch}.

\mapentry[fWeaponStore]{Weapon Storage}

\begin{exampletext}
  Whenever \pgls{guard} asks about \glspl{weapon}, \pgls{jotter} tells them about the great store of swords, spears, and shields, saved up for a coming raid which will solve all the problems, and make everyone in the land much, much, safer.
  This room holds that grand store of weapons for the up-coming, legendary push-back against the forest, to expand civilization.
\end{exampletext}

\paragraph{Opening the door}
requires an \roll{Intelligence}{Crafts} roll at \tn[12].
But once the door opens, they find nothing.
\emph{The room lies empty.}

\mapentry[fHibernation]{Hibernation Cells}

The \gls{templeOfBeasts} keep hibernating \glspl{monster} in here, and the cold keeps them in hibernation.
The \gls{templeOfBeasts} do not believe they are responsible for maintenance, so they have not fixed the door, which is currently hanging on by a single hinge.
Any vigorous pushing against the door will make it topple.

In total, the chamber contains:

\chitincrawler

\chitincrawler

\umberhulk[\npc{\T[2]\A}{\arabic{noAppearing} \Glspl{hulk}}]

\umberhulk[\npc{\T[2]\A}{\arabic{noAppearing} \Glspl{hulk}}]

\basilisk

\dragon

\showStdSpells

%%%%%%%%%%%%

\mapentry[fCheeseStorage]{Hall of Cheese}

\begin{exampletext}
  The other \glspl{temple} guard their possessions with locked doors.
  The \gls{templeOfFrost} guards their goods a snowman and a lot of lies.
  They tell everyone that the snowman can see them; he's magical, and speaks when nobody's looking.
  And nobody believed them until someone stole some cheese, and they found out by the end of the day.

  To identify the thief, they relied on their standard gossip network, and the fact that it is impossible to come home with a wheel of cheese without anyone noticing.
  What could anyone even do with it in secret?
  Eat it?
  An entire wheel of cheese?
\end{exampletext}

The entrance hall looks like someone carved the walls into a perfect pattern of repeating cylinder-sides.
But in fact the walls just have shelves which bulge with cheese wheels.%
\footnote{If people at your table have previously played the video game \textit{Skyrim}, you will need to gently explain that eating forty cheese-wheels will not close their wounds, nor return lost blood.}

\mapentry[fFabrics]{Fabric Room}

The \gls{templeOfFrost} owns one room, in addition to the \nameref{fCheeseStorage} (\vpageref[next][last]{fCheeseStorage} room).
This place is littered with every type of fabric imaginable, a few barrels of wine, and a few more snowmen dotted around the walls (mostly clothed, other anatomically exaggerated).

\mapentry[fGrain]{Flour Store}

Barrels of flour line the room.
They go on and on forever.

\paragraph{Setting fire to the barrels}
doesn't work, as the small room does not have enough life in its air.
But if the barrels move outside, and \emph{then} catch fire, the dust-fire could leap across the great hall in half \pgls{ap}.

\mapentry[fSpirits]{Spirit Store}

An ominous sign says `\textsc{Spirit Containment}'.
Of course, this means whiskey in every barrel.

%%%%%%%%%%%%%%%%%%%

\mapentry[fLibrarians]{Wolf, Goat, and Three Librarians}

Three \glspl{seeker} each have a pile of their most precious books, and none will let them go.

\begin{boxtext}
  A tubby man, in green robes, steps back as your boat hits the stairs; he holds a pile of books like a baby.
  A tall woman steps back with him, holding her own precious pile.
  The last is a frightened-looking man, with a pointed hat and cubic sack.
  ``\textit{Looks like they're staying!}'', he says.
  ``\textit{Let me hop on\ldots}''
  But the tubby man shouts, ``\textit{No, ladies first!
  Or just let the lady on too!}''.

  But you haven't the room, for books, him and you.
\end{boxtext}

A boat can hold \pgls{seeker} and their books, or two \glspl{seeker} without any books.
But if two \glspl{seeker}, and the \glspl{pc} return for the third, then the boat \emph{will} have room for \pgls{seeker} and their books.
Therefore, all \glspl{seeker} will not enter the boat on the first run, without their books.

\humandiplomat[\npc{\glsentrysymbol{yonder}\M\Hu}{Tubby \composeHumanName}]

\humanalchemist[\npc{\glsentrysymbol{yonder}\F\Hu}{Tall \composeHumanName}]

\humandiplomat[\npc{\glsentrysymbol{yonder}\M\Hu}{Timid \composeHumanName}]

\mapentry[fGiftschrank]{Giftschrank}

Banned but well-worn books and pamphlets remain here, where uneducated people won't be in danger of getting ideas.

\begin{itemize}
  \item
  \textit{Mungrit's Lewd Limericks, Volume IV} (3 \glspl{sp})
  \item
  \textit{On the Metaphysics of Government as a Super Organism} (10 \glspl{gp}, but buyers are hard to find)
  \item
  \textit{The Sexual Lives of Gnolls} (1 \gls{gp})
  \item
  \textit{Tsog's Vile Verses, Volume VIII} (2 sp)
  \item
  \textit{Wars of \Glsentrytext{fenestra}, Vol. II} (1 \gls{gp})
  \item
  \textit{Life in an Elven Stronghold} (2 \glspl{sp})
  \item
  \textit{The Daring Adventures of Scrib the Dullard} (3 \glspl{cp})
\end{itemize}

\paragraph{Unlocking the door}
requires an \roll{Intelligence}{Crafts} roll at \tn[13].

\mapentry[fBrack]{Tat Cupboard}

This place has everything nobody wanted to buy.
Gnomish lanterns, a barrel of gnomish lantern oil, \glspl{torch} made from \gls{basilisk}-fat, soap, tinder boxes (plenty), chalk, magnifying globes (heavy), and one copper bell.

\paragraph{Unlocking the door}
requires an \roll{Intelligence}{Crafts} roll at \tn[13].

\mapentry[fLibrary]{Treasured Library}

This library has the lot, including four smokeless reading lanterns and five small barrels of paraffin.
The books are so fancy that they have jewels in the cover, and contents so valuable that this only \emph{doubles} their price.


%%%%%%%%%%%%

\mapentry[fDoor]{Double Door Ascent}

\begin{exampletext}
  A stairway leads up to the local \gls{sunGuard} station.
  Before the thaw, the \glspl{sunGuard} would fetch frozen beasts, and pull them gently up wooden planks, into this room, and shunt them into the \nameref{fWaiting}.
  Once the beasts woke, it only had one place to go: into the \gls{court}, to amuse the crowds.
\end{exampletext}

\paragraph{Ascending the smooth, slate, stairs}
demands a \roll{Dexterity}{Athletics} roll at \tn[2].
\emph{Every time} that someone walks up with their wet feet, the \gls{tn} increases by~1, to a maximum of 10.

\setcounter{weight}{2}
\setcounter{freeHP}{8}
\Gls{tn}: \hpStat

\paragraph{To lift the wooden bar blocking the double-doors,}
the roll is \roll{Strength}{Crafts} at \tn[7].

\begin{boxtext}
  The great double-doors at the top of the high stairs have a heavy bar over the front.
  It pulls aside with some difficulty, then falls down the steps with a big \emph{thudd} and \emph{massive} splash.

  Pulling open the right-hand door, lantern light touches the room inside.
  It's coming from a gated door in front.

  To the left, two small boats are stacked against the wall.
\end{boxtext}

The \glspl{sunGuard} won't feel happy with any \glspl{guard} entering here, and will tell them to close the door.

\paragraph{If the flood rises to the top,}
water will pour into the \gls{court}.

\mapentry[fWaiting]{Waiting Room}

A recently-thawed \gls{crawler} scurries around the \gls{court} (\vpageref[next]{fCourt} room).
Any vibrations here, even a single footstep, will draw its attention.

\mapentry[fCourt]{\Glsfmttext{court}}

\begin{exampletext}
  This room is empty, but full of anticipation, screaming, cheering, accusations, excuses, and pregnant silences.
  Here, the town's \gls{warden} sits beside a jester, listening to those below.
\end{exampletext}

Currently the crowd stands confused, as the \gls{crawler} came out early, before the accuser could speak.
It really killed the show.

\chitincrawler

\end{multicols}
