\subsection{My First Day}

The module might work for new \glspl{pc}.
Once the \gls{warden} hears of the fridge melting, he makes an expedient pre-court ruling that these new prisoners must join the \gls{guard}, and save the \glspl{monster} at once.

Or perhaps \pgls{pc} has entered the \gls{court} and received a lethal sentence.
But when the \gls{warden} sentences them to death and the gate at the side of the \gls{court} rises, \pgls{sunGuard} pokes his head out and say ``\textit{my \gls{warden}, we have a problem}''.

Or perhaps it works for neither.
Not a single word has been tempered in the fire of the table.

\subsection{Time's Relentless March}

The session should track time from start to finish, because the fridge is melting.
Just measure broad turns with \glspl{ap} as usual, \emph{but} characters spend \pgls{ap} to:

\begin{itemize}
  \item
  Move anywhere in line of site.
  \item
  Get into a boat.
  \item
  Engage in a conversation.
  \item
  Open the heavy doors to room \vref{fDoor}.
  \item
  Engage in combat (even spending 4 \glspl{round} just counts as one turn).
\end{itemize}

Whenever the troupe moves, every \gls{pc} must spend \pgls{ap} or they lag behind.
However, their pace depends on decisions more than the \glspl{pc}' Speed Bonuses.

\begin{multicols}{2}
  \noindent
  The \glspl{pc} may move forward, relentlessly, blinded to their surroundings by the pace.
  \begin{description}
    \item[3 \glspl{ap}]
    Down the stairs (room \vref{fStairs}), they head through the \nameref{fDip}.
    \item[2 \glspl{ap}]
    Some \glspl{pc} enter the boat (room~\vref{fBoat}).
    \item[1 \glspl{ap}]
    They row out to the \nameref{fGuardDoor} (room~\ref{fGuardDoor}).
    \item[3 \glspl{ap}]
    Next turn, they open the door.
    \item[2 \glspl{ap}]
    They move past the \nameref{fCrossbow}.
    \item[1 \glspl{ap}]
    And open the door to room~\ref{fHibernation}.
  \end{description}
  \columnbreak
  Or they may move casually, gaining information and resources.
  \begin{description}
    \item[3 \glspl{ap}]
    Down the stairs (room \vref{fStairs}), go to investigate the \nameref{fHallStore}.
    \item[2 \glspl{ap}]
    Upon investigation, the \nameref{fHallStore} looks dangerous.
    \item[1 \glspl{ap}]
    The troupe checks room~\ref{fLift}.
    \item[3 \glspl{ap}]
    Next turn, they chat with the \glspl{weaver}, above.
    \item[2 \glspl{ap}]
    The \glspl{weaver} might have \pgls{torch} for them.
    \item[1 \glspl{ap}]
    The \glspl{weaver} throw down two \glspl{torch}.
  \end{description}
\end{multicols}

\subsubsection[Upon Entry]{1: Upon Entry}
the water has begun to melt.
The dark centre of the map has waist-high water to a human, and the grey sections of the map (\vpageref{Dyson_Logos/bowels}) have no water yet.

\subsubsection[Water Rises]{4: Water Rises}
through the grey areas, but stops to fill up the \nameref{fDip} in room~\vref{fDip}.

\subsubsection[Oil Spills]{5: Oil Spills}
from the barrels in room \ref{fDip}, and water moves through the rest of the grey areas.

\subsubsection[An Escaped Prisoner]{7: An Escaped Prisoner}
jumps from the \gls{court} and out the doors of room \vref{fDoor}, then heads towards the nearest light (e.g. any \gls{torch} \pgls{pc} carries).
\Pgls{sunGuard} bolts after him a moment later with his own \gls{torch}, and the prisoner has to swim to safety.

The \glspl{pc} might not want to waste their time on this ruckus, but the \gls{sunGuard} wants their boat (he's not getting into that filthy water with his nice, white, vestige).

\humanthief[\NPC{\M\Hu}{\composeHumanName}%
  {one eye, bald}% DESCRIPTION
  {spinning thumbs}% MANNERISM
  {find valuables, then bolt}% WANTS
  \npcQuote{Wouldn't want to join you lads, no offence yea?}]% QUOTE

\sunguard[\NPC{\glsentrysymbol{paik}\M\Hu}{\Glsfmttext{sunGuard} \composeHumanName}%
  {muscular frame, sticky-out ears}% DESCRIPTION
  {stands uncomfortably close while yelling}% MANNERISM
  {to catch the criminal}% WANTS
  \npcQuote{Where is the cretin?
  Give me that boat!}]% QUOTE

\subsubsection[Waist Height Water]{9: Waist Height Water}
means everyone has to spend 2~\glspl{ap} to move when not in a boat.

\subsubsection[Water Moves up the Stairs]{11: Water Moves up the Stairs}
entering the rooms \vpagerefrange{fDoor}{fCourt}, and moves towards the \nameref{fCourt} \vpagerefrange{fGuardDoor}{fHibernation}.

\subsubsection[The \Glsfmtplural{monster} Wake]{12: The \Glsfmtplural{monster} Wake}
and the first moves to the door at room \vref{fGuardDoor}.
The \glspl{monster} also spend \glspl{ap} to move.
