\thread[ftown]{Before the Thaw}

\begin{speechtext}
  \noindent
  They say the fridge began when a dragon attacked, and battled with \pgls{witch} on a rainy day.
  It fell, they say, with enough force to smash \emph{through} the ground and make a hole.
  Then the \gls{witch} froze the rain, trapping it in ice.
  They also say it has a pot of gold somewhere, and has to give you a wish if you answer its riddle.
  `They' are generally full of shit, of course.
\end{speechtext}

\begin{exampletext}
  Everyone in \gls{ftown} relies on the fridge, especially the local guilds.
  It sits underground, accessible from the street, through an open cavity in the ground, down a badly-maintained set of stairs which sometimes slope up, and sometimes slope down.
  Sometimes would-be thieves enter, then wander down some dark and crooked stairs until the darknes and silence make them turn back.
  Very few go all the way down, and even fewer come back up.
  They might c\"oordinate a long passage in the dark, but once into the massive, frosty hall, it's hard to move without a light.
  And if any thief carries a light then everyone else in the fridge can see them.

  Everyone who has business in the fridge knows everyone else who has business in the fridge.
  The \glspl{weaver} go down rarely, to store a cheese-wheel or woolly clothes.
  The \glspl{scribe} sometimes to hide or hunt for dangerous books.
  The \glspl{guard} enter regularly, to check the frozen beasts don't wake up.
  The \glspl{server} enter daily, to check their private rooms, where the flour dries.
  And \pgls{proctor} enters whenever anyone else does, so make sure no more arguments flare up about who owns what, who stole what, and who left a mess.
\end{exampletext}

The \glspl{thread} below establish that the \glspl{temple} in \gls{ftown} argue over the fridge, and hint at the contents.
In \gls{segment}~\ref{meltBegins} (\vpageref{meltBegins}), the fridge begins to melt, the \glspl{temple} stop arguing and unanimously decide that the \gls{templeOfBeasts} has complete responsibility for the fridge, and has to fix the problem immediately.

\ifnum\value{temperature}=0
  \hibernationSegment
\fi

\segment[\squash]{ftown}% AREA
{Juris Diction}% NAME
{Three Guild leaders argue over the dominion of honey}% SUMMARY

\Glsentrydesc{temple}.
So who owns the honey?

\begin{itemize}
  \item
  \Gls{chef}~Ginshank reminds everyong that bees have a poisonous sting, so the \gls{templeOfPoison} --- which protects peoplel from rot by use of \emph{preservatives} --- has exclusive rights to the use of honey.
  \item
  \Gls{notary}~\composeHumanName\ replies that bees do have poison, but `venom', and that honey preserves food, so the \gls{templeOfFrost} has clear and natural dominion over honey.
  \item
  \Gls{counter}~\composeHumanName\ stamps her crutches at every syllable, stating that if anyone claims rights over honey, the \gls{templeOfSickness} won't help them or their families when the time comes to go with \gls{eldren}.
\end{itemize}

If \pgls{pc} takes sides, all \glspl{npc} will remember the decision, and ensure they receive good service or no service.

Begin the next available \gls{segment} in \gls{ftown} immediately.

\segment[\squash]{ftown}% AREA
{Even Lager}% NAME
{\Pgls{chef} has invented a new type of ale}% SUMMARY

\Gls{chef}~Ginshank shouts to everyone that'll hear him about his new type of ale, and hands out free drinks from his barrel.

\begin{speechtext}
  It's ale, but it's not.
  It's gonna be even larger than ale.
  Much larger.
  Have taste and tell your friends!
  It's my secret recipe!
\end{speechtext}

People only receive a thimble-full, but it definitely tastes better than standard ale.
A bit thin, but very refreshing, and strangely cold.

\humandiplomat[\NPC{\glsentrysymbol{abderian}\M\Hu}% Symbol
  {\Glsfmttext{chef} Ginshank}% Name
  {massive hat, grubby overalls, brown teeth}% Description
  {unwanted bear-hugs}% Mannerism
  {to find out how his own ale is made}]% Wants

Unfortunately, Ginshank made the ale by accident after he abandoned it in the fridge, under \gls{ftown}.
Ale doesn't usually ferment in the cold, but this new type ferments very well.

\ifcase\thecycle%
  \hibernationSegment
\or%
\or%
  \ifnum\thefenestraDay>20%
    \hibernationSegment
  \fi%
\else%
  \hibernationSegment
\fi%

\segment{ftown}% AREA
{The Melt Begins}% NAME
{Various guild leaders need the troupe to the problems with the fridge post-haste}% SUMMARY
\label{meltBegins}

The fridge has begun to melt, and the troupe must descend.
\emph{Best run this \gls{segment} only at the start of a session!}

\Gls{chef}~Ginshank has left his new type of ale fermenting at a nice cool temperature in the town's fridge.
This has raised the temperature rapidly, which has then escalated the fermentation process, which raised the temperature again, and makes the barrels of ale give off a lot of flammable gas.
The fridge now contains a mixture of flammable gasses, overly-pressurized ale caskets, and a lot of caskets of dried flour.%
\footnote{Flour is not flammable, but it is explosive.}

\begin{boxtext}
  A lanky man, two \glspl{step} tall, sprints towards you like a beast.
  Two men watch and follow, far behind.
  He stops suddenly, and wheezes through his news.

  \begin{speechtext}
    Yous need to come to the fridge.
    It's melting.
    It's melting and it's full of \glspl{monster} in the farthest section.
    Sleeping \glspl{monster} owned by the \gls{templeOfBeasts}.
    Once they wake, they'll kill all they can.
    Gut them while they sleep in the snow down there, in the room at the end.
    Do it quick, before they wake, or it'll be lives lost up here.

    Yous hear me?
  \end{speechtext}
\end{boxtext}

\Gls{proctor} Cainpelt will lead the troupe down to the fridge (\vpageref{theFridge}) if they ask, but he only joins them as far as his office (\vpageref{fOffice}), then returns.

\paragraph{While they walk,}
\gls{seneschal} Saunter walks briskly to the \gls{pc} at the back of the group, and quietly, rapidly, tells them not to kill the beasts.

\begin{speechtext}
  \small
  Oh, but they're worth a fortune.
  Such a waste!
  Just find what's causing the snow to melt and stop it.
  Don't waste time ruining what could be saved.
\end{speechtext}

\paragraph{Just before the entrance,}
\gls{librarian} Giltvine plants \pgls{gp} in \pgls{pc}'s pocket, along with a note.
Only this \gls{pc} is aware of the note.

\begin{speechtext}
  \scshape
  Save \glspl{seeker} first.
  \Pgls{gp} each.
\end{speechtext}

If the \glspl{pc} manage to save the three \glspl{seeker} (room \vref{fLibrarians}) they receive an extra 2~\glspl{gp} (since they already received one).

