\section{Wrapping Up}
\label{fridgeEnd}

\begin{multicols}{2}

\subsection{At the Exit}
Ten \glspl{sunGuard} stand, ready to kill any \glspl{monster} which emerge.

\sunguard[\npc{\T[10]\glsentrysymbol{paik}\Hu}{\thenoAppearing~\Glsfmttext{sunGuard}}]

\paragraph{\Gls{seneschal} Saunter}
stands behind the \glspl{sunGuard}, angry about any \glspl{monster} which had to die.

\paragraph{\Gls{chef}~Ginshank}
hangs about, just wondering what happened, and if his ale is okay.

\paragraph{\Gls{proctor} Cainpelt}
hopes to hear the troupe completed their jobs okay.
He will support the \glspl{pc}, as long as they did not let more then two \glspl{monster} out, and even point out that they have rights to any \gls{monster} corpses.

\paragraph{\Gls{librarian} Giltvine}
waited for his colleagues (from room \vref{fLibrarians}), and left if they emerged.
He will pay any debts if the \glspl{pc} inquire at \gls{ftown}'s \gls{templeOfCuriosity}.

\subsection{Stolen Goods}

The \glspl{pc} might smuggle out valuables from rooms \ref{fIngStore}, \ref{fGiftschrank}, or \ref{fLibrary}.
The troupe won't sell rare items easily, as anyone interested in the items know exactly where they come from.
Finding an interested buyer needs an \roll{Intelligence}{Empathy} roll at \tn[10].
Selling without getting caught requires \roll{Intelligence}{Stealth} at \tn[10].
The \gls{tn} to sell decreases by 2 in any town beyond \gls{ftown} (but buyers only exist in towns).

These rolls can wait until \gls{downtime}.
And of course, failure means prosecution in the \gls{court}.%
\exRef{judgement}{Judgement}{pitOfJustice}

\subsubsection{But if the Dragon}
flies free, everyone's toast.
It kills the \gls{seneschal} and half the \glspl{sunGuard}.
And it will most likely return to cause more havoc before long.

\end{multicols}
