\section{Monsters \& Malthus}

\begin{multicols}{2}

\subsection*{The Primordial Forest}

\begin{exampletext}

  The forest wants to eat you, so pay attention.

  You've had an easy life, in your secluded, little patch, but here we live above ground, where giant creatures crawl everywhere, and I can assure you I'm the gentlest one out here.
  They're probably looking at you and licking their lips.
Well, not `lips'.
  I don't think a single one has lips.
  Some have beaks, others have `mandibles', but you get my point.

  The chitincrawlers lay webs, but don't think they need to wait for you.
  If they get hungry enough, they'll run straight at you, grab you with their claws, and just start eating!

  And remember to be on the lookout for moving trees.
  If you see something shift that's meant to stay still, it could be a woodspy -- like an octopus, but\ldots do they have octopuses where you're from?
  No I didn't think so.
  But I'm glad to have you with us.
  I hear your people can see better in the dark than we can.

  Most of the world sits in darkness, just like this
  Most of the world lacks roads, beer, beds, and everything that makes life worth living.
  For this reason, we exist, to push back the darkness, and make way for more civilization.

  \subsubsection*{Bandits}

  Notice the trees.
  \ifnum\value{temperature}<1
    They just sitting there in the \seasonDesc n wind right now, but once the warmer seasons hit, they'll be full of fruits, and you'll see things growing all around.
  \else
    There's good eating up there if you can climb.
    The forest is laying a trap for us, but it's a tasty one!
  \fi
  Humans could live out here like some kind of paradise, never working, just taking food from the trees -- at least over C\'alea and Laiqea, and the other warmer seasons.
  Even Toldea has plenty to eat if you know where to look!

  And all this means the forest has laid another trap for us.
  Thieves, cut-throats, and black-alchemists who want to escape the law come into the forest, and she treats them well.
  They live here, tax-free, robbing villages over the cold seasons.

  And obviously it's \emph{our} job to come out here and route them out, by fire and sword.


  \subsubsection*{And worse\ldots}

  Magic's a horrid thing.
  Once someone knows enough of it, they can destroy a city.
  And you can never spot someone who knows it.
  Well sometimes I think you can -- a shifty look in the eye, especially if they've been away from the \gls{alchemists} too long.
  They may as well stay out there, as far as I'm concerned.

  Sometimes we lose whole cities, to curses which bring beasts out of the forests, or maybe they'll just turn a town's walls to ice on a warm C\'alean day and watch them melt while the beasts comes in to eat the village.

  You don't know any magic do you?

  Shame.
  We could probably use some out here.
  At least a little blessing or something would be nice.
\end{exampletext}

\subsection*{The \Glsentrytext{edge}}

\begin{exampletext}
  We're coming up to civilization at last!
  See that patch of lawful land?
  No trees, vines, bushes, or anything?
  That means the town must be close.

  I was in a proper \textit{big burn} once.
  We covered the area in oils last Laiquea \ifnum\value{temperature}=3 on a day as warm as this one \else on a scorcher of a day \fi and it burnt so high I swear it reached \gls{ainumar} and made the gods stink for a week of woodsmoke!

  \subsubsection*{Walled Villages}

  The great clear areas around the \gls{edge} gives us a buffer. 
  Anything that wants to skitter over here in the daylight gets filled with arrows.
  Of course that won't always stop a basilisk, but it can drive them off.

  We'll see some sheep inside, maybe even cows, but there's never much meat here.
  You can take them out grazing only a short way, where it's safe, and the sight of so many animals always tempts something out of the forest sooner or later.
  Mostly, meat comes from the inside, while the outer circles send back wood, or forest fruits, or anything from the fields outside their walls.

  Most humans on the \gls{edge} learn the bow, or at least how to use a crossbow.
  Anyone who doesn't puts everyone else in danger.
  I grew up in a village like this one -- and we had a lot of sleepless nights, telling each other stories of famous adventurers from back when that sort of thing was still legal.

  Simpler times.
  But also dangerous -- we're much safer now with \gls{king} in charge of everything, and don't let anyone hear you say otherwise.
  Any time these places need help, \gls{king} sends us out to help out.

  We can't stay long, so get some rations, and we'll be on the road soon, headed inwards.

  \subsubsection*{Lonely Roads}

  That's another one of your duties, recruit -- maintain the roads.
  Sometimes bandits slip past the outer villages and camp at the side.
  Sometimes critters in the forest do the same.
  A lot of them are smart enough to know where we go, so they'll sit at the side of the road, picking off traders who carry meat, or just any trader.
  Traders \emph{are} meat as far as the forest is concerned.

  I wanted to be a trader once, but honestly couldn't summon up the courage.
  I survived a couple of trips, but I knew I wouldn't make it for long.
  So if you're ever travelling out to the \gls{edge}, remember to let the word around town, just like I did there.
  The more people who travel together, the safer.
  And if you end up getting eaten by something, maybe the beast will leave the trader alone, and let him get to market.
  Then you'll die a hero!

  Everyone dies a hero in the \gls{guard}.

  See that crossroads ahead?
  That's a good sign.
  We passed the \gls{edge}, now we have two roads, meaning at least two villages around us.
  They'll have walls of their own too, but the farther inwards we get, the safer.

  Sometimes these outer roads break.
  At first you notice nobody is coming to visit the town from that direction.
  Then you notice that nobody who went that way returns\ldots
  A couple of weeks later, and people hold the wakes for anyone who went that way, and hopefully they have a different road out.

  When a village just has one road, then no roads, they just sit there like an island.
  Hopefully someone notices, and \glspl{guard} get to them; but until then they live on alone, without iron, coal, or any other help from the outside.
  If that goes on too long, it's another win for the forest, and a loss for civilization.
  We'll keep on pushing out, but when we reach too far out, the forest eats our fingers.

  \subsubsection*{Quiet Hamlets}

  We're getting closer.
  See that little hamlet?
  No walls, or nothing -- just stone houses for emergencies.
  Very little makes it in this far.

  Whether it's beasts or bandits, they get tempted by the smells along the road, and end up in an altercation with one of the settlements further out.
  Even if something nasty made it in here, they eat the sheep before the people.
  Mostly.

  These inner lands provide most of the meat of Fenestra.
  I bet you've even had some back home.
  No?
  Well lets go up and say `hello'!
  Villagers always give hospitality to the \gls{guard} when they see us.

  \subsubsection*{Little Masters}

  Each area has its own master.
  It's not true what they say about humans -- we don't need leaders telling us what to do, but we have them anyway.
  They don't really do much, but I guess they look nice and fancy.
  Village masters own a few villages, and town masters own a town.
  Then they send their taxes back and have a nice dinner.

  Must be nice.
  Pointless, but nice.
\end{exampletext}

\subsection*{Hungry Towns}

\begin{exampletext}

  I suppose you never seen a big city like Arthur's Wing before.
  No monsters live here, so everyone can rest easy, aside from the cutthroats and thieves, who of course have to worry about the likes of \emph{me} dragging them into our merry little crew and our glorious missions.
  Look at that pathetic beggar over there, asking for food.
  He can clearly walk, but refuses to sign up with us and fight for the crown.
  Even if he got eaten by something, it'll slow that something down while everyone else kills it or gets to safety.
  It's a good deed.
  He could be a hero.
  Everyone dies a hero in the \gls{guard}.

  \subsubsection*{Guilded Temples}

  The various guild-temples obtained their monopolies long before the current Rex.
  The priests of V\'er\"e have always taken care of the court houses, and the priests of Ohta have always dealt with weapons.
  And of course, nobody else is allowed to.

  They all support \gls{king}, and he supports them.
  Of course in the olden days -- back when we still had adventurers -- people would rise through divine gifts to be a top priest, and give blessings to all the people.

  Nowadays is different.
  The guild higher-ups just want to make money.
  Some of them still hear the real call, get gifts from the gods, but then their superiors send them out on a grand mission.
  We can't have useful people doing paperwork all day, so I suppose it's for the best.
  And it's good to have them out with us in the \gls{guard}.

  Let's get some rest.
  You've got a mission already.
  Nobody's come from Greenwell in a week, and someone needs to find out why.
  I've found a few other new recruits, so you won't be lonely.

  Time to be a hero.

\end{exampletext}

\end{multicols}

\section{Circles of Civilization}

\begin{multicols}{2}

\subsection{\Gls{shatteredcastle}}

\Gls{king} rules the land completely from every area at once, with the single exception of Liberty.
Not long ago, these lands had separate rulers, and in those times each lord of a land -- large or small -- created a personal militia to deal with problems.
Nowadays, all personal armies have been made illegal.
People may hold weapons, but nobody may have a standing army.

In place of the local militias, \gls{king} has created the \gls{guard} to look after the realm.
Anyone unable to find proper work, goes to work in the \gls{guard}.


\subsection{Temple Guilds}
\label{guilds}

Within every town in Fenestra, divine monopolies are officially enforced.
People must seek legal rulings in a temple of V\'{e}r\"{e}, swords are only sold from the temples of Ohta, and every tavern, at least in some official sense, is a temple to Alass\"{e}.

\paragraph{Alass\"{e}} governs `the ale guild', and all manner of taverns.  Officially, all taverns are temples to Alass\"{e}.

Her journeymen once travelled and sang, but since bards received so many new laws under \gls{king}, they mostly just research better methods for brewing beer.

\paragraph{C\'{a}l\"{e}'s} temples doubles as paper-producing guilds, and provided all townmasters and areamasters with seneschals to count up the lord's holdings and due taxes.
Many are now being replaced with the \gls{king}'s own accountants.

The `Temple of Light' occasionally sends recruits along with the \gls{guard} to find information from beyond the \gls{edge}, especially where a road has broken recently.

\paragraph{Laiqu\"{e}} was at one point in charge of various tasks related to farming.
The temple have since abandoned all such activities, and mostly abandoned any buildings they once held in towns.
The priesthood have stated their intention to work purely on theological matters; as a result they hold the highest portion of efficacious miracle workers.

Most people feel at least a little suspicion from them.
Laique herself symbolizes the forest, and it seems the `temple' never entirely takes humanity's side.

\paragraph{Ohta} rules over warfare and many call the temple `the Sword Guild', as it has exclusive jurisdiction over the sale of all weapons.

\paragraph{Qualm\"{e}} does not deal with much beside funerals, and once dealt with death-payments, made when a murderer must make pay the expected value of the victim to the victim's family.
These services were not popular, and death-payments were soon taken over by the Verean temples.
After that, the church had borrowed too much so the temples were sold or abandoned.
A few remote priests decided to pass into undeath and remain in their abandoned monasteries in a sad, robotic, and bitter state.

\paragraph{V\'{e}r\"{e}} has become central figure of `the Justice Guild'.
People approach the temples of V\'{e}r\"{e} for marriages, court rulings, and to make public business deals.

\subsection{Villages \& Walls}

Villages have numerous ways to stay safe, from staying on small islands to building massive wooden walls.
Many build massive moats around their lands to keep their animals safe, while others keep their animals in barns, and post watchmen with bows to guard them through the night.
Predatory creatures do not always come out during the night, but during the day people have less to fear because they can see danger a long way off, and fire arrows before it arrives.

Villages almost universally cut down all vegetation in the area to give themselves better visibility.
Farther afield, villagers allow trees to grow so they can grow them into the correct shapes for quarterstaffs, or use them to make arrows.
Massive orchards can be left safely outside, as the animals of the forest already have plenty of fruits to eat.

Villages typically surround a town in every direction, meaning that those close to a town or city can rest easy;
any creatures wandering from the forest will typically encounter trouble with those in the outer layers before getting anywhere near the inner circle.
Meanwhile, those poor villages in the outer circle can see a dark, primordial forest every day.

If a village defends itself well, it can grow, and one day may create another village farther out, pushing further into the deep forest.
However, this push-and-pull game does not always go so well for people.
When a village has too many young archers die, or too many livestock stolen to feed itself properly, it can no longer defend itself, and the remaining inhabitants must flee to neighbouring villages, or into a town, where most will have to join the \gls{guard}.

\paragraph{Dwarves} tend to live underground, with tight fortifications, and almost always maintain a direct, safe, tunnel to some nearby dwarvish city.

Despite their relative safety, dwarvish parties must still venture out in order to hunt for more seams, or establish fertile mushroom gardens.

\paragraph{Elves} build small villages almost exclusively.
Each one needs only one or two powerful spellcasters and the rest can remain safe.
The exact magics employed vary from village to village, but they might include a spellcaster who can sense any nearby dangers and incinerate them, or someone who can bless all other villagers with luck when they leave.

\paragraph{Gnomes} tend towards hidden villages, but a few cities remain within Fenestra.
They rely extensively on traps both underground and above ground.

\paragraph{Gnolls} keep plenty of fierce guard dogs around their area to alert them to wandering monsters.
Every gnoll in a village knows they must run and hunt at the first sign of danger.
Gnolls welcome such incursions more than any other race, as they enjoy meats of any creature.


\subsubsection{Chaos at the Edge of Civilization}

Exactly what lies in wait for people outside the small civilized lands depends upon the area.
Mount Arthur has bears, giant arachnids, griffins and more.
The frozen Eastlake area in the North tends to have a lot of undead.
Quennome has every creature one can name, in addition to strange monsters which defy classification.

Long roads, connecting different civilizations, wander through the forest for many miles.
These long roads are only taken by the suicidal, or by groups of armed men.
The exact number of soldiers depends upon the area, but typically six to twelve can keep themselves safe if they take turns at watch during the night.

The forests hold so much edible material -- fruits, vegetables, roots, and game -- that people could live easily within them were it not for the creatures which hunt them.
For this reason, outlaws commonly make little liveable spots, either in a self-made shelter, an abandoned stone building in the forest, or anywhere else they can put up enough of a wall to stay safe.
In this way, any group can keep themselves fed until the food in the local area runs out, or until a cold season hits.
In general, such groups do not have the organizational skills to survive, so they either die one by one, as the forest eats them, or they turn to banditry, and someone comes for their heads.

From the point of view of civilization, the greatest dangers come from any element which can organize the creatures of the forest.
Sometimes this is a necromancer, able to summon the dead, and intent on taking out villages.
At other times, an old elf has become irritated with humanity's encroachment on a nice forest, and decides to organize the creatures of the forest to attack, and trees to grow tall and reclaim the land.
These `forest masters', or `beast masters' pose such a danger that local lords must send specialized hunters after them.
Sometimes a full army will go, but smaller teams are often preferred.
When necromancers kill large armies, the lot can be turned undead, and when priests of the forest sing enchantment spells over a wide area, the extra numbers offered by an army do nothing to help the battle.

\subsection{The Outer Darkness}

While maps of Fenestra, made by men, focus on cities and villages, fields and coppiced trees; the truth is that the world is mostly made of dangerous terrain, running wild with dangerous creatures.

Throughout most of Europe's history, we were the nastiest, scariest things around.
We could wander freely, and use the land as we saw fit.
We tamed forests by cutting down the under brush so we could hunt game more easily, and farm any land we could.
Fenestra, by comparison, sustains a much smaller population per square mile.
The continent remains wild, and only little blotches of civilization exist, with rare roads running between them.

These regions of outer darkness, placed on maps using guesswork and rumours, form the larger part of the world.


\end{multicols}
