\chapter{History}\index{History}

\section{A Creation Tale}\index{Creation Tale}

\begin{multicols}{2}

\begin{exampletext}

\ifnum\value{r4}=1
  \noindent
  In the beginning, we traipsed through mud.
  The world felt weak, and filth clung to our beards.
  So pulled away the muck and filth, creating a great chasm as we searched for something real.
  And at the bottom, we found \emph{stone}.

  We pulled it up, learned to work with it, and shape it.
  We built dry land, then piled it up to create houses, then castles, and structures even greater.
  Weapons came soon after, and the stone let us best all the unarmed creatures of the world.

  We felt ourselves lords.
  Nobody could stand against the might of stone.

  But we boasted too much, and felt too much pride.
  And soon the gods laughed at us.

  The gods shook the earth, and the shaking pulled stone from beneath the muck, more and more, until mountains formed.
  Soon every man and maggot could see great bodies of stone, larger than any which we could have imagined.
  We had lost our edge.

  We continue to live around stone, and work with it well.
  And now we dig up iron, silver, platinum, and gold.
  But we must take care never to boast about them too much, because the gods will laugh at us again, and send down the golden monster, which will pour gold over all of the lands, and turn the hard-earned savings of every clan to nothing.
\else
  \noindent
  Deep beneath the earth, worms the size of many castles roam, occasionally rising to push up the earth into great mountain-like mole-hills.
  These worms were created by the gods.

  Making such big creatures was difficult, so they all had to work together.
  However, the creature proved too complex, so most of the parts were removed before the gods woke it up.
  At first, it had too many eyes, so the eyes were removed, but the gods did not notice that they were already a little bit awake, so they walked away, and mated with each other, and eventually those eyes turned into elves.
  The worm's face looked like a dragon, so of course it had a beard, but the gods decided this was unnecessary and cut it off.
  Where the beard fell, a little dwarf was formed, and quickly escaped by burrowing underground.
  When the creature was seen to be too smart, they removed some of its heart to make it less able to think (because our thoughts and souls reside in our hearts).
  The heart burrowed underground to become warmer, and turned into a little gnome.
  Finally, they decided that its arms were only holding it back in the narrow tunnels it made, so they took its muscular arms off and threw them away.
  They scuttled off and later became strong humans.

  The gods later noticed all the creatures running around the world.
  They wanted one to rule the world, governing it properly, and put fire in their blood so that they could command every part of the world.
  At first the job was given to the elves, but they did nothing but sit about in their forests reading poetry.
  The gods declared that they would no longer rule and at that moment all the elves in the world swore never to recognise them again.
  Later, the gods requests that the dwarves would rule the world, and gave them the gift of runic magic to help rule.
  But the dwarves just mined and mined, ignoring everything which was happening on the surface, favouring exploration of the great worm tunnels.
  Next the gods gave dominion to the gnomes, and to help them govern wisely told them many secrets about how the universe works and how to bend it to their will.
  The gnomes tried in earnest to figure out how best to govern, and the discussion grew and grew but they could not reach a conclusion.
  While the gnomes continue to argue to this day, the gods did not have all day, so they decided it was time to speak directly.
  They selected humans to commune with, speaking with them directly through divine prayer.

  And that, little one, is why we live here, instead of them, and why your dad needs to go and get more land for all the little brothers and sisters you might have later; because if we want to govern the land we have to take it first.
\fi

\end{exampletext}

\end{multicols}

\section{The War of Lies}

\begin{multicols}{2}

\begin{exampletext}
  The war of lies was not a real war, but a simple legal error, compounded until records became a mess.
  Scholars called out for a distinction between fact and fable, which implied a consistent view of history.
  This meant that once a `fact' had been established, scholars changed all other records and accounts of history to become consistent.
  Every new `corrected' book had to make the long journey from city to city, where others would rewrite it, and pass copies along.

  Scholars guessed at what they did not know -- consistency eventually became more important than truth, and every spelling mistake and inkblot which made its way past the rushed editorial process dragged history in a new direction.

  The end of this era came with various stories about how the initial confusion arose, which themselves had to create a consistent structure, so scholars began to explain the fake books entirely by reference to the adventurers who would pay incredible sums for books about long-lost treasures, without knowing if the book was true or false.
  As a result, early history contains more guesswork than real records.

\end{exampletext}

\begin{exampletext}
  After the great elven crime, they began to cover up their lies with myriad false histories.
  Of course, writing simple falsehoods would fool no-one, so instead the elves produced thousands of copies of nearly-correct works, obscuring their own misdeeds in a mountain of conflicting histories, with names changing here, a king swapping his generals, or warriors slotting into the wrong century.

  For their final trick, the elves published various books on how the confusion and lies all began with misguided rules for consistency, which began a fashion among scholars of laughing at the notion that any of their history has to contain `facts', or that it should `make sense when taken with the other stories'.
  The lie has worked, and to this day, nobody can say for sure exactly what the elves did that required so much work to cover up.

\end{exampletext}

\begin{exampletext}
  Many a travelling adventurer crew made their money from researching fallen cities, then raiding locations which had the best loot.
  This produced a market for antiquated books which mentioned golden thrones, magical wonders, lakes of eternal strength, and trapped scrolls which reveal the answer to any question posed.

  Unethical scholars soon began creating a horde of `hidden histories' books, which looked mostly authentic, but contained subtle clues to glorious treasures.
  Of course, claiming the `lost city of gold' was over the hill meant angry adventurers returning soon after.
  The smartest of the con-artists always hinted at treasures lying in the most dangerous areas of the deep forest.

  Scholars mostly invented any historical details they did not know about, but since adventurers never read much about history, the books' accuracy hardly mattered.
  Sensationalism trumps truth when it comes to sales, so the writers relied on creating tales to cater to their readers' prejudices and hopes.
  And when angry adventurers returned from some desolate swamp, empty-handed, the charlatans could always blame their book's errors on `an elvish conspiracy'.

\end{exampletext}

\end{multicols}

\section{The Rise of Rex Dalius Quennome}
\index{Rex!Dalius}

\begin{multicols}{2}
\label{h_dalius}\index{Dalius Quennome}

\noindent The following letter was found in the grave of a women who lived with the elves until her death.
It was taken to \gls{college} as an item of historical interest.

\begin{quotation}

  Dearest Clara, I write to you in haste to instruct you to hide yourself.
  Land Master Dalius of Quennome is mad.
  He he entreated the many elves there to aid him on his mad mission.
  We have upset them too much -- our incursions were only to strike at Dalius, but we have rattled the hornets nest and now the elves mean to wipe us out through assassination and I am sure that Dalius is at the heart of all of this.
  During this season of Qualmea, the elves have so far murdered three town masters and a number of village masters.
  Four more gone missing, and a further three went mad and ran away to join the elves under some enchantment.
  They have mostly been caught and put into a prison, but with the leadership in such disarray, many have joined forces with Dalius of Quennome.
  At this rate we shall all be ruled by elvish masters before long, with Dalius as their pet pretending to lead the show.

  As Land Master of Mt Arthur, you are in grave danger.
  Please make haste to the Shale before the elvish assassins find, then kill or capture, you too.
  Hopefully now that Qualmea is almost ended, the elves will trouble us no longer.

  Yours forever,

  Your faithful servant, Ghent.

\end{quotation}

From the history books of the Quennome region, written by Runcible Shaw, commissioned by Dalius IV of Fenestra.

\begin{quotation}

  After Dalius struck a deal with the elves, they left him to his own devices.
  It was mostly alone that he managed to conquer the Mt Arthur region, and shortly afterwards conquer Whiteplains.
  With half of the area North of the Rexway mountains under his control, Dalius declared himself `Dalius Rex' -- ruler of all of the North.
  The nobles of each region soon began to pay his taxes rather than fight.
  He used his increasing wealth to purchase more suits of armour from the dwarves of first the Mt Arthur region and later from the dwarves of the Shale.
  Before long, he ruled over all the lands of the North.

  Dalius II, was a genius by all accounts, and created the first cash economy.
  He bought coinage from the dwarves and later commissioned additional coinage with his face upon the coin, to show that he would guarantee the money be paid in food if ever the coin was handed in.

  Unfortunately, Dalius II's reign was not easy.
  Earthquakes came, larger and harsher than any which usually come during the stormy seasons.
  They came during Qualmea, at the height of the eclipse, and most of the kingdom was levelled by the earthquakes.

  Dalius II was as good at making friends as his father, and quickly commissioned the gnomes to create portals using conjuration magics.
  With these portals, stone could be transported from place to place quickly and conveniently.
  The portals were also dotted around the kingdom to increase trade.
  Suddenly, people deep in Quennome could purchase fish from the Shale.
  Elves would use the portals to visit the Pebbles Islands.
  Dwarves would come from deep underground to trade their gold, their armour and return with mutton and leather from above.
  Some trade even began with the barbaric peoples of the South.

  This was a time of great expansion, and before two generations had passed, every town had a wall and every noble family had a castle once more.

\end{quotation}

\subsection{The Reign of Rex Hunter Quennome}

\index{Rex!Hunter}
\index{The War of Portals}\label{r_hunter}

From the notes of Hokl of Whiteplains, gnomish chronicler and Master of Forces.

\begin{quotation}

  We told the firyar and their sorcerer king not to play with fire.
  We told them to build their castles, be grateful for the help and then leave off.
  We suffered from the grand earthquake as well as any and remembered it for longer than any of them.  

  They asked us to show them how it was done -- how we opened portals from one place to the next.
  We warned them that a gnome may spend many decades studying such materials and still fail to produce results, yet still they insisted on pushing their young ones to the point of breaking through arduous study so that they could eventually learn our tricks.
  They made a building, ostensibly dedicated to learning of all kinds, but in practice focussed almost exclusively to studying the Conjuration sphere so that they could have more portals.
  It was built next to the only gnomish portal which still stood, and bared their engravings.
  It stands there to this day, in Eastlake -- taller than any building which has gone before.
  They managed to hold a lot of it up with force magic.
  Some say that if the bottom half falls in an earthquake, the top half will stay where it is.

  We have all heard the stories from the dwarves about what happens deep below the earth -- the strange and terrible creatures which live there, always hungry, always eating.
  We knew well enough not to touch them.
  But humans cannot understand something in the abstract.
  You tell them that a horse kicks and they have to smack its arse before they believe you.

  The humans travelled far and wide, and opened up portals to strange other worlds.
  Worlds where dragons rule.
  Worlds where secret things happen which nobody was meant to know about until the right time.
  Worlds of deep, dark, hungry beasts.
  We called them `nura', `the deep ones'.
  The humans opened up portals all over the kingdom, sent in explorers and often had to shut the portals off before anyone could get out again after some strange beasts exited and decided to make a meal of this new land.
  For a time it was chaos, but the chaos was at least largely restricted to Whiteplains.

  Then the device came.
  Some say it was invented by a human.
  Some of the humans say it was invented by a gnome out of vengeance for abusing their art.
  I believe this is wrong but will not pass moral judgement if it is right.
  The device, any time it was used, opened a portal from where it was to the nura world.
  Hordes of creatures which were nothing but mouth and arms came through and ate and ate all they could, and then dragged the humans back through the portals.
  They came back changed.
  The portal could not be closed except by opening another one -- this was the intentional design of the device.

  The device might have been tossed on an island and forgotten about, but humans' first question in regards to every item is how to use it to kill, so they began to make war.
  Troops journeyed South across the Rexway Mountains, including Rex Hunter himself (who gave the mountains that name at the time).
  They made war, and before armies could be rallied on the other side, they left a man with the device to open all the portals he could, one after another, while marching along the mountainside and fleeing before any creature could attack him.

  As portals opened to the nura land, people were pulled back in, kicking and screaming, then ejected out as horrifying giants, three or four times the size of a gnome.
  They would exit and march relentlessly, fall upon a village, eat all the food within (including half of the inhabitants) and then return to the portal, if it was still there, with captives.

  It was theorised that some powerful magic existed in there which forced people to change form.
  When interrogated, it was found to everyone's horror that the ogres which came out could still remember being men.
  They had such hunger that they were forced to persistently raid.
  They were instructed to bring others back by some unknown creatures inside the portal -- some kind of terrible worshippers of dark gods which live under the earth.

  While the nura were ravaging everything in the Southern Kingdom, Rex Hunter continued to wage war, making sure to keep his distance from the exact spot where the nura were active at that time.
  The device was soon captured by people in \gls{outerKingdom}, but instead of throwing it away or destroying it, they decided to take it back and continue the war with their new weapon.
  The war went on, and the only creatures to really benefit were the nura as the device was past back and forth like a ball in some sport.

  Even we were not immune.
  Our many colonies in \gls{outerKingdom} were often targeted accidentally.
  We did not hide in those days but had many structures above the ground.
  The nura came, and took dozens of us at a time, and returned us a horrific little creatures, no longer with a love of growing and learning, but only for eating.
  They are the creatures now known as `goblins'.

  The armies of \gls{outerKingdom} took the device over the mountains, and since the dwarves had made a pact with the King of the North, they attempted to stop the mountain crossing.
  The device was used there, and the armies fled back as raging hordes of nura greeted the dwarves at the entrance to one of their tunnels.
  The dwarves were soon captured en mass and taken into the portal, to return much taller, and stronger, and stupider.

  The two armies battled back and forth, and the device was captured many time by both sides, like some unwholesome football.
  Meanwhile the humans had trained up two alchemists who were able to summon permanent magical portals.
  Rather than use them for trade, they were used to move army regiments in order to better attack \gls{outerKingdom}.

  This was known as the `war of portals', and it from an old word for portal which we derive the name of the North Kingdom as it is known today -- `Fenestra'.

\end{quotation}

\subsubsection{The Epic of Logan}

This is literal, rather than lyrical, translation of the last verses of the Epic of Logan.

\begin{verse}

  His hands finally around the device,

  He had fought many months to obtain it.

  Logan remembered his oath to the king,

  He swore to combat all the enemies of the kingdom,

  But the enemy was Fenestra itself.

  Armies of nura swirled about him

  Afraid of the last touch they had taken,

  But they grew hungry for the device

  There Logan knew what had to be done.

  He entered the portal to the other world,

  And opened it from within.

\end{verse}

\subsubsection{A Recount of the First Logan's Meet}

- From the pen of Carnestel, Elvish Bard

\begin{quotation}

  We really only came together to sing his song.
  He was a popular man so when his epic was finally completed we all came together on the first anniversary of Logan's sacrifice.
  Of course once there, we got to talking -- mainly there were elves and bards at that first recounting as the bards were eager to know the song early as they knew it would be popular everywhere in the land.

  Once everyone was talking, we started sharing knowledge of the magic of song -- how to sing a blessing or strum an enchantment upon people.
  Of course teaching rarely comes for free.
  Bards expect to be paid, even by other bards.
  That night was an exception -- a lot of powerful magical techniques were passed along many over the week's feasting and drinking.

  Around the end of the festivities, word came on the wind that a nearby elven settlement had been attacked by nura hordes.
  Despite the device being gone they continued to come from under the mountain, having made their own portal there.
  That was the only time I know of that elves have been dragged underground to become nura.
  We all feared how they might come out -- many were old and had grown incredibly fast and dangerous.
  It was decided among us that the entire camp of bards would go out and help with our new songs of power.

  The battle was short, but it took half our number and a great many of the remaining elves of the area.
  We destroyed their portal through a great song of undoing which was constructed so that it could only be sung by a choir of six men and six elves.
  Then we collected the dead, and began to mourn.

  During the meeting the various bards had wrapped pieces of their own souls into their songs to animate them and give them magical powers.
  They intended to break those songs in time and regain whatever piece of themselves they had placed into their music.
  But they had died, and would never unmake their own songs.
  They were only songs to summon entertaining illusions, or to cast minor blessings, but they would not be forgotten or misused by any of the survivors.

  We decided then that it would be our duty to fight off nura incursions, or at least to keep an eye on them.
  Many cycles can pass without a meeting, but when a meeting is called, everyone comes.
  It is all done through the power of the spell-songs.
  Let me explain.
  A number of people all know the same song.
  Each one sings the song and activates the spell only once per season -- everyone has his designated day so that they do not break the spell by over use.
  If any one of them ever wants to hold a meet-up, the spell-song is sung until it is broken and the magic is all gone.
  When the others attempt to sing the song, they realise one after the other that the magic is gone and that they must go to the meeting on the next cycle at a prearranged date.

  The humans have all since died of course, but we make sure to teach other humans to replace them.
  We gather and sing songs, we eat and tell the Epic of Logan, and we all exchange any information on encroaching nura, bubbling up from underground.

  Now if you will excuse me, tonight is my night, and I must sing to make sure that my spell-song is still intact and that I am not needed.

\end{quotation}

\end{multicols}

\section{The Rise of Rex Nolan Shale}
\label{nolan}
\index{Rex!Nolan}

\begin{multicols}{2}

\noindent From the banned Histories of the Shale Family, by Rafferty Quennome.

\begin{quotation}

  Legend has it that Nolan of the Shale was the worst dinner guest in history.
  He would spit as he talked and always boasted of either himself, his powerful friends, or how much he resented the kingdom being orchestrated from a castle in Quennome.
  He had some friends among the dwarves of the Shale, mainly because they did not trust the elves.

  He spent a good weight in gold bribing secretive bardic organisations to aid him in sending quick signals.
  The bards have some special magic for it which nobody can figure out.
  He also promised \gls{alchemists} that they would receive extensive funding if they would join him.

  Nolan used portals to march his army quickly to the Rex's castle in Quennome.
  The Rex had become complacent after ruling for more than a century unchallenged.
  The castle was sacked there and everyone inside killed.
  At the point when the king died, the high priest of Qualm\"{e} was staying within the castle.
  He made a curse upon the king's body, so that anyone who died in the presence of the king would rise up as one of the hungry dead and feed on human souls.
  The spell did not stop drive back the attackers unfortunately, it only succeeded in making that castle a cursed place forever afterwards.
  It lies in ruin and rot even now, its road since destroyed, somewhere in the middle of the great Quennome forest.
  Nolan was crowned before the end of the week.

  Nolan's reign was terrible.
  His first act was to begin the building of a castle.
  The central hub was built in Whiteplains and has no door or window, only a narrow hatch to access the top.
  He then commissioned different castle `wings', each in completely different areas in Fenestra.
  Finally, \gls{alchemists} was charged with the task of creating portals from the various `castle wings' to the central hub in Whiteplains.
  By the time of this writing he has succeeded, though he is of course long dead.
  It is known as `\gls{shatteredcastle}', and has pieces and openings all across the land.
  Each piece has one portal which goes to the central hub, so from any castle one can travel, with the king's permission, to any other part.

  The king's next act was to destroy the band of bards who had given him advanced warning of the movements of the Quennome army.
  He managed to kill only a few, as the rest quickly went into hiding.
  To this day, the organisation is illegal and its continued existence is one of the greatest failings of Rex Nolan Shale.

  Finally, Nolan worried about the dwarves.
  They too could set up portals, and had been known to forge alliances between themselves.
  He worried about being surrounded by dwarvish mountains on all sides, so he began a war with them too.

  The war with the dwarves was difficult and protracted.
  They kept blocking tunnels to fend off his attacks, and while they did not like having to make do without being able to trade with humans for their food, they can be stubborn.
  Eventually, however, a peace treaty was signed.
  The dwarves agreed to pay a small tax, to recognise Nolan's crown and to disband all of their portals.
  In return, they would be given cheap foods and enjoy the protection of human armies if the nura ever arose.

\end{quotation}

\subsubsection{The Reign of \gls{king}}

Continued from the banned Histories of the Shale Family.

\begin{quotation}

  \Gls{king} was a more timorous character than his ancestors and would constantly worry out loud about elvish assassins sneaking into the imperial court.
  He designed childish password systems to ensure no shape changers could enter the building without detection, though the passwords were so impractical that most guards just told each other rather than throwing those who could not remember into jail.

  The bards, meanwhile, were still operational and unhappy.
  They started taking gnolls into their ranks and helping them prepare for war.
  They gathered information about all the portal locations they could.
  Gnolls from Liberty (then just called `Dogland') and Whiteplains went out to war.

  At this time, a war had already started in Liberty between the gnolls and the humans.
  The bards only had to pass information on to Whiteplains and secretly take emissaries from one to the other to strengthen ties between tribes.
  The war grew and \gls{king} a fatal flaw in the design of \gls{shatteredcastle}: the portals meant his army could travel anywhere, but it also meant that they could be attacked from all sides at the same time.
  The gnolls pushed in and met in the castle's centre.
  Troops arrived from Eastlake through \gls{shatteredcastle} wing there.

  \Gls{king}, paranoid and short-sighted as ever, could not break with family tradition for a moment and soon after issued another blundering, kingdom-wrecking decree: all magical portals were banned except those officially commissioned by the crown, and all must go to the heart of \gls{shatteredcastle} so that he could control them.
  We could have created a flourishing kingdom with easy transport and excellent trade to every part of the world.
  We could have stopped ten thousand blisters on the feet of all the tradesmen in the realm who wander through monstrous forest Paths to transport goods.
  But \gls{king} could have none of that, so he has sacked gnomish villages and killed alchemists on the suspicion that some new portal is their doing.
  Profit-seeking alchemists, take note! -- flee to a better kingdom, one where your talents can be accepted.

\end{quotation}

\subsubsection{From the Journals of Gregor Liberty}

\index{Gregor Liberty}

\begin{quotation}

  At the time of the tragic passing of \gls{king}'s father, I was stationed as a mere commander of one hundred men.
  I quickly rose through the ranks to take my place by my father's side during the Dog Wars overseen by His Majesty.
  Many battles were planned meticulously -- I oversaw most of them personally.
  I was in charge of pushing back the invasions from Liberty.
  It was there most of the trouble started.
  Pushing gnoll hordes back, time after time, became wearisome, so eventually I happened upon the plan which won the war: we would create a portal in the centre of Liberty and wage perpetual warfare without allowing the gnolls to fall back to safety at any point.
  Naturally, it connected with the Heart of \gls{shatteredcastle}, as all portals must.

  While previously we had been putting captured gnolls to the sword, \emph{I} set them to work.
  Before the war was won we had a flourishing guild all based around the labour of gnolls.
  Their first task was to build another wing of \gls{shatteredcastle} in Liberty.

  As the months went by I built up a healthy guild of gnolls.
  We would take them out, set them to work on someone's castle, or digging some hole, and then charge a very reasonable fee.
  Within a decade we had built up a good civilisation within Liberty, for all manner of families to come and purchase affordable farms.

  Of course, it was not long until trouble started anew as ships from the South came to invade our islands and threaten our shore.
  The War of the Pebbles began before the last gnoll in Liberty was put to use.

  \label{whiteplainsWar}
  The War is almost over, but the crown has further troubles in Whiteplains.
  The nobles there have attempted to erect a magical portal to trade with those in Mt Arthur and refused royal decrees to destroy it.
  \Gls{king} has dealt with the situation, and Whiteplains is to have no more nobility until those people can be trusted to toe the line.
  It shall remain ruled by independent village masters but by nothing higher and a bounty has been placed on the heads of any surviving nobility from Whiteplains.

\end{quotation}

\subsubsection{Current Affairs from the Bard Lennox.}

\begin{quotation}

  Peace!

  Ladies and gentlemen, it has been ten long years of warfare at sea with \gls{outerKingdom}, but at long last we have peace.
  The gnoll raiders in Whiteplains have fallen back to lick their wounds, and the wicked nobles there who attempted to split the kingdom have been vanquished.
  The ships of \gls{outerKingdom} have vanished back to their own ports or sunk in the sea.

  At last, peace reigns.

\end{quotation}

\end{multicols}
