
\ifcase\value{temperature}

  \segment{Town}% AREA
  {Pet Griffins}% NAME
  {The nobles keep pet griffins}% SUMMARY

  The fleeting fashions of nobles really shouldn't have any relation to the \glspl{guard}, but they \emph{do}.
  \Gls{fashionista} kept a pet griffin, in order to show off to other nobles.
  This began a crazy of other nobles keeping a pet griffin.

  The \glspl{pc} hear the locals disagree about this.

  \begin{itemize}
    \it
    \item
    I hear griffins eat the meat for ten mens' meals -- it's an outrage!
    \item
    Lord Cairnhold said that he wants to `tame the forst'.
    He's fighting back, civilizing the beasts!
    \item
    I hear they pay 5 \glspl{gp} for a fertile egg, or baby griffin.
  \end{itemize}

  The rumours about the 5~\glspl{gp} reward are true.

\or

  \segment{Villages}% AREA
  {Warm Coats}% NAME
  {The \glsfmtplural{warden} all want warm coats of griffin feathers}% SUMMARY

  \Gls{fashionista} has found griffin feathers to be very warm, and had a coat stuffed with them.
  The other nobles want a griffin-feather coat, and will pay 5~\glspl{sp} for about one griffin's worth of feathers.

\else

  \segment{Town}% AREA
  {Bloomers}% NAME
  {Young \glsfmtplural{warden} dress in solidarity with \glsfmttext{fashionista}}% SUMMARY

  \begin{exampletext}
    \Gls{fashionista}'s griffin attacked her, exposing her back and some of her bloomers to the attendees at her gala.
    She quickly took our her elven rapier, and (while her servants held the beast down) slew it.
  \end{exampletext}

  Young \glspl{warden} now show their solidarity and daring lifestyles by wearing shirts and dresses with a cut across the back (but properly hemmed) which exposes their own undergarments for the world to see.
  All of their parents hate this.
\fi
