\chapter{Call to Adventure}
\epigraph{
  There was an old lady who swallowed a fly,

  I don't know why she swallowed a fly – perhaps she'll die!
}


\label{callToAdventure}

\input{fey/overview.tex}

\input{fey/snails_attack.tex}

\commentary{
  \begin{description}
    \item[\gls{gm}:]
    The path comes to a cross-roads.
    Do you continue forwards, or go left, or right?
    \item[Player 1:]
    (Sinkmaul)
    Left?
    \item[Player 2:]
    (Mildrain)
    Left\ldots
    \item[\gls{gm}:]
    Now the path has a turn-off to the right\ldots
  \end{description}
}{
  This is awful, and the \gls{gm} should be fed to goblins, feet-first.

  If the players had previously said they want to approach `the tower', then the next part of that process is either arriving at the tower, or failing.
  The \gls{gm} can roll out a long description, to make the length of the march clear, and they can run various distracting \glspl{segment} on the way, but the moment the \glspl{pc} are free to act, the journey to their goal should resume.
}

\chapter{Elven Forests}
\epigraph{
  There was an old lady who swallowed a goat;

  Just opened her throat and swallowed a goat!

  She swallowed the goat to catch the dog,

  She swallowed the dog to catch the cat,

  She swallowed the cat to catch the bird,

  She swallowed the bird to catch the spider

  That wriggled and jiggled and tickled inside her,

  She swallowed the spider to catch the fly;

  I don't know why she swallowed a fly – perhaps she'll die!
}


\label{elvenForests}

\input{fey/ravencops.tex}

\commentary{
  \begin{description}
  \item[\gls{gm}:]
  Some miles on, and the Sun's high.
  You keep walking with that `squelch, squelch, squelch' along the snail-path, and spot a potato on the path ahead, held in someone's arms.
  \item[Player 3:]
  (Cleftbarb)
  Hide!
  \item[\gls{gm}:]
  That's \roll{Wits}{Stealth}, \glsentrylong{tn}~6.
  \item[Player 2:]
  (Mildrain)
  Scared of a potato-wielding marauder?
  Okay then\ldots\dicef{6}.
  \item[\gls{gm}:]
  Hiding in the thick darkness which surrounds the road, you wait until the small figure passes, grumbling to himself in the \gls{tradeTongue}.
  Once silence returns, the march continues, through hours of nearly-identical woodland.
  By the time you approach the tower, it's nearly night, but the dusk's light highlights a stone structure in the forest.
  A long cube, a rumbling stone box.
  \item[Player 2:]
  (Mildrain)
  Ignore!
  \item[Player 1:]
  (Sinkmaul)
  Right, `tower'.
  \item[Player 3:]
  What if it's treasure?
  \item[\gls{gm}:]
  The road opens, revealing a shining lake, with the tower in the centre\ldots
  \end{description}
}{
  The pacing sped up suddenly as the players hopped through three \glspl{segment} within a couple of minutes.
  The players still received information (giant snails reduce the \gls{monster} population), gained questions (though nobody has actively asked about the source of the giant potato), and have a new location on the map which they might return to.

  That last \gls{segment} with the `stone box' (sepulchre) has the `\gls{vlg}' symbol listed next to it.
  This indicates it should go onto the map, with the assumption it's always been there.
}

\input{fey/oathtower.tex}

\commentary{
  \begin{description}
  \item[Player 1:]
  Okay, so the elf at the quarry asked us to deliver this letter, and showed he knows some kind of levitation spell, now this elf is growing floating plants.
  Is she the one?
  \end{description}
}{
  The eye of the story moves North, and finds a stone sepulchre.
  Then it moves West and finds a quarry, surrounded by giant snails.
  The land seems full of life, because the \glspl{sq} from \autoref{elvenForests} place the living things in front of the players.
}

\input{fey/sunway.tex}

\input{fey/shadepaths.tex}

\begin{multicols}{2}
\input{fey/snail_tamer.tex}
\end{multicols}

\input{fey/plateaus.tex}

\chapter{Loose Threads}
\epigraph{
  Can you name the nameless one?

  Can you shoe a snail?
}

\label{looseThreads}


\section{Meandering Tails}

\begin{multicols}{2}

\input{fey/rnj.tex}

\input{fey/ranger.tex}

\input{fey/enchanted_enc.tex}

\input{fey/garden_enc.tex}

\stopcontents[sq]

\end{multicols}

%\input{fey/mound.tex}

\input{fey/conclusion.tex}


