\thread[ravencops,oathtower]{Events in \glsfmttext{enchantedLands}}

\segment[\gls{afternoon}]{ravencops}% AREA
{Silencing the Starlings}% NAME
{A goblin takes aim at a starling for the crime of high-pitched song}% SUMMARY

\histEvent{39}{3}{%
  Soon \glsfmttext{enchantedLands} filled with \glspl{sepulchre}, and every high-pitched noise woke the sleeping \glsfmtplural{ogre} inside.
  \Glsfmttext{MindElder} banned all high-pitched noises, including most birds.
  Of course, the ravens and crows remain, which started the name `\glsfmttext{ravencops}'%
}

\begin{exampletext}
  When \gls{MindElder} banned high-pitched voices, this included most birds, because bird-song can wake the sleeping \glspl{ogre}.
  The ecosystem has become strange since then, as it has very few birds, except ravens, crows, and magpies.
  This is why people call the local forest `\gls{ravencops}'.
\end{exampletext}

Mora has heard a starling sing, and thought to herself `that's illegal!', then readied her \gls{projectile}.
After a `\emph{thunk}!', the bird lies dead, so she finishes the job with a rock, and sucks out the starling's brains.

Mora will happily speak with the troupe, in a low-pitched voice (to avoid waking any \glspl{ogre}), but if she sees them doing something criminal (like carrying weapons without a licence) she flees to sound the alarm (but quietly).

\enchantedGoblin[\npc{\F\N}{Mora}]

\segment[\gls{morning}]{oathtower}% AREA
{Meat Salad}% NAME
{As a snail approaches the tower, \gls{MindElder} turns its mind towards thoughts of meaty salad}% SUMMARY

\histEvent{95}{3}{%
  When giant snails barged into \glsfmttext{enchantedLands}, \glsfmttext{MindElder} decided to kill two birds with one stone, and twisted their little minds to crave flesh.
  The giant snails stalk the woods, looking for some meat to eat with their leaves, and occasionally find \glsfmtplural{crawler}%
}

A giant snail approaches \gls{oathtower}, along a snail road, on a clear day.
The \glspl{pc} probably won't see the snail, unless they're on the road out, but they will certainly see \gls{MindElder} observing the land from the balcony at the top of \gls{oathtower}, and singing a spell to twist the mind of the snail.

\spell{Quamahta}% Name
  {Detailed, Distant}% Enhancements
  {Warp}% Action
  {Fate, Water}% Spheres
  {fullness of snail belly}% Resist with
  {The caster describes meat-based salads, and a target snail at \spellRange\ begins to hunt for crawling things in the bushes as much as they do the bushes themselves}% Description
  {They don't eat people who are not surrounded by bushes (because they expect meat only in regular food).}

\paragraph{Any characters who understand Elvish}
will understand the song relates to a meat-based salad, and that the snail should add meat to its salad.

\segment{ravencops}% AREA
{Shoeing a Snail}% NAME
{The goblins hound a snail into a dead-end to kill it}% SUMMARY

Mora shoes a snail into a dead-end road.
She organizes a dozen goblins to spread out, and occasionally stab it (while keeping their distance), in order to hound it to the clearing where six more goblins wait with spears on the road.

The scene proceeds just as the game-mechanics suggest.
Goblins throwing javelins deal around $1D6-1$ to $1D6+1$ Damage, and the giant snails have a lot of \gls{dr}, even in their most vulnerable location.
Some javelins hit the shell and shatter, others stick into the snail's `skin' harmlessly, and a few dig into its body enough to inflict a minor wound (perhaps 1 or 2 Damage).

Bringing down the snail will take the entire \gls{interval}, as the goblins run away, some double back to collect javelins which fell on the ground, and others run ahead to climb trees and throw javelins from above.
The goblins have to constantly stop the snail entering the forest.

\paragraph{Once the snail dies,}
the goblins take it apart in three stages.

\begin{enumerate}
  \item
  The barbecue, as every goblin feels famished, and must eat immediately.
  \item
  The dissection, where bloated goblins with pot-bellies cut and cure the snail-meat, then make ropes from its innards.
  \item
  The great smashing, where they use rocks to crack the shell into smaller dishes, then use the dishes to transport remaining meat back to the Icebox House.
\end{enumerate}

Each stage takes another \gls{interval}, so the troupe may see the goblins again if they pass through the area the next day.

\segment[\gls{afternoon}]{ravencops}% AREA
{On the Menu}% NAME
{Another giant snail approaches, but ignores anyone on the road}% SUMMARY

Another great snail approaches, with its mind focussed on a meaty-salad.
It ignores anyone on the road.
It only attacks living things in the bushes where it feeds.

The snail stops here and there, inspecting the bushes, and sometimes vomiting on them as part of its external digestion, but never stays for long.

A biting wind blows, bringing one extra \gls{mp} at the end of the \gls{interval}.

\segment{oathtower}% AREA
{Promises, Promises}% NAME
{Goblins gather to renew their oaths}% SUMMARY

Twelve goblins solemnly approach \gls{oathtower} to renew their oaths to uphold the laws of \gls{ravencops}.
They stand at the edge of the lake and sing aloud, while \gls{MindElder} enchants their oath.

\label{ravencopsLaws}
\begin{speechtext}
  We shall not shriek nor giggle, \\
  We silence high-pitched sounds. \\
  And when we find a screamer, \\
  We stop the screaming sounds. \\
  \\
  We tell no jokes nor funny tales, \\
  We quash what makes folk laugh. \\
  And if we hear a singer, \\
  We check they sing in key. \\
  \\
  No weapons without licence, \\
  No running without clothes, \\
  And when we spot a criminal, \\
  We stop the crime with glee!
\end{speechtext}

\segment{oathtower}% AREA
{Cry for the \Glsfmttext{basilisk}}% NAME
{Goblins fight to the death over the honour of eating a complete \glsfmttext{basilisk}, and another \glsfmttext{ogre} is born}% SUMMARY

A goblin smells \pgls{basilisk}, and gives the traditional low-pitched cry.
The others echo it along the path, until all the goblins in the forest are giving the low, crow-like groan.
Soon, those near the \gls{basilisk} shoot arrows in the air, or throw rocks -- anything to draw attention to the location.

Back in \gls{oathtower}, goblins frantically run up and down the stairs, pointing out the window and staring, hungrily.
The hobgoblins are the worst; they leave immediately, and the rest leave soon after.

Soon, \gls{oathtower} becomes silent.
None remain but the elves.

If the \glspl{pc} join the goblins, they see the horde descend upon the \gls{basilisk} with no thought for their own survival.
A hobgoblin from \gls{oathtower} strikes the killing blow (which gives them the right to eat the entire corpse).
Other hobgoblins contest the kill (as is tradition), which prompts a bloody fight.

\basilisk

Once a single hobgoblin remains, it eats through the lizardy-corpse while surrounding goblins try to steal bits of meat (as is tradition).
The feasting takes a full day, and the hobgoblin must eat through the night.
The area becomes covered in shit quickly as the goblin metabolisms devour the corpse at rapid speed.

By the end, a newborn \gls{ogre} stomps back, its stomach bloated with meat, and sleeps an enchanted sleep in \pgls{sepulchre}.
\Gls{MindElder} gives a short speech about the honour of protecting the land, and the grand feast which the \gls{ogre} will awaken to.

\segment{ravencops}% AREA
{The Unmerry Band}% NAME
{A dozen goblins patrol the lands, and their three leaders bicker, insinuating the others have committed crimes as a way to assert dominance}% SUMMARY

Grawl, Majiscule, and Brev patrol the land, looking for trouble-makers, mapping new paths the snails have brought, and noting local monsters.
And as they walk they bicker; Grawl accuses Majiscule of jokes.%
\footnote{\Glsentrytext{MindElder} banned jokes to stop the high-pitched goblin laughter waking the \glsfmtplural{ogre} sleeping in their \glsfmtplural{sepulchre}.}

\begin{speechtext}
  If we find a snail, we'll need to go back to get the others, and shoe it towards a dead-end.

  \textbf{How do you `shoe a snail', Grawl?}

  Is that a \underline{joke}?
  Are you trying to be \underline{funny}?

  \textbf{You said it!
  You said we might shoe a snail!
  You were making the joke, not me!
  Confess!}

  Your face is a joke.
  It is a crime, your face.
  You have a funny-looking, criminal face.
\end{speechtext}

Majiscule hides his face until nightfall.

\enchantedGoblin[\NPC{\T[4]\F\N}{Grawl}%
  {blotches of green skin sprout moles, which host long, thick, black hairs across her back and neck}% DESCRIPTION
  {stares over your shoulder, like there's something there}% MANNERISM
  {to lead the group, unchallenged}% WANTS
  \npcQuote{You being\ldots \underline{funny}?}]

\enchantedGoblin[\NPC{\T[4]\M\N}{Majiscule}%
  {shaking from hunger}% DESCRIPTION
  {strokes and fondles food, down to the last scraps}% MANNERISM
  {to lead everyone to food}% WANTS
  \npcQuote{watcha looking at?}]

\enchantedGoblin[\NPC{\T[4]\M\N}{Brev}%
  {head larger than his shoulders, like an ice-cream cone}% DESCRIPTION
  {nasal voice, due to index finger shunted down his nostril}% MANNERISM
  {deer with cheese}% WANTS
  \npcQuote{Don't you \underline{dare} pull my finger}]

\paragraph{If the \glspl{pc} gift the goblins a couple of \glspl{ration},}
they calm down, and stop bickering.
And if the \glspl{ration} are foreign (e.g. human food), they start to ask every details about the origins of the food.
It soon becomes apparent that a fascination has grown in the goblins for that particular taste, and they would like to taste it again.

\paragraph{If the goblins continue hungry,}
they become tetchy, then raspy, then painstakingly polite, then violent (the standard four stages of goblin aggression).
If the \glspl{pc} irritate them at all, the goblins attack.
Otherwise, the goblins eventually turn on each other, viciously, leaving half of the goblins dead, and the other half badly wounded.
Any \glspl{sepulchre} within a mile hear their squealing and screeches, and the \glspl{ogre} sleeping inside awaken, and begin their breakfast-chants.
