\sidequest[shadepaths,sunway,plateauGardens]{An Ordinary Week in \Glsfmttext{sunderedForest}}

\sqpart[\gls{morning}]{shadepaths}% AREA
{The Picnic Choir}% NAME
{Elven songs descend from the plateau above}% SUMMARY

A little group of elves sing a love-song to the clear-blue skies, but don't respond to shouts or calls.
They will, however, respond to someone singing with them.

If someone sings well, they will lower the vines of a broken bridge to help them up.
If they sing poorly, the elves leave silently.

Singing \glspl{pc} should roll \roll{Intelligence}{Performance} (\tn[10]) to understand the elven song, and imitate it.

\sqpart{plateauGardens}% AREA
{Interview with the Tao Mistress}% NAME
{\Glsfmttext{LifeElder} wanders \glsfmttext{shadepaths}, and her answers seem strange}% SUMMARY

\Gls{LifeElder} wanders, and sometimes sings in a way that somehow matches the breeze.
She stops to ponder a flower, then alters a seed so the flower will grow purple.

While the \glspl{pc} are in \gls{plateauGardens}, they see a small, red-haired elf below, wandering naked and humming to herself.
Other elves may identify her as the source of all the change in the landscape, but will not give her a name (except to say `Hi').

She speaks quickly, and cryptically, and gives deep thought to every word someone says, but quickly tires of conversation.

\begin{itemize}
  \item\it
  What's your name?
  \item[\adforn{54}]\bf
  I'm not really into labels.
  People just say `hi'.
  \item\it
  Did you make the snails?
  \item[\adforn{54}]\bf
  Nobody can make a snail.
  Snails are completely impossible, unless you have a snail.
  Is that a paradox?
  \item\it
  Did you make the snails grow big?
  \item[\adforn{54}]\bf
  No, but I suggested it.
  Making someone do something sounds violent.

  Have you had a ride on a snail yet?
  \item\it
  Could you stop making large snails?
  \item[\adforn{54}]\bf
  Certainly.
  I've not been doing a lot of things, what's one more thing to not do?

  But who will all the little biters to the East eat?
  \item\it
  I don't care what the goblins eat, could you just stop\ldots wait did you say `who'?
  \item[\adforn{54}]\bf
  No, wrong word.
  I haven't spoken Gnomish in a while.

  It should be `whom', shouldn't it?
  \item\it
  We're human.
  Wait! what do you mean `whom the goblins will eat'?
  \item[\adforn{54}]\bf
  You speak Gnomish well for a human.
  You know so much!
  \item\it
  Thanks.
  Listen.
  I know you're busy, but\ldots
  \item[\adforn{54}]\bf
  Come back when you know nothing, and I will teach you nothing.
\end{itemize}

\LifeElder

\showStdSpells

\sqpart{shadepaths}% AREA
{Misty Way}% NAME
{The mists fill the causeway}% SUMMARY

The \glspl{pc} can see nothing in the causeway, as mist hangs low.
They may have to make a navigation roll just to move about, and projectiles suffer double the normal range penalties.

\sqpart{plateauGardens}% AREA
{Mist Below}% NAME
{\Glsfmttext{shadepaths}, only eyestalks roam above}% SUMMARY

Mist always falls into \gls{shadepaths} by \gls{plateauGardens}.
The disorienting environment is quite lethal to newcomers, as people who don't know where the land lies can easily make a misstep, tumbling down the sides.

\begin{boxtext}
  Mist has risen, but is falling into \gls{shadepaths} beside the plateau gardens.
  Soon \gls{shadepaths} look like fluffy-white rivers.
  A figure in the distance seems to hover between two plateau gardens, but the walking reveals they must be on a vine-bridge which has fallen just below.
\end{boxtext}

Fast movements require a \roll{Wits}{Survival} roll (\tn[8]) to avoid falling into the misty canals below.
Longer journeys require an \roll{Intelligence}{Survival} roll (\tn[10]).

The mists fade after \pgls{interval}.

\sqpart{sunway}% AREA
{A Voice from on Hi}% NAME
{\Glsfmttext{LifeElder} looks down at the party, ready to converse again}% SUMMARY

While the troupe wander through \gls{shadepaths}, \gls{LifeElder} has been searching \glspl{plateauGardens} for \glspl{ingredient} to make \glspl{boon} so she can make more giant snails.

\begin{boxtext}
  A voice from above floats down into the shadows.
  The grey-haired elf looks down, eyes wide-open, and she says ``\textit{I wonder if you found enough food}''.
\end{boxtext}

As before, she remains distracted and indifferent, but potentially helpful if the \glspl{pc} seem peaceful.

Here are the kinds of things she says:

\begin{itemize}
  \item[\adforn{54}]\bf
  What is a human's favourite type of rain?

  Or do humans like all rain equally?
  \item[\adforn{54}]\bf
  Endings have nothing to do with what happens.
  An ending is ultimately a manifestation of values.
  \item[\adforn{54}]\bf
  If it never rained, the plants wouldn't grow.
  \item[\adforn{54}]\bf
  Possessions are just \emph{things}.
  Don't let your things control you -- be free!
\end{itemize}

\paragraph{She leaves soon after,}
hoping to gather \glspl{ingredient} to make 2~\glspl{boon} so she can make more giant snails.

\spell{Hyalmahta}% Name
  {Detailed, Devious, Duplicated}% Enhancements
  {Wax}% Action
  {Earth, Water}% Spheres
  {available plant quantities}% Resist with
  {The caster thinks about past dreams of salad and \arabic{spellTargets} snails in the vicinity begin to eat and grow.
  Within a few days, they reach the size of a human, and after one \showOnset\ reach the size of a house.
  Each snail has Strength~+5, Speed~-3, \gls{dr}~5 (or 10 on the shell), along with the abilities Viscid and Acid Spray}% Description
  {}

\sqpart[\gls{vlg}]{plateauGardens}% AREA
{Drooping Bean Vines}% NAME
{Bean vines on a tree provide a route down}% SUMMARY

The bridge-vines don't always grow into bridges.
This one just climbed a tree, which is now covered in bean-producing vines.

If the troupe pull the long vines down, they can safely descend to \gls{shadepaths} below.
However, leaving the vines there means creatures below can crawl up; so if the troupe use the vines to descend, the next time they arrive here, they see \pgls{crawler} ascending to the plateau garden, and assaulting anyone there.

\sidequest[shadepaths,sunway,plateauGardens]{Grumbling Gardens}

\sqpart{shadepaths}% AREA
{Excuse Me!}% NAME
{A giant snail blocks the path}% SUMMARY

`Brownie' the snail grew much bigger than the others, and pregnancy did not help.
Now when she moves through \glspl{shadepaths}, she blocks them entirely.
Anyone walking inside \glspl{shadepaths} must simply turn back and find another route.
This may delay the troupe in whatever they wanted to do by \pgls{interval}.

If the troupe try to squeeze past her, have them roll \roll{Dexterity}{Stealth} (\gls{tn}~8 plus the character's Strength Bonus).
Failure inflicts $2D6$ Damage as her massive body smooshes them against \gls{shadepaths} walls.

\sqpart{sunway}% AREA
{Stampede}% NAME
{An auroch stampede passes through}% SUMMARY

Elves watch from a garden plateau as aurochs stampede through the causeway.
Behind them, \pgls{crawler} chases, but it's losing steam, and soon retreats into the forest.

\sqpart{plateauGardens}% AREA
{\Glsfmttext{woodspy} Spotted}% NAME
{\Glsfmttext{woodspy} stalks the garden}% SUMMARY

Most large animals have trouble reaching up the tall walls to the plateau gardens, but \glspl{woodspy} are clever, and sometimes manage to pull themselves up using a snail, or finding a piece of vine hanging down.
This \gls{woodspy} has reached the \gls{plateauGardens}, and stays close to a vine-bridge, to wait for someone to pass.
Once it grabs someone, it pulls them down onto \gls{shadepaths}.

If the \gls{woodspy} attacks \pgls{npc}, you can resolve the fight by just comparing each \gls{npc}'s \gls{cr} (highest wins), instead of rolling dice alone.

\sqpart[\glsentrysymbol{day}]{sunway}% AREA
{Slow Wander Home}% NAME
{The aurochs are returning\ldots slowly}% SUMMARY

Aurochs move slowly through a mild breeze, grazing at \gls{sunway} grass as they go.
Any sudden movement makes the aurochs flee, creating a stampede.

\paragraph{If the troupe wants to pass quietly,}
have them roll \roll{Dexterity}{Empathy} at \tn[8].

The troupe may prefer to wait, but this will take the rest of the \gls{interval}, as the aurochs move slowly, while grazing.

\sqpart[\glsentrytext{evening}]{sunway}% AREA
{Full Moon \& Sacks}% NAME
{Goblins approach the \glsfmttext{plateauGardens}, sniffing loudly for food}% SUMMARY

A guilt of goblins approach \gls{plateauGardens}, two sacks each, looking to steal vegetables.
They spent their time searching the walls before finding roots to climb, slowly up.
Once up they split and search, sniffing so loud it almost sounds like a whistle.
And soon they regroup, as Glottal found some carrots.

\paragraph{If the \glspl{pc} stop them,}
they object.

\begin{speechtext}
  Go fuck a badger, we came for grub!
  It's not `theirs', okay?
  They said the trees belong to the ground, so now they're ours.
  They said all good tea is theft, okay!?
  So we have not broken our oaths.
\end{speechtext}

\enchantedGoblin[\NPC{\F\N}{Glottal}%
  {Long nose, stubby fingers}% DESCRIPTION
  {Fingers her sack}% MANNERISM
  {to find the longest carrot}% WANTS
  \npcQuote{Off and get your own nobody's-carrots}]

\enchantedGoblin[\NPC{\M\N}{Grawl}%
  {Emaciated and agitated}% DESCRIPTION
  {Claws cheeks down till the eye-veins show}% MANNERISM
  {eat or fight, now, now, NOW}% WANTS
  \npcQuote{They said `proper tea', not `good tea', idiot}]

\paragraph{If the \glspl{pc} ask the elves,}
the elves say they don't feel happy with goblins taking their food, but refuse to acknowledge the concept of property.

