\section*{Broch}

\printcontents[Broch]{f}{2}{\setcounter{tocdepth}{3}}

\section*{Ravencops}

\printcontents[Ravencops]{f}{2}{\setcounter{tocdepth}{3}}

\section{Oathtower}

\printcontents[Oathtower]{f}{2}{\setcounter{tocdepth}{3}}

\begin{multicols}{2}

The Oathtower stands in a lake, with reeds growing from it.
The local elves use the reeds to make paper.

\begin{boxtext}
  The tall, wooden, door has an identical shade of grim-brown to stone walls of the tall tower.
  The tower stands a stone's throw away, at the centre of a lake, brimming with leaves.
  A single tall figure in a green tunic stands in a boat by the tower.
  Her ears are long, her skin maggoty-white, and her eyes full of suspicion.
\end{boxtext}

The door inside leads to a hall with a staircase and three doors.
Each floor is identical, so if the \glspl{pc} enter a door, roll to determine what's inside the first three floors:

\begin{dlist}
  \item
  1D6 hobgoblins with a deck of cards, debating the importance of rules in games.
  \item
  1D6 hobgoblins mending a boat for the lake around the Oathtower.
  \item
  Hobgoblin weapon storage.
  \item
  A kitchen with 1D6 elves.  They debate politely, but their words hide fierce insinuations about how the others once burnt an egg, or used the wrong type of wood for cooking snail meat.
  \item
  Light supplies, with 1D6x3 torches, and 1D6x4 candles.
  \item
  Meat storage, with 1D6 meals' of \gls{crawler} eggs in salt, 1D6 meals of snail-meat, 1D6 meals' worth of vegetables (half of them gigantic and stolen by goblins from the garden plateaus), and 1D6 meals of auroch meat.
\end{dlist}

Storeys 3, 4 and 5 have different contents:

\begin{dlist}
  \item
  A small study, with books on poetry, and fantastical erotica where rabbits ride foxes to battle snakes.
  \item
  A cupboard of \glspl{ingredient}, including 1D6-3 woodspy beaks (Water \glspl{ingredient}), 1D6-3 \gls{crawler} spinnerets (Fate \glspl{ingredient}), and 1D6-3 phials of human blood (also Fate \glspl{ingredient}).
  \item
  A lounge, with 1D6-3 elves lounging.  These elves have taken so many oaths concerning good behaviour that they are practically incapable of action, and need goblins to tend to them daily.
  \item
  A toilet which empties into the lake below.
  \item
  Romeo's bedroom (with three books on poetry placed randomly around the room).
  \item
  \gls{MindElder}'s bedroom (with a balcony, a bed, and nothing else).
\end{dlist}

\end{multicols}

\section*{Causeway}

\printcontents[Causeway]{f}{2}{\setcounter{tocdepth}{3}}

\section*{plateauGardens}

\printcontents[plateauGardens]{f}{2}{\setcounter{tocdepth}{3}}

\section*{Canals}

\printcontents[Canals]{f}{2}{\setcounter{tocdepth}{3}}

