\section{\Glsfmttext{oathtower}}
\label{oathtower}

\Gls{oathtower} stands four storeys tall in the centre of a shining lake.
People can see the tower's top for miles around wherever the tree coverage is not too thick.

\Gls{segment} set by \gls{oathtower} begin when the \glspl{pc} \emph{approach} -- they do not have to enter the tower.
Many of these \glspl{segment} concern the nearby forest, or events which can be seen or heard once close to the tower.

Nobody near \gls{oathtower} can breathe a wisp of mana, because mana always heads towards the largest vacuum.
And since \gls{MindElder} spends so many \glsentrylongpl{mp} each day, he usually has the most missing by the end; so everything flocks towards him, leaving the air feeling stagnant even in a storm.

\printThreadsInRegion{oathtower}

\begin{multicols}{2}

\subsection{The Lake}

\noindent
Whistling cane grows around the South side of the lake, giving it an eerie sound whenever the wind blows.%
\exRef{judgement}{Judgement}{whistlingCane}
The elves use it to make paper, while the hobgoblins use it as an instrument.

\begin{boxtext}
  \Gls{oathtower}'s tall, wooden, door has an identical shade of grim-brown to its stone walls.
  The tower stands a stone's throw away, at the centre of a shining lake.
  A single, short, figure in a green tunic stands in a boat by the tower.
  Her ears are long, her skin maggoty-white, and her eyes full of suspicion.
\end{boxtext}

\subsubsection{Parley with the Boat Goblin}
is the only means of communication with \gls{oathtower}, and takes \pgls{interval}.
She will row across, listen to the \glspl{pc}, and then respond to \emph{every} question with ``\textit{I dunno\ldots I'll ask}''; then she rows back to \gls{oathtower}, walks up the stairs (the \glspl{pc} see her passing by a window), asks \gls{MindElder}, then returns with the reply.

\paragraph{If the \glspl{pc} want information}
then \gls{MindElder}, or the elves in room \vref{towerKitchen} furnish Ha\^{c}ek with replies.
The replies will be accurate, but curt, with no additional information.

\paragraph{If the \glspl{pc} ask for anything more}
(such as weapons' licences or entry to \gls{oathtower})
\Gls{MindElder} stands on his balcony (noted \vpageref{MindElderRoom}) and calls down to them, asking them about their own intentions, and asks them to promise to obey the local laws.
The \glspl{pc} should suspect his request has cast \pgls{spell} on them, and you should ask them to roll \roll{Wits}{Academics} at \tn[13].
However, a successful roll lets the character know that \gls{MindElder} has cast nothing at all (he's nearly out of \glspl{mp}, and too busy to bother with the \glspl{pc}).

\enchantedGoblin[\npc{\F\N}{Ha\^{c}ek the Boat Goblin}]

\paragraph{If the \glspl{pc} break the local laws here,}
\gls{MindElder} will step onto his balcony, and begin casting offensive spells.

\subsection{Stony Spiral}

The tall structure's stone floor, wooden doors, pitch \glspl{torch}, and loitering hobgoblins give the false impression of a classic `dungeon'.
However, everyone in \gls{enchantedLands} keeps their promises, so they have no need for currency, or locks.
Every door opens at a push, but the only chests contain bottles of brandy or hobgoblin underwear, waiting to be washed.

\subsubsection{The entrance door}
opens into a hall with a staircase and three doors.
If the troupe enter, Majuscule the hobgoblin asks them to place their weapons in room \vref{towerWeaponStorage}.

\begin{description}
  \item[The ground floor]
  has a grand hall with two doors and a spiral staircase.

  In the centre of the hall, $1D6$ hobgoblins sit with a deck of cards, debating the importance of rules in games.
  The hobgoblin guards have \gls{armour} made from \pgls{basilisk}'s hide \gls{covering} their torso, and a mottled-brown helmet made from the shell of a giant snail.
  \begin{enumerate}
    \item
    \textbf{Banging noises} echo from $1D6$ hobgoblins mending a boat for the lake around \glsfmttext{oathtower}.
    \item
    \textbf{Loud snoring} emanates from a sleeping hobgoblin.
    The room has a dozen hammocks made from cured snail-skin.
    \item
    Hobgoblin weapon storage, with $1D6 \times 2$ shortswords mounted on the walls.
    \label{towerWeaponStorage}
  \end{enumerate}
  \item[The first floor]\label{towerKitchen}
  has a massive kitchen with two storage rooms at opposite sides.
  $1D6$ elves debate politely, but their words hide fierce insinuations about how the others once burnt an egg, or used the wrong type of wood for cooking snail meat.
  \begin{enumerate}
    \item
    Light supplies, with $1D6 \times 3$ \glspl{torch}, and $1D6 \times 4$ candles.
    \item
    Meat storage, with $1D6$ meals' of \gls{crawler} eggs in salt, $1D6$ meals of snail-meat, $1D6$ meals' worth of vegetables (half of them gigantic and stolen by goblins from the garden plateaus), and $1D6$ meals of auroch meat.
  \end{enumerate}
  \item[The second floor]
  has an open room, full of cushions, with $1D6-3$ elves lounging.
  These elves have taken so many oaths that they are practically incapable of action, and need goblins to tend to them daily.

  The spiral staircase ends here.
  One door and two entrances sit behind the elves -- both open but gloomy.
  \begin{enumerate}
    \item
    Behind the door, an in-house outhouse which empties into the lake below.
    \item
    The first alcove houses a small study, with books on poetry, and fantastical erotica where rabbits ride foxes to battle snakes.

    The bookcase serves as a ladder to room \ref{MindElderRoom}, above.
    \item
    A cupboard of \glspl{ingredient}, including $1D6-3$ \gls{woodspy} beaks (Water \glspl{ingredient}), $1D6-3$ \gls{crawler} spinnerets (Fate \glspl{ingredient}), and $1D6-3$ phials of human blood (also Fate \glspl{ingredient}).

    The cupboard lifts back to reveal stairs up to room \ref{RomeoRoom}, above.
  \end{enumerate}
  \item[The third floor]
  has two rooms, accessible only from hidden entrances, below.
  \begin{enumerate}
    \item
    \Gls{MindElder}'s bedroom has a balcony, a bed with white sheets, and nothing else.\label{MindElderRoom}
    \item
    \Gls{romeo}'s bedroom has three books on poetry placed randomly around the room; each of them discuss the importance of a Mathematical underpinning.\label{RomeoRoom}
  \end{enumerate}
\end{description}

\enchantedHobgoblin[\npc{\T[3]\N}{Hobgoblins}]

\paragraph{Names:}
Glyph, Sillabery, and Majiscule.

\enchantedHobgoblin[\npc{\T[3]\N}{Hobgoblins}]

\subsubsection{An Interview with \glsfmtname{MindElder}}
will not be easy, because \Gls{MindElder} does not trust unknown travellers who carry weapons.
If the \glspl{pc} give him a good reason to speak, he speaks with one alone, in \gls{oathtower}, after they have left any weapons in with the hobgoblins on the ground floor, in room \vref{towerWeaponStorage}.

\begin{itemize}
  \item\it
  What's your name?
  \item[\adforn{54}]\bf
  `\Glsfmtname{MindElder}'.
  How can I help?
  \item\it
  Can you stop forcing people into oaths?
  \item[\adforn{54}]\bf
  Which oaths in particular?
  Do you want people to break their promises?
  Are you here to attack the peaceful people here in \gls{ravencops}?
  \item\it
  No, we won't attack anyone, but\ldots
  \item[\adforn{54}]\bf
  I thank you for your oath of peace.
  \item\it
  That wasn't an oath!
  \item[\adforn{54}]\bf
  In fact, it was.
  \item\it
  People should be able to sing, even if it's not in key!
  \item[\adforn{54}]\bf
  If you enjoy that sort of thing, do it elsewhere.
  I don't want to hear any more tuneless songs from the consumers.
  \item\it
  Can you stop making snails attack people?
  \item[\adforn{54}]\bf
  Aggressive snails eat \glspl{crawler}, and even attack \glspl{basilisk}.
  They make the land a lot safer on average.
  \item\it
  Well we're going to stop the snails being made in the first place!
  \item[\adforn{54}]\bf
  That saddens me, but I shall not stop you.
  I don't own the snails until they enter my land -- they come from the West, where lawless elves live.
  What will you do after you leave my tower?
  \item\it
  I suppose we'll go West, and then\ldots
  \item[\adforn{54}]\bf
  I accept your oath to travel West, and wish you well on the journey.
  Travel well!
  \item\it
  Thanks!
  I'm really excited to get going!
\end{itemize}

\MindElder

\showStdSpells[
  \input{config/spells/Water4.tex}
  \input{config/spells/Mind3.tex}
  \input{config/spells/Mind3.tex}
  \input{config/spells/Mind2.tex}
]

\thread[plateauGardens,ravencops,oathtower,sunway]{Oathless Lovers}
\label{oathlessLovers}

\noindent
\Gls{romeo} and \gls{juliet} have never met, but still managed to fall in love through a series of song-spells carved into trees in \gls{sunway}.
Each one spends months composing a new verse, then commits to the dangerous journey to carve more glyphs in their secret, shared spot.%
\footnote{Find that spot \vpageref{sunwayGlyphs}.}

Unfortunately, \gls{romeo} only understands love as an oath to be kept, while \gls{juliet} only understands oaths as an insult and a violation.
If the \glspl{pc} nudge them together, the two soon form a plan to meld the two elven lands into one, with \pgls{spell} so powerful it will destabilize the land, and bring all other threads to a dramatic and confusing conclusion.

\paragraph{If the couple unite,}
they discuss the grand spell, and request the \glspl{pc} help them gather the necessary \glspl{ingredient}.
They need sixteen in total:

%
\null
\begin{itemize}
  \item
  \Repeat{4}{\sgn}\quad 4 Earth \glspl{ingredient}
  \item
  \Repeat{4}{\sgn}\quad 4 Fire \glspl{ingredient}
  \item
  \Repeat{4}{\sgn}\quad 4 Water \glspl{ingredient}
  \item
  \Repeat{4}{\sgn}\quad 4 Fate \glspl{ingredient}
\end{itemize}

\Glspl{pc} can gather \glspl{ingredient} in various places,%
\footnote{There are \glsfmttext{griffin} wings \vpageref{sunwayAirIngredients} for an Air \glsfmttext{ingredient}, loads of \glsfmtplural{ingredient} in \glsfmttext{oathtower}, auroch hooves for Earth \glsfmtplural{ingredient} \vpageref{sunwayEarthIngredients}, \glsfmtplural{glowshroom} for Fire \glsfmtplural{ingredient} \vpageref{ravencopsFireIngredients}, and \pgls{woodspy} for a Water \glsfmttext{ingredient} \vpageref{plateauWaterIngredient}.}
but that's their problem, and the \gls{gm} shouldn't worry about helping them.
If this \gls{thread} does not complete successfully, there are other options.

\begin{exampletext}
  We want to unite the lands, making all one, a single people and place.
  This \gls{spell} will reframe everyone's problems into a memory.
\end{exampletext}

The couple won't be able to explain the \gls{spell}, except in abstract terms (`united, entirely', `the grand crossing of perspective').
See `\nameref{grandSpell}' \vpageref{grandSpell} for the complete description.

\segment[\gls{vlg}]{ravencops}% AREA
{Goblins in the Quarry}% NAME
{\Glsfmttext{romeo} should be working, but needs to complete the perfect poem}% SUMMARY
\label{goblinQuarry}

\Gls{MindElder} has sent his son \gls{romeo} to oversee the goblins, excavating rock at the quarry, and cutting long slabs to construct more sepulchres.
But \gls{romeo} can't think of anything but the poem he needs to write, to tell his beloved how he feels.
Unfortunately, his father raised him to be a perfectionist, which means he can't write perfect poetry, or good poetry, or bad poetry, or any poetry at all.

\begin{speechtext}
  What rhymes with snail?
  Mail, sail, bail\ldots hay-bail?
  Are hey-bails a thing?
  But `hey' is too informal.
  Better to say `hello'.
  `Hello-bail'\ldots no it sounds non-committal.
\end{speechtext}

So he stands looking at a blank sheet of paper, while twenty goblins ignore him, and bicker about pick-axes and the proper way to use a cart.

\begin{boxtext}
  In the near-distance, around this corner (or possibly two), someone, or something, is hitting metal on rocks.
  The metal sounds strange, butt probably iron.
  The rocks give that satisfying crack that rocks give with a long, clean cut.
\end{boxtext}

\paragraph{If the \glspl{pc} ask about the poem's recipient,}
\gls{romeo} explains he has no idea whom he loves, so he can only describe their mind.

\begin{speechtext}
  A quick wit, and very insightful in material matters -- able to tell the weight of a stone, bird, or an entire tree just by looking at it.
  And a deep critical thinker, not in any malicious sense, but nevertheless with cutting questions, whenever the need arises.
  This someone has wisdom beyond their years, though I don't know how old they might be, but still I'm sure of it\ldots

  I read their writing, and learned so much.
  They taught me how to move, and how things move.
  We write back and forth, we know each other so well.

  \ldots and yet, I cannot describe a face.
  But what's a face?
  Who cares?
  I just want to explain how I feel, and marriage to seal the deal.

  `Seal the deal'\\
  `An oath would make me less morose\ldots'

\end{speechtext}

\Gls{romeo} does not know whom it's for, and explains he learned from his teacher by reading, and fell in love utterly.
His father doesn't approve of the `oathless' types, and he feels ashamed of loving such a lawless person, despite all she's taught him.

\paragraph{If the \glspl{pc} help him with the poem,}
then he perks up and quickly finishes it, then asks them if they might try to find the recipient.

\paragraph{If the \glspl{pc} do nothing,}
\gls{romeo} remains at the quarry, thinking of the perfect words.

\paragraph{If the troupe commit crimes,}
the goblins will ignore them as long as they can.
They have taken an oath to dig rocks, and they will continue to dig until something shakes them from their oath.

\romeo

\showStdSpells

\paragraph{As the troupe leave,}
they notice \gls{romeo} using the Force \gls{sphere} to make the goblins' rocks lighter.%
\footnote{This tells the \glspl{pc} that \gls{romeo} understands the Force \gls{sphere}, which indicates a link to \gls{juliet}.}

\segment[\gls{vlg}]{sunway}% AREA
{Sunlit Glyphs}% NAME
{Elven glyphs, carved in wood, describe a song of ritual magic}% SUMMARY
\label{sunwayGlyphs}

\begin{exampletext}
  \Gls{juliet} began coming here to practice using the Force \gls{sphere} through ritual songs, and carved the notes into local trees to remember them.
  Years later, \gls{romeo} found the patch of glyphs and felt fascinated.
  So little by little, he taught himself the songs, and the rituals of the Force \gls{sphere}.

  And eventually, he began to carve his own glyphs, adding to her songs, or copying them with variation.
\end{exampletext}

\histEvent{45}{4}{%
  \Glsfmttext{juliet} begins carving song-rituals of the Force \gls{sphere} into a patch of trees in \glsfmttext{sunway}%
}

\histEvent{40}{3}{%
  \Glsfmttext{romeo} finds \glsfmttext{juliet}'s song-ritual glyphs, and slowly learns the Force \glsentrytext{sphere} from them%
}

\begin{boxtext}
  The dusky Sunlight makes little rune-shadows across the glyphs carved into all the surrounding trees.
\end{boxtext}

The elves who pass through here don't notice the glyphs because they normally ride giant snails.
The goblins who pass through here don't notice the glyphs, because they don't care.

\paragraph{Understanding the glyphs}
requires an \roll{Intelligence}{Academics} roll.

\begin{boxtable}

  \textbf{Roll} & \textbf{Result} \\\hline

   6 & These glyphs mean musical notes.  \\

   7 & \ldots and a few words in Elvish.  \\

   8 & Two people carved them.  \\

   9 & The `lyrics' pertain to Fire.  \\

  10 & \ldots and Earth.  \\

  11 & They form a magical ritual.  \\

  12 & Actually, they make many rituals.  \\

  13 & The rituals are of the Force \gls{sphere}.  \\

  14 & The ritual won't work without the missing words.  \\

\end{boxtable}


\segment{plateauGardens}% AREA
{\Glsfmttext{juliet}'s Flowers}% NAME
{Once dry, each works as a Force \glsfmttext{talisman}}% SUMMARY

A dozen elves spot clouds, describing what the clouds look like, and telling stories about the exploits of the `cloud-willow', and the `cloud-river'.
They think in terms of moving forests, and complete species, rather than individuals.

\Gls{juliet} takes no interest, as she is busy drying her flowers by pressing them onto a rock, then hanging them from a bush.
These are the `flowers of enlightenment', which she grows to make things float.
Check the details \vpageref{flowerOfEnlightenment}.

\juliet

\showStdSpells

%!
%\null
\begin{speechtext}
  You met the author?
  Can you deliver a message, saying to meet where we write?

  What eye-colour do they have?
\end{speechtext}

\paragraph{If the \glspl{pc} look tired,}
\gls{juliet} hands them a flower or two, and just says `\textit{flower of enlightenment, for the weight of the world}' (she doesn't speak the \gls{tradeTongue} well, but anyone who speaks Elvish can clarify).

\paragraph{If the \glspl{pc} identify \gls{juliet}}
(and potentially deliver a poem, written \vpageref{goblinQuarry}) she asks them to deliver a message back to \gls{romeo} (although she does not know his name).


\segment{oathtower}% AREA
{Wherefore Not?}% NAME
{\Glsfmttext{MindElder} tells \glsentrytext{romeo}, `there is no love without an oath'}% SUMMARY

\Gls{MindElder} has no respect for anyone's privacy, and has been \gls{casting} \textit{Witness Mind} to figure out why his son, \gls{romeo}, seems so detached lately.
Once he discovered that \gls{romeo} has become infatuated with an unknown elf from \gls{plateauGardens}, he felt enraged, and began to argue with \gls{romeo} about the character of the `unruly' elves who live over there.

\begin{boxtext}
  An angry voice echoes from \gls{oathtower}, then another.
  Fast, bitter words in Elvish ring out, then the tower goes silent.
\end{boxtext}

\Gls{romeo} eventually leaves through the tower's front door.
He feels angry, but has accepted his father's idea: ``love means nothing without an oath''.

The \glspl{pc} may try to change his mind, but it won't be easy.

\begin{speechtext}
  Why is she so committed to avoiding an oath of love?

  Why would you refuse an oath of love if you don't plan on violating the oath?

  If I can promise to love her, shouldn't she reciprocate?
  Do I have more duties than she?

  Why does she never wear clothes?
  What's wrong with those people?
\end{speechtext}

\paragraph{If the \glspl{pc} can provide rational answers,}
\gls{romeo} calms down and starts thinking clearly.

\paragraph{If the grand plan has already begun,}
(at the \gls{thread}'s start, \vpageref{oathlessLovers})
\Gls{romeo} reveals he has collected one \gls{ingredient} of each type required (Earth, Fire, Water, and Fate) from a storage room in \gls{oathtower}.
The plan requires twelve more \glspl{ingredient} in total.

If the plan has not begun, \gls{romeo} proposes it only if the \glspl{pc} calm him down.

\segment{sunway}% AREA
{Insider Knowledge}% NAME
{\Glsfmttext{romeo} gathers \glsfmtplural{ingredient}, while talking about the problems with goblins}% SUMMARY

\Gls{romeo} has found a spot of \glspl{marchingMushroom}, which he puts in his snail-gut satchel.
He has enough for two Earth \glspl{ingredient} so far, and hopes to find more.

\paragraph{If he trusts the troupe,}
he speaks openly to them about the problems in \gls{enchantedLands}.
Otherwise, he hides.

\begin{speechtext}
  My father has played a dangerous game with the consumers.
  He controls them only as long as they consume, but when they consume, they breed.
  He sends them to kill dangerous creatures, or bring news from far away, but most survive.
  Every \gls{sepulchre} with sleeping \glspl{ogre} marks another failure to control the population.%
  \footnote{\Gls{romeo} won't say so, but he holds the same view of local humans, and would destroy \gls{coppernut} and everyone inside if they looked threatening.}

  If the consumers suspected that \gls{MindElder} wants to reduce them, they would begin to push back against their oaths, or simply leave the area, past where he can reach them.

  We do not want that.
  We do not want to abandoned the consumers to their instincts.
\end{speechtext}

\segment{sunway}% AREA
{Change of Plan}% NAME
{\Glsfmttext{juliet} has figured out the spell will not work unless the troupe engineer an artificial \glsfmttext{flood}}% SUMMARY

\Gls{juliet} enters \gls{sunway}, studying her old notes, and realizes the plan for the great spell to unite everyone's perspectives across the two lands will not work; she needs \pgls{flood}.
The troupe will have to scout out \gls{shadepaths} to find where the water comes from,%
\footnote{Check the river locations \vpagerefrange{shadePool}{shadeDamn}.}
and engineer \pgls{flood}, perhaps with a damn.

How the \glspl{pc} engineer \pgls{flood} depends on them, but \gls{shadepaths}'s high walls, and secret streams allows them more opportunities than most locations.
Whatever their plan, it should be abstracted to \pgls{natural}; and if the roll fails, they will simply have to change their methods and tools to provide bigger bonuses.

\segment[\squash]{ravencops}% AREA
{The Goblin Hunting Party}% NAME
{A dozen goblins hunt for \glsfmttext{romeo}}% SUMMARY

\Gls{romeo} left his father's lands, to search among \gls{plateauGardens} for \gls{juliet}.
\Gls{MindElder} has sent out bands of goblins to find his location, and report back.

Majiscule leads six other goblins through the woods.
Soon they will reach \gls{plateauGardens}.
The goblins speak frankly with the troupe about their mission, and will not deviate from it, even for a moment, unless their lives are in danger.

\paragraph{If the grand plan has begun,}
then \gls{romeo} will have used this time to collect one more \gls{ingredient} of each type required.

\segment{sunway}% AREA
{The Grand Spell}% NAME
{The two unite the elves, goblins, \glsfmtplural{pc}, and snails as one}% SUMMARY
\label{grandSpell}

This climactic spell, planned by \gls{romeo} and \gls{juliet} will require 16~\glspl{ingredient} in total, plus \pgls{flood} and \pgls{earthquake}.
The \glspl{pc} will have to engineer the \gls{flood} inside \gls{shadepaths}, where rivers have been artificially diverted.
The \gls{earthquake} will occur naturally, as it always does, at the end of each \glsentrytext{cycle}.
This means the plan must wait until the end of \showCycle, so it can only occur at the end of the session.

Once the couple have all they need, they make each one into \pgls{boon}, and begin a grand ritual of song-magic, harmonizing together.

\spell{Shattered Identity in G Minor}% Name
  {Detailed, Distant, Divergent, Duplicated}% Enhancements
  {Warp}% Action
  {Water, Fate, Earth, Fire}% Spheres
  {\roll{Wits}{Survival}}% Resist with
  {The chorus alone is twenty minutes long, but once the spell really gets going, it binds everyone and everything at \spellRange\ into a self-jealous, psychic lump of non-space.}% Description
  {Everyone involved drops pieces of their memories, desires, and their perception of time into the fractured space, and each space involved loops together.

  Everyone involved gains a +\arabic{spellPlusOne}~Bonus to Empathy tasks, and a -\arabic{spellPlusOne}~Penalty to all other tasks.
  Both the Bonus and the Penalty reduce by one each time someone travels between one of the connected area, and finds another part of themselves.}

\begin{boxtext}
  Elves from \gls{plateauGardens} look down at you, scared and confused.
  Their usual blank-eyed cool has gone and you remember kissing \pgls{ogre} good night, and hoping to become like him one day.

  \Gls{oathtower}'s stairs lead up to \gls{plateauGardens}, or you can go into the elven home through the trees.
  Two consumers stand on the cooker holding your memory of learning the Air \gls{sphere}.

    ``\textit{Sorry, wrong person!}'', the little consumer says in Elvish.

\end{boxtext}

\begin{boxtext}
    Up the stairs, and out of the elvish house, \gls{oathtower}'s library is full of goblins, all of them are \gls{MindElder}, shouting for everyone to leave his home.
    Then they turn and point at you in unison, and begin to shriek, ``\textit{Identify! Identify}''.

    An eyeball reaches in, to return your name, if you want it\ldots ?
    The other eyeball searches the room upstairs.

    The name feels right, but the snail has another, and that one feels right too.
    The snail wonders how you know which name is your name, and wonders if you would share.
    The snail feels ashamed of its nakedness, so the eyeball slithers back out of the window.

\end{boxtext}

\begin{boxtext}

    The goblin hunger was deep, now infectious.
    You could enjoy the taste of your own arm, and it wouldn't actually be \emph{yours}, so it would be fine, and you would get a whole arm to yourself.

    \Gls{SnailTamer} arrives, telling you he loves all of you, and you're all doing really well.
    He's not really there, as he was taking a nap, and just decided to have a little dream with you guys to help everyone out.
\end{boxtext}

The spell ends before long, or it should if the players aren't into Dada-Taoism.
Or if they're getting the vibe, continue handing out memories until they've pieced their characters together.

After the experience ends, \gls{LifeElder} wanders past, stops to observe a spider-web, then continues.

\paragraph{Once the spell wears off}
the two groups of elves and the goblins will begin to work together, planning a route forward, each more fully aware of the others points of view.


\end{multicols}
