\sidequest[plateauGardens,ravencops,oathtower,sunway]{Oathless Lovers}
\label{oathlessLovers}

\noindent
\Gls{romeo} and \gls{juliet} have never met, but still managed to fall in love through a series of song-spells carved into trees in \gls{sunway}.
Each one spends months composing a new verse, then commits to the dangerous journey to carve more glyphs in their secret, shared spot.%
\footnote{Find that spot \vpageref{sunwayGlyphs}.}

Unfortunately, \gls{romeo} only understands love as an oath to be kept, while \gls{juliet} only understands oaths as an insult and a violation.
If the \glspl{pc} nudge them together, the two soon form a plan to meld the two elven lands into one, with \pgls{spell} so powerful it will destabilize the land, and bring all other threads to a dramatic and confusing conclusion.

\paragraph{If the couple unite,}
they discuss the grand spell, and request the \glspl{pc} help them gather the necessary \glspl{ingredient}.
They need sixteen in total:

%
\null
\begin{itemize}
  \item
  \Repeat{4}{\sqn}\quad 4 Earth \glspl{ingredient}
  \item
  \Repeat{4}{\sqn}\quad 4 Fire \glspl{ingredient}
  \item
  \Repeat{4}{\sqn}\quad 4 Water \glspl{ingredient}
  \item
  \Repeat{4}{\sqn}\quad 4 Fate \glspl{ingredient}
\end{itemize}

\begin{exampletext}
  We want to unite the lands, making all one, a single people and place.
  This \gls{spell} will reframe everyone's problems into a memory.
\end{exampletext}

The couple won't be able to explain the \gls{spell}, except in abstract terms (`united, entirely', `the grand crossing of perspective').
See `\nameref{grandSpell}' \vpageref{grandSpell} for the complete description.

\sqpart[\gls{vlg}]{oathtower}% AREA
{Groaning \Glsfmtplural{sepulchre}}% NAME
{The troupe must walk quietly and avoid the groaning sepulchres}% SUMMARY

\histEvent{20}{2}{%
  \Glsentrytext{MindElder} finds the goblins overpopulating the area, and fears the day they run out of food; even the most powerful enchantments cannot withstand goblin hunger.
  He rewards loyal goblins with obscene amounts of food, which lets them grow and grow, into hobgoblins, and eventually into \glsfmtplural{ogre}.
  Trials include service in \glsfmttext{oathtower}, hunting dangerous creatures, and plenty of duels (which really helps reduce the population).
  Once the goblins ascend, \glsentrytext{MindElder} places them in an enchanted sleep inside a stone sepulchre (where the goblins can check on them)}

\begin{exampletext}
  Goblins have no natural height limits, so when they eat too much, they just grow and grow, until one day, without any clear cut-off point, people call them a `hobgoblin', and soon after, `\gls{ogre}'.

  When the goblins become \glspl{ogre}, \gls{MindElder} puts them into an enchanted sleep, and tells the goblins their big brothers will awaken when the time of grand feasting comes.
  The goblins must see the sleeping \glspl{ogre} from time to time, or they will suspect murder and betrayal, and even their oaths will not keep them passive.

  So the forest around \gls{oathtower} has slowly filled up with snoring sepulchres, and everyone must tread quietly, lest they wake and ask for breakfast\ldots
\end{exampletext}

\begin{boxtext}
  Past the trees, an arrow's flight away, a mossy tower stands as tall as a feasting hall turns on its end.
  A low groaning noise, like a distant earthquake, floods through the trees, surrounding you.
\end{boxtext}

\Gls{oathtower} has a lot of sepulchres dotted around it, often hidden by trees, and always with little paths leading towards them.
Each one has three \glspl{ogre}, cramped in together.
The sound of snoring emanates for a few hours each day.

\iftoggle{verbose}{
  \gls{vlg}~Once the \glspl{pc} approach \gls{oathtower}, place \pgls{sepulchre} on the map on \vpageref{feylands}.
  It should be inside the forest, not far from the \glspl{pc}.
}{}

\paragraph{High-pitched noises near \gls{oathtower}}
have a 1 in 6 chance of waking \pgls{ogre}.
The chances increase by~1 for loud noises, or noises closer to the \glspl{sepulchre}.

In order to avoid waking the \glspl{ogre}, \gls{MindElder} has told the goblins to slay anyone making high-pitched noises, such as whistling or laughing.
Farting is also banned, as it makes the goblins giggle, which then wakes the \glspl{ogre}.

\enchantedOgre[\NPC{\M\N}{`The Grave'}%
  {slate-coloured skin, with bright-blue eyes}% DESCRIPTION
  {stretches calves}% MANNERISM
  {deer with cheese}% WANTS
  \npcQuote{only asking, only asking\ldots}]

\enchantedOgre[\NPC{\F\N}{Kerning}%
  {bra made from human faces (it helps with running, not modesty)}% DESCRIPTION
  {chews leaves, then spits them out}% MANNERISM
  {\gls{crawler} soup}% WANTS
  \npcQuote{the road goes ever on, unless it doesn't.
  `Dead end', they call it}]


\sqpart[\gls{vlg}]{ravencops}% AREA
{Goblins in the Quarry}% NAME
{\Glsfmttext{romeo} should be working, but needs to complete the perfect poem}% SUMMARY
\label{goblinQuarry}

\Gls{MindElder} has sent his son \gls{romeo} to oversee the goblins, excavating rock at the quarry, and cutting long slabs to construct more sepulchres.
But \gls{romeo} can't think of anything but the poem he needs to write, to tell his beloved how he feels.
Unfortunately, his father raised him to be a perfectionist, which means he can't write perfect poetry, or good poetry, or bad poetry, or any poetry at all.

\begin{speechtext}
  What rhymes with snail?
  Mail, sail, bail\ldots hay-bail?
  Are hey-bails a thing?
  But `hey' is too informal.
  Better to say `hello'.
  `Hello-bail'\ldots no it sounds non-committal.
\end{speechtext}

So he stands looking at a blank sheet of paper, while twenty goblins ignore him, and bicker about pick-axes and the proper way to use a cart.

\begin{boxtext}
  In the near-distance, around this corner (or possibly two), someone, or something, is hitting metal on rocks.
  The metal sounds strange, butt probably iron.
  The rocks give that satisfying crack that rocks give with a long, clean cut.
\end{boxtext}

\paragraph{If the \glspl{pc} ask about the poem's recipient,}
\gls{romeo} explains he has no idea whom he loves, so he can only describe their mind.

\begin{speechtext}
  A quick wit, and very insightful in material matters -- able to tell the weight of a stone, bird, or an entire tree just by looking at it.
  And a deep critical thinker, not in any malicious sense, but nevertheless with cutting questions, whenever the need arises.
  This someone has wisdom beyond their years, though I don't know how old they might be, but still I'm sure of it\ldots

  I read their writing, and learned so much.
  They taught me how to move, and how things move.
  We write back and forth, we know each other so well.

  \ldots and yet, I cannot describe a face.
  But what's a face?
  Who cares?
  I just want to explain how I feel, and marriage to seal the deal.

  `Seal the deal'\\
  `An oath would make me less morose\ldots'

\end{speechtext}

\Gls{romeo} does not know whom it's for, and explains he learned from his teacher by reading, and fell in love utterly.
His father doesn't approve of the `oathless' types, and he feels ashamed of loving such a lawless person, despite all she's taught him.

\paragraph{If the \glspl{pc} help him with the poem,}
then he perks up and quickly finishes it, then asks them if they might try to find the recipient.

\paragraph{If the \glspl{pc} do nothing,}
\gls{romeo} remains at the quarry, thinking of the perfect words.

\paragraph{If the troupe commit crimes,}
the goblins will ignore them as long as they can.
They have taken an oath to dig rocks, and they will continue to dig until something shakes them from their oath.

\romeo

\showStdSpells

\paragraph{As the troupe leave,}
they notice \gls{romeo} using the Force \gls{sphere} to make the goblins' rocks lighter.%
\footnote{This tells the \glspl{pc} that \gls{romeo} understands the Force \gls{sphere}, which indicates a link to \gls{juliet}.}

\sqpart[\gls{vlg}]{sunway}% AREA
{Sunlit Glyphs}% NAME
{Elven glyphs, carved in wood, describe a song of ritual magic}% SUMMARY
\label{sunwayGlyphs}

\begin{exampletext}
  \Gls{juliet} began coming here to practice using the Force \gls{sphere} through ritual songs, and carved the notes into local trees to remember them.
  Years later, \gls{romeo} found the patch of glyphs and felt fascinated.
  So little by little, he taught himself the songs, and the rituals of the Force \gls{sphere}.

  And eventually, he began to carve his own glyphs, adding to her songs, or copying them with variation.
\end{exampletext}

\histEvent{45}{4}{%
  \Glsfmttext{juliet} begins carving song-rituals of the Force \gls{sphere} into a patch of trees in \glsfmttext{sunway}%
}

\histEvent{40}{3}{%
  \Glsfmttext{romeo} finds \glsfmttext{juliet}'s song-ritual glyphs, and slowly learns the Force \glsentrytext{sphere} from them%
}

\begin{boxtext}
  The dusky Sunlight makes little rune-shadows across the glyphs carved into all the surrounding trees.
\end{boxtext}

The elves who pass through here don't notice the glyphs because they normally ride giant snails.
The goblins who pass through here don't notice the glyphs, because they don't care.

\paragraph{Understanding the glyphs}
requires an \roll{Intelligence}{Academics} roll.

\begin{boxtable}

  \textbf{Roll} & \textbf{Result} \\\hline

   6 & These glyphs mean musical notes.  \\

   7 & \ldots and a few words in Elvish.  \\

   8 & Two people carved them.  \\

   9 & The `lyrics' pertain to Fire.  \\

  10 & \ldots and Earth.  \\

  11 & They form a magical ritual.  \\

  12 & Actually, they make many rituals.  \\

  13 & The rituals are of the Force \gls{sphere}.  \\

  14 & The ritual won't work without the missing words.  \\

\end{boxtable}


\sqpart{plateauGardens}% AREA
{\Glsfmttext{juliet}'s Flowers}% NAME
{Once dry, each works as a Force \glsfmttext{talisman}}% SUMMARY

A dozen elves spot clouds, describing what the clouds look like, and telling stories about the exploits of the `cloud-willow', and the `cloud-river'.
They think in terms of moving forests, and complete species, rather than individuals.

\Gls{juliet} takes no interest, as she is busy drying her flowers by pressing them onto a rock, then hanging them from a bush.
These are the `flowers of enlightenment', which she grows to make things float.
Check the details \vpageref{flowerOfEnlightenment}.

\juliet

\paragraph{If the \glspl{pc} look tired,}
\gls{juliet} hands them a flower or two, and just says `\textit{flower of enlightenment, for the weight of the world}' (she doesn't speak the \gls{tradeTongue} well, but anyone who speaks Elvish can clarify).

\paragraph{If the \glspl{pc} identify \gls{juliet}}
(and potentially deliver a poem, written \vpageref{goblinQuarry}) she asks them to deliver a message back to \gls{romeo} (although she does not know his name).

\begin{speechtext}
  You met the author?
  Can you deliver a message back for me?

  Tell them to meet at our special place.
  They'll know what that means.

  What eye-colour do they have?
  Are they a man or woman?
\end{speechtext}

\paragraph{If the troupe mention that \gls{romeo} lives in \gls{ravencops},}
she curls her nose in disgust, and seems sad, but perseveres in trying to meet with him.%
\footnote{The two elven groups have incompatible values.}

\sqpart{oathtower}% AREA
{Wherefore Not?}% NAME
{\Glsfmttext{MindElder} tells \glsentrytext{romeo}, `there is no love without an oath'}% SUMMARY

\Gls{MindElder} has no respect for anyone's privacy, and has been \gls{casting} \textit{Witness Mind} to figure out why his son, \gls{romeo}, seems so detached lately.
Once he discovered that \gls{romeo} has become infatuated with an unknown elf from \gls{plateauGardens}, he felt enraged, and began to argue with \gls{romeo} about the character of the `unruly' elves who live over there.

\begin{boxtext}
  An angry voice echoes from \gls{oathtower}, then another.
  Fast, bitter words in Elvish ring out, then the tower goes silent.
\end{boxtext}

\Gls{romeo} eventually leaves through the tower's front door.
He feels angry, but has accepted his father's idea: ``love means nothing without an oath''.

The \glspl{pc} may try to change his mind, but it won't be easy.

\begin{speechtext}
  Why is she so committed to avoiding an oath of love?

  Why would you refuse an oath of love if you don't plan on violating the oath?

  If I can promise to love her, shouldn't she reciprocate?
  Do I have more duties than she?

  Why does she never wear clothes?
  What's wrong with those people?
\end{speechtext}

\paragraph{If the \glspl{pc} can provide rational answers,}
\gls{romeo} calms down and starts thinking clearly.

\paragraph{If the grand plan has already begun,}
\Gls{romeo} reveals he has collected one \gls{ingredient} of each type required (Earth, Fire, Water, and Fate) from a storage room in \gls{oathtower}.
The plan requires twelve more \glspl{ingredient} in total.

\sqpart{sunway}% AREA
{Insider Knowledge}% NAME
{\Glsfmttext{romeo} gathers \glsfmtplural{ingredient}, while talking about the problems with goblins}% SUMMARY

\Gls{romeo} has found a spot of \glspl{marchingMushroom}, which he puts in his snail-gut satchel.
He has enough for two Earth \glspl{ingredient} so far, and hopes to find more.

\paragraph{If he trusts the troupe,}
he speaks openly to them about the problems in the \gls{enchantedLands}.
Otherwise, he hides.

\begin{speechtext}
  My father has played a dangerous game with all these goblins.
  He controls them only as long as they remain fed, but if they feed, they will breed.
  He sends them to kill dangerous creatures, or bring news from far away, but most survive.
  Every \gls{sepulchre} with sleeping \glspl{ogre} marks another failure to control the population.%
  \footnote{\Gls{romeo} won't say so, but he holds the same view of local humans, and would destroy \gls{coppernut} and everyone inside if they looked threatening.}

  If the goblins suspected that \gls{MindElder} wants to reduce them, they would begin to push back against their oaths, or simply leave the area, past where he can reach them.

  We do not want that.
  We do not want to see the goblins abandoned, in great number, to their own hunger.
\end{speechtext}

\sqpart{sunway}% AREA
{Change of Plan}% NAME
{\Glsfmttext{juliet} has figured out the spell will not work unless the troupe engineer an artificial \gls{flood}}% SUMMARY

\Gls{juliet} enters \gls{sunway}, studying her old notes, and realizes the plan for the great spell to unite everyone's perspectives across the two lands will not work; she needs \pgls{flood}.
The troupe will have to scout out \gls{shadepaths} to find where the water comes from,%
\footnote{Check the river locations \vpagerefrange{shadePool}{shadeDamn}.}
and engineer \pgls{flood}, perhaps with a damn.

How the \glspl{pc} engineer \pgls{flood} depends on them, but \gls{shadepaths}'s high walls, and secret streams allows them more opportunities than most locations.
Whatever their plan, it should be abstracted to \pgls{natural}; and if the roll fails, they will simply have to change their methods and tools to provide bigger bonuses.

\sqpart[\squash]{ravencops}% AREA
{The Goblin Hunting Party}% NAME
{A dozen goblins hunt for \glsfmttext{romeo}}% SUMMARY

\Gls{romeo} left his father's lands, to search among the \gls{plateauGardens} for \gls{juliet}.
\Gls{MindElder} has sent out bands of goblins to find his location, and report back.

Majiscule leads six other goblins through the woods.
Soon they will reach \gls{plateauGardens}.
The goblins speak frankly with the troupe about their mission, and will not deviate from it, even for a moment, unless their lives are in danger.

\paragraph{If the grand plan has begun,}
then \gls{romeo} will have used this time to collect one more \gls{ingredient} of each type required.

