\sidequest[Broch,Canals,plateauGardens,Causeway,Oathtower]{The Snail Tamer}

\sqpart{Broch}% AREA
{A Logistical Discussion}% NAME
{A monster approaches a distant bailey, prompting arguments over duty vs stupidity}% SUMMARY

Over the sea of dense green, something strange and monstrous moves, taller than the canopy, approaching the bailey below.

According to the \gls{jotter}:

\begin{speechtext}
  You are the night guard.  That is a beast.  Go guard!
\end{speechtext}

But according to the archer:

\begin{speechtext}
  The bailey lies two miles away, which requires half an hour at a good march.
  The beast looks to be two miles beyond the bailey, so it will arrive long before us, and when we arrive, the situation will already have ended\ldots one way or another.
\end{speechtext}

And according to the ranger:

\begin{speechtext}
  You can run two miles in ten minutes, with gear.
  Beasts always take it slow, so we can beat it; but best not.
  We don't know what it is.
\end{speechtext}

In truth, \gls{SnailTamer} has lost control of his mount, `Nettlerash', who recently acquired a taste for living flesh in the Kingdom of Oaths.

If the troupe watch from the safety of the broch,
they will intermittently see the eyestalks rise again above the canopy, and soon notice how slow the thing is (but will not be able to identify it from the two bulbous tentacles) poking into view for those small moments.

If the troupe warn the bailey,
the farmers let them in, then cover the wall in archers -- 20 in total.

Once the troupe see the snail,
\gls{SnailTamer} call down to them, with his standard sloth and naïveté.

\begin{boxtext}
  The wind dies down, leaving a distant crashing noise which becomes louder, and sounds like crushed bushes and snapped branches, mixed with  retching.
  A rock-like surface emerges, the size of a cottage, with a smooth, brown neck hovering above, which tapers to a point above with two tentacles pointed straight in the air.
  An elf sits on the rock-like surface above, waving slowly.
\end{boxtext}

If the snail sees the troupe (or anyone) it begins projectile-vomiting acid while \gls{SnailTamer} -- the elf on the snail's enormous shell -- begins to explain himself:

\begin{description}
  \item[round 1] So, hey, um\ldots everyone there.
  Can you hear me?

  \item[round 2] That's good, good that you can hear me, because I really want to ask you if you might not hurt Nettlerash here.
  She's a good girl usually.

  \item[round 3] Actually, I named Nettles after a human, or more like, humans in general, because, as you can see, she is very tall, like a human.
  Are you sure you need to do that with the sword?
  I think it's upsetting her.

  \item[round 4] So, it all started a while ago, we went a little too close to the tower, and\ldots wait, let me back up a bit.
  It's actually important that you understand some of the earlier events (I suppose you might call these events `history', or is that prejudiced?
  Sorry, I never actually met humans, unless you count dwarves\ldots).

  \item[round 5] Do you count dwarves?
  I mean no judgement, I just wondered, because\ldots wait, Nettlerash really is not looking good.
  Okay, this is serious, we really need to discuss the use of swords and how to respect differences, and\ldots
\end{description}

\paragraph{If the \glspl{pc} attempt a peaceful resolution,}
they can manipulate the snail easily.

\paragraph{If Nettlerash dies,}
\gls{SnailTamer} leaves, saddened, and will not speak with the troupe until he collects his thoughts.

\paragraph{Once the combat dies down,}
the \gls{jotter} appears, and asks the troupe what happened to the ranger.
Someone needs to find the source of that giant snail, and make sure no more will be coming this way.

\paragraph{If the troupe ask \gls{SnailTamer} about his plans,}
he says he has a letter for \gls{LifeElder}, from \gls{MindElder}, and must deliver it at once.

\begin{speechtext}
  I have to go. \\
  I have this letter for someone.

  Who is it? \\
  She's just a person, like anyone, but she's not into labels.

  I don't call her anything. \\
  Well, maybe I call her `hi'.
\end{speechtext}

\sqpart{Canals}% AREA
{Debating Parsnips}% NAME
{\Glsentrytext{SnailTamer} meanders home, wondering if he should make parsnip or potato soup}% SUMMARY

\Gls{SnailTamer} has missed a turn-off in the maze of canyons, and can't remember which direction he's going.
However, he's more concerned with the question of what kind of soup to make.

\begin{boxtext}
A voice in the distance sounds like it's practising Elvish vocabulary, and you recognize the word for `parsnip'.
Around the corner, \gls{SnailTamer} appears, as naked as the Sun, then stops and stares, as if trying to recognize you.
\end{boxtext}

\begin{speechtext}
  So, like.  Guys?  Yea, you guys.  Okay then.  So we don't make people do things here, no laws about fabrics over your body -- you just walk about however you like, okay?

  Potatoes are always great, but you can't eat potatoes all day.
  It's been ages since I had parsnip soup actually.
  But the parsnips are fresher\ldots I don't want the potatoes to go bad.
  That would be a waste\ldots
\end{speechtext}

\Gls{SnailTamer} continues meandering and mumbling until someone snaps him out of it.
If \pgls{pc} makes a clear request help getting out of the canyon and onto the plateau, he agrees, says `\textit{let's go then}', and starts walking while whistling a sad song.
Soon enough, an elf in one of the raised gardens will hear him, and lower a few ripe bean-vines.

The bean-vines can take a total \gls{weight} of 10.
The \glspl{pc} can gauge the weight limit with an \roll{Intelligence}{Survival} roll at \tn[8].

\sqpart{plateauGardens}% AREA
{Shell Escape}% NAME
{Giant potato sprouts vomit from \gls{SnailTamer}'s house, it may fall any moment!}% SUMMARY

If you leave potatoes in a bag with a little Sunlight, they will grow arms and try to find earth to implant themselves into.
Giant, magical, elf-potatoes are entirely normal in this regard, but much bigger.

\begin{speechtext}
  Across the next bean-vine bridge, a giant snail-shell stands, dangerously close to falling into the canyon.
  Sunlight shines through the shell, displaying an upper floor with a desk, hammock, some cloaks, and jars.

  Pale tendrils push out the door, then have a hard twist down into the earth, like a limb trying to grasp the ground.
\end{speechtext}

\paragraph{Get an item from the house:}
\roll{Dexterity}{Stealth} at \tn[12] (+1 per item).

Righting the house, so it will not fall over:
\roll{Strength}{Cultivation}, \tn[12].

\paragraph{Using the vines to secure the house:}
\roll{Intelligence}{Cultivation}, \tn[10].


\sqpart{plateauGardens}% AREA
{Choosing a Shell}% NAME
{\Glsentrytext{SnailTamer} to pick a new snail}% SUMMARY

\Gls{SnailTamer} needs to pick a new snail so he can get about safely, but he can't decide which snail is best.
One seems to unbalanced for him to put the wooden carrot-rod onto, another looks `too old', then the next is pregnant (which will cause problems before long).

If \gls{SnailTamer} still suffers from the \gls{disgnome} stings,
the \glspl{pc} can usher him along with a \roll{Charisma}{Cultivation} roll at \tn[10].
Otherwise, he'll pick a snail within the \gls{interval}.

\sqpart{Causeway}% AREA
{Urgent Delivery}% NAME
{\gls{SnailTamer} races to deliver latter on back of snail}% SUMMARY


\Gls{LifeElder} was bound to send `her most reliable messenger', and that's \gls{SnailTamer}.
He's not the fastest, or well-spoken, nor is he routinely sure of where he is.
However, he has delivered two messages, which means his success rate is $2/2$, and 100\%.

So she charged him with delivering the letter to \gls{MindElder}, and he's finally en route.

\sqpart{Oathtower}% AREA
{The Letter of the Law}% NAME
{\gls{SnailTamer} wants someone else to deliver the letter, so he can avoid any binding enchantments}% SUMMARY

At the tower, \gls{SnailTamer} has changed his mind about entering, and decided to hide behind his new mount.
This plan would have worked well, except for the large wooden pole mounted on the snail's back, indicating that it has a rider.

If the \glspl{pc} agree to deliver the letter,
\gls{SnailTamer} hands his satchel, which feels dank and smells musty, as it's full of moss.
Once opened, the moss is in the shape of the letter `F' (which is the letter \gls{LifeElder} wanted to send).

