These random side quests should work with any area which contains chitincrawlers, such as Mt Arthur, Dogland, or Quennome.  See page \pageref{sidequests} for how to run them.  If your area doesn't contain chitincrawlers, just remove those encounters which features them.

If you have a plot-arc in mind already, you can pull out a couple of these random encounters as flavouring for your campaign, or just slot them in when you've not prepared enough material for the night.  Alternatively, you can lump the lot together, and run them as a major plot-arc.  The overall plot contains two major arcs:

\section{Expanding Wilderness}
A long time ago, humans build too far into the forest, distubring elven settlements with their noise, and polluting a mana lake by building a city around it.  The elves gave them repeated polite letters, demanding they leave the area, but the city and surrounding hamlets kept on growing.  Eventually, the elves decided to destroy the city.

The coordinated attack happened on a single night.  One group felled trees and used magic to grow plants over the largest trade routes.  Another group assassinated military leaders, or poisoned entire guard houses.  Finally, they gathered to summon all the most dangerous animals of the forest to attack the town.

Centuries on, the city and nearby villages lie abandoned, and overgrown.  However, \gls{baron} has decided he wants to move up in the world, and he can only do that by making new territory.  He's already found the location of the lost city.  It has fertile ground, lots of stone (from abandoned walls and buildings), and he wants to start by building a military outpost.  He's banking on the idea that the new king will not stand for elven interference this time, and the elves will back away from a war.


\gls{forestpriest} has learned of \gls{baron}'s plans.  He's a friend to elves of the region, and wants to stop the coming war.  Being a long-term thinker, he's quite willing to kill a few people for the greater good, and in particualr, he's been using his magic to turn anyone he finds involved in the conspiracy into animals.

\section{The Nura Conspiracy}
The Whiteland nobles, after losing their homes, land and power, have gathered in the town's sewer's.  They found a portal there to the nura realm, which they could open at will.  \gls{sewerking} immediately started thinking of how he could use this to spark a rebellion:

\begin{speechtext}
	We have connections all over, but you have to think local, even when trying to overthrow a king.  The town's small, but it's a major hub for trade, and information, especially since we got \gls{pigowner} on side.  She's a fierce one.

	Setting up in a sewer sounds horrific, but we have a nice area at the side we've decked out like a normal house.  It even has fancy cushions.  Of course we explored deeper and deeper.  I think the original home must have been owned by some mad wizard.  Gods only know what has happened to him -- he probably wandered into that portal.

	Right, the portal!  You just say the word (an no, I won't tell you what it is), and the great round stone circle opens up to another world.  And that world stinks.  It stinks in there, really bad.  There's a full castle, and you can look down and see a full world of these horrible goblin creatures wandering about.

	We gathered men from above.  We'd gone on some raids, gathered some money (everyone needs to survive, okay?) and we'd started recruiting.  Truth be told we were drunk when we first opened the portal with others, and at the other side, a man stands there, asking if he can come in.  Just this bald guy, with a funnyfucker accent.  He asks us if he could have a bit of the pork we were eating, so I tossed it towards him (because who wants to actually invite someone in from a magical portal?) and he eats it slowly, then says ``thank you'', like we're all sitting down to tea.

	Well that was it rolling.  He tells us he's the king there, and that he will gift us magical items for our generosity.  We're not stupid, so obviously, we tried the ring out on a street kid from above.  He read this scroll aloud and \emph{bang}, fire rises in front of him, and throws itself onto the nearest wall.  The burn-marks are still there.

	Since then it's been rings that can keep you awake and tireless for days, daggers that summon fire, and this ogre-dust that turns someone into a warmachine for a moment (though usually when we use it, we just end up eating a lot).

	He's a clever one alright, \gls{patron}, and we don't altogether trust him.  And I know what you're thinking -- we're being duped, and something bad's going to happen.  But what you need to understand is that the portal closes after a few minutes, and we're the only ones with the word to open it, and it has to be opened from this side.

\end{speechtext}

Since that time, \gls{banditking} has decided to live outside with a few other men, coordinating things from outside.

\begin{speechtext}
	At the start it was all innocent -- I'd dress up as a travelling alchemist and sell a magical item for a small price to a few of the local lords.  Then I started selling those scrolls, where you read them and they only work once.  I came back one day just for dinner, and spoke with a few of them.

	They all had the same complaints -- the taxes are too high, \gls{king}'s gone mad, and worst of all were the constant envoys checking on this and that, making sure everything was accounted for, and they never had any power.  They couldn't even keep their own guards, just had to rely on the Night Guard to look after their area, and the Night Guard aren't always were people need them.

	So we got talking more, and said we had more items, and overall the plan was to wait and gather others, with an understanding that when the time was right, we'd all kick off at the same time.  We could free all of Fenestra.

	It's been getting better faster than we thought.  \Gls{patron} got us some hobgoblins one day, with manacled feet so we didn't have to worry about them turning on us.  They worked and worked for weeks, and we got a full dungeon now.  We've been filling them with the dead.  Anyone in town who dies eventually gets brought down here, locked in a vault, and \gls{patron} raises them for us.  When the time comes, we can ruin the town.

	I know it sounds a bit bloody, but it's nothing worse than what \gls{king} did to us when he took our land.  He made those rules, and we're just playing the game.  So when push comes to shove it's going to be one wasted town, trade routes destroyed, and we're the only ones in the habit of hoarding food.  We'll wait for his armies to tire, all of them thinking the nura and the dead have risen from the ground, and when they're panting and bleeding, we'll arrive with heavy chain and swords.

Till then, we wait, we build our armies, and we speak gently with the other Villagemasters.

	Recently, we've obtained a new base of operations in the forest.  It's been the strangest time, I can tell you, really has.  Some of the rings -- these magical rings -- which came through, well they make you feel kind of \emph{cold} and stop you feeling tired.  Then one day in the forest we find this abandoned temple to Qualme, so we investigate and find it's full of the walking dead, and they just\ldots don't do anything.  It's like we're invisible.  Soon enough we found it's the rings.

	So we stayed there for a bit\ldots okay we didn't -- we fucking ran away, okay?  The smell of death, the wandering dead people, we couldn't take it.  But it's amazing what you get used to in time, you know?

	Well we explored around, and found the mausoleaum's abandoned.  How weird is that?  The living are in the mausoleum, while the temple's occupied by the dead.  So we're camped out there, with just one neighbour to speak of.

	It's been a funny arrangement, but then we're living in funny circumstances.  We approached the central tower, the actual temple itself, and there was this awful screaming, and it sounded like someone saying ``Get out now!''.  So we got out.  And some time later, one of the dead approached us with a piece of paper and a quill and ink.  And it asked us who we were.

	And you know what?  I explained the lot.  A leap of faith with a monster, and I told it the full plan, full and honest.

	And he let us stay.

	When our horses died, he pulled them back to life, and they do just what they're told, though they have a strange occupation with dead bodies.  And all he's really asking in return is that once in a while, we bring someone back, still breathing, and we leave them in the tower for him.  And that's not too much to ask, considering his hospitality.

	And I know what you're thinking -- but the dead can't see us, and we're ready to vanish from there at a moment's notice if things go sour.  I got this under control, and when it all starts to go wrong, we'll be safely here while the fighting starts with the king against the dead.  And once everything's ash, we'll rise again, and we'll be ready to liberate the land, and split Fenestra up into six Kingdoms, just like the old days, and everyone can just rule over their own part of \gls{shatteredcastle}, and we'll finally be free.

\end{speechtext}




