% !TeX root = ../main.tex
\index{Nura}\label{nura}
The nura are twisted versions of natural creatures -- humans and dwarves, spiders and cats; anything can be twisted with dark magic into a nura creature.

The nura always want food, but never feel satisfied.  Whether spiders, ogres or nura horses, they will destroy a population's food resources quickly before burning out and dying of hunger, or returning underground.

\section{Nura Encounters}\index{Nura!Encounters}

The nura are ever present, but with differing degrees.  Sometimes they are running amok across an area, while at other times they have been mostly killed and the land is quiet.  The number of Nura in the area can be given a rating - anywhere from 1 upwards, where `1' represents almost no nura creatures in the environment and `3' represents a few but not many, and so on.  Whenever you roll on the encounter tables, if roll equal or lower than the Nura Rating, the nura encounter occurs.

	For example, if the nura rating is `5' and the PCs are wandering in a forest, rolling 3-5 will mean a nura encounter, rolling 6-7 will mean no encounter and rolling 8 or more means a normal forest encounter.

	You can set the nura rating to fit your current story, but for a standard `background' level, set the nura level to 3, and then raise it by 1 every time the PCs encounter the nura.  This will allow the campaign to gently increase in danger bit by bit.

The players can lower the general nura rating by `plugging holes'.  Nura come from below the ground or through magical portals.  Once an entrance to Fenestra has been sealed off, the nura represent less of a threat.

\begin{encounters}{with the Nura}
	\setcounter{enc}{18}
	& & \textbf{Nura Encounters} \\\hline
	\li & Goblins riding nura spiders. \\
	\li & Lava Man (page \pageref{lavaman}). \\
	\li & Ghast (page \pageref{ghast}). \\
	\li & Nura Horse (page \pageref{nura_horse}) \\
	\li & $(2D6)^{2}$ ghouls (page \pageref{ghoul}).\footnote{This encounter indicates that a nearby village has been slain, and spare corpses raised as ghouls.} \\
	\li & Goblins riding ogres. \\
	\li & Hobgoblins riding nura horses. \\
	\li & Goblins riding nura wolves. \\
	\li & Nura Cat (page \pageref{nura_cat}) \\
	\li & Nura Spider (page \pageref{nura_spider}) \\
	\li & Nura Wolves (page \pageref{nura_spider}) \\
	\li & Ogre (page \pageref{hobgoblin}) \\
	\li & Hobgoblin (page \pageref{hobgoblin}) \\
	\li & Goblins (page \pageref{goblin}) \\
	\li & Nura Slugs (page \pageref{nura_slug}) \\
\end{encounters}

As nura emerge from magical portals, they find more and more opportunity to take creatures back with them underground, where those creatures too turn into nura.  Once people become less common, nura start taking any animals they can back underground with them.  Eventually, nuramancers emerge from underground and begin to raise the dead.  Entire villages are sometimes killed and pulled back from death, just to roam the landscape and consume souls.

If the nura rating is high enough that it coincides with a regular encounter, drop the encounter and just put nura there -- they eat everything in the landscape, so it makes sense that regular creatures would be seen less often.

\section{Nura Magic}\index{Nura!Magic}\index{Saurecanta}\label{saurecanta}

\subsection{New Path: The Path of Nura}

\textit{Spheres: Conjuration, Invocation, Metamagic, Necromancy, Saurecanta}

\noindent Occasionally, the strange creatures of the deeps emerge with apparently inborn magical abilities fuelled by the corruption in their bodies.  The Nura humanoids such as goblins and ogres occasionally learn such magics, though it can be difficult as they are never very intelligent, and while the Path of Nura is a strange Path of magic, it is still based upon one's Intelligence.  It is also possible to learn such magics through memorization of corrupt thoughts alone - books uttering extreme and surreal crimes are known to exist which can teach anyone how to step onto the Path of Nura.

\paragraph{Signs:} When cast, inky black mist, speckled with violent red appears around wherever the spell brings something into existence.  The mana stones of the Path of Nura are always instruments of change.

\paragraph{Mana Stones:} Unorthodox books, cutting knives, revolutionary art - anything which can promote drastic change.  Such mana stones are activated by a condition if they have a spell imbued in them, and their spells are cancelled by a condition.

\sphere{Saurecanta}

This new sphere of magic comes from the foul realm under the earth where strange creatures breed and eat at a dizzying pace.  It bears a passing similarity to Polymorph but with fewer restrictions on form and without any ability to disguise oneself as a natural creature.  While a creature is affected by this sphere, they must eat a minimum of thrice the normal amount; this need not mean constant intake of food - a single massive meal will suffice.  Failure to eat inflicts the usual Fatigue Points.

	Each level of Saurecanta is a double-edged sword, allowing targets extra abilities at a cost.  While those affected can gain a lot of power, they are also afflicted with unending hunger.  Any scene in which they do not eat, the characters heal no Fatigue Points, and gain 1.  Meanwhile, the character can heal a number of Fatigue Points each scene equal to their maximum HP, simply by gorging on food.

	Saurecanta spell effects never stack with each other, or with Polymorph -- only the highest bonus counts.

\spelllevel \label{saurecantaone}
\spell{Hunger Pains}{Continuous}{Beast Ken}
The target is affected by a ravenous hunger and extreme stomach pains.  They suffer 1 HP Damage per 2 Fatigue points they currently have - FP cannot be spent to mitigate this.

\spelllevel
\spell{Brawnform}{Continuous}{Medicine}
The caster pulls out the inner beast in anyone, polluting their soul and improving their body.

The target gains \arabic{spelllevel} points + the caster's Intelligence Bonus to add to Strength.  Each point so gained reduces the target's Charisma, and with it the target's FP.

The targets gain only 1 Attribute point per round as muscle-mass slowly grows, but lose their Charisma instantly.  Targets who reach 0 Charisma begin to look obviously unnatural.

Characters who gain more than a single point of Strength break out of their armour, taking 1 point of Damage for each level of DR the armour provided.

\spell{Speed Demon}{Continuous}{Medicine}
As above, except the character gains a +\arabic{spelllevel} bonus to speek + the caster's Intelligence Bonus, while the target's Intelligence is reduced by the same amount.

As above, targets gain the Intelligence penalty instantly, but only gain their bonus to Speed at a rate of 1 point per round.  Targets with less than -5 Intelligence lose all control, and act as animals.

\spelllevel
\spell{Ultraform}{Continuous}{Beast Ken}
This spell functions as per {\it Speed Demon} and {\it Brawnform} above, combined.  The target gains a number of points equal to \arabic{spelllevel} + the caster's Intelligence Bonus, which the caster can assign to either of Strength or Speed.  Additionally, any negative Attribute can be raised to 0 by spending a single point of the total bonus.  Each effect takes a single round.

The target gains an equal penalty to {\it both} Intelligence and Charisma.

Animals can also be affected by this spell, including animals which would normally be considered too small to have a Strength or Speed Attribute can be turned into little monsters.

\spelllevel
\spell{Muscle Explosion}{Continuous}{Medicine}
This functions as per {\it Ultraform}, above, except that the bonuses occur instantly, and the total points gained and lost is equal to \arabic{spelllevel} plus the caster's Intelligence Bonus.

\spell{Herd Mind}{Continuous}{Deceit}
After long enough in the nura form, all nura notice a kind of kinship which, at the very least, stops them eating each other.  This spell pushes that effect forward, making it instant.  It also discourages the target from engaging in strictly non-nura aims, such as quests.

Any time the target wants to attack the nuramancer, or any of the nuramancer's companions, they must make a Wits + Vigilance roll, with a TN equal to 11 plus the caster's Deceit bonus.

\spelllevel
\spell{Demonic Form}{Continuous}{Ether Lore}

This functions as above, plus the ability to grant innate abilities.  The caster can implant a spell inside any nura creature at the cost of a number of ability points equal to the spell's level.  The creature gains a number of natural Mana Points equal to the spell's level, which can only be spent on these in-built nura spells.

\textit{For example, one nuramancer has Intelligence +2, Saurecanta 5 and Invocation 3.  Turns a nearby dog into a rabid hellhound which shoots fireballs.  The dog's Strength is -2 and its speed is +3.  The nuramancer has 7 points to spend in total.  The first is used to bring the dog's Strength up to 0.  Three more points are used to grant the dog a third-level Invocation spell -- it can now bark fire at its opponents, dealing $2D6+2$ Damage (as per the nuramancer's would-be Damage for the spell).  The last three points are used to bring the dog's Strength up to +3.}

\textit{The dog now has +3 Strength, +2 Speed, and has 3MP to cast spells.}

\noindent Creatures can regenerate up to three of their internal MP only by gorging upon food.  A complete meal is equal to 1MP.

