
\section{Random Encounters}
% !TeX root = ../main.tex
\index{Encounters}

Whether you're in the middle of an adventure or the PCs are just randomly wandering the world without any respect for local laws or plot, a random encounter can always add some excitement.  Each time the players pass through a new type of terrain, roll $3D6$ on the encounter table and create an encounter from the result.  Each area has a specific encounter table, but they look a little like this:


\begin{encounters}{Random Places}

Random Woods & Random Roads & Result \\\hline
	\li & Mana Lake. \\
	\li & Bassilisk. \\
	\li & Chitincrawler. \\
	\li & Woodspy. \\
	\li \lii Griffins. \\
	\li \lii Bear. \\
	\li \lii Bandits. \\
	\li \lii Wolves. \\
	& \lii Human traders. \\

\end{encounters}

If you're on the roads, your party can walk with relative safety.  The most likely encounter is human traders -- a harmless bunch, and then wolves (an equally harmless bunch).  Meanwhile, the woods contain deadlier creatures, and no chance of finding human traders.  Despite the different tables, the overlap provides some cohesion to the area.

If you reach a result which is not listed, there is no encounter.  If you roll trips -- three of the same number -- roll again, and if you get another encounter, combine the two.  If you get a griffin and a bandit, perhaps the players stumble upon bandits in the woods, attempting to pilfer griffin eggs for a patron.  If you roll wolves and a chitincrawler, perhaps the players hear persisten wolf-cries in the distance as a chitincrawler has caught some wolves in its web while the others watch and bark helplessly.

Each creature listed has some suggested encounters, and each area has its own particular encounter tables.\footnote{See page \pageref{encounters}.}

	You may want to set up your random encounter before the start of the session, allowing you to review monsters' stats and perhaps tie the encounters together, or integrate them with active characters from players Story Points, or recent events in the campaign.

	Encounters do not have to be deadly, or even confrontations.  Wandering wolves might attempt to steal the PCs' rations, a bear might stalk them for a while, or the players could find a human corpse hanging from a chitincrawler's web and simply have to make a decision to give a proper burriel or leave the corpse where it is.


%\subsection{Towns}
%
%\begin{longtable}{|rp{.7\textwidth}}
%\setcounter{enc}{18}
%
%	Roll & Result \\
%	\hline
%	\arabic{enc} & {Someone poisons the PCs' drinks in an attempt to steal their things.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {A thief attempts a confidence trick on the characters.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {A monster is running amok through town.  Roll 2D6+6 on the Forest encounter table.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {Monsters are running amok around the outskirts of the city.  Roll thrice on 	the forest encounter table.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {The characters are asked to protect a convoy travelling to another town or city.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {The guards ask the characters to hunt down a known murderer in the city.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {The guards ask the characters to hunt down a known thief in the city.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {Roll 3D6 twice and add encounters.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {1D3 Pickpocketing children target the character.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {Local guards demand a bribe after creating a faked charge.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {2D6 thugs surround the group.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {A guild tries to hire the characters for an in-city job such as murder, theft or just casual labour.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {A random person a member of the group has met before appears.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {A guild tries to hire the characters for an out-of-city job - retreiving an item from a noble's house, murdering someone while travelling or infiltrating a secret meeting are all possible.} \\
%	\addtocounter{enc}{-1}
%	\arabic{enc} & {The characters find a mysterious magical item.} \\
%	\addtocounter{enc}{-1}
%\end{longtable}


\section{Side Quests}\label{sidequests}

Another way to add impromptu elements into your game is Side Quests.  These are short encounters which slowly feed elements into the background of your game.  They're good for foreshadowing without too much planning, and good for adding things to the Path of players who simply want to run around in a sandbox, without the constraint of a full-on plot-arc.  

Any given area can have Side Quests, such as `the baron's bandit problem', or `Why all the monsters?'.  When players enter the area, they encounter a random quest and start with part 1.  If they bump into the same encounter a second time, they get part 2, then part 3, until the entire mini-arc has completed.

Side Quests divide the landscape into various areas.  If you want to host a campaign in a city, you might make the areas \textit{the slums}, \textit{the citadel}, \textit{the market}, and \textit{the sewers} -- each with their own set of encounters.  For a grander campaign set in the Pebbles, you might have a series of Side Quests where each island and the sea itself contains a series of tales.

Quests jump from area to area.  Characters journeying into the town's sewers may find it full of green, animate slime-creatures, and the encounter jumps to the slums.  Each time they journey to the slums there's a chance to see the green monster claw its way up from below to devour peasants.  The encounter moves next to the market, where the characters hear terrible rumours of slime coming up to devour peasants.  Finally, if the characters don't journey down to erradicate the threat, the slums are completely enveloped by a horde of these acidic monsters.

If you want to run Side Quests as a secondary part of your game, you can just run them any time the group doesn't get a random encounter.

If you want them to be the primary mover in your campaign, you can run a Side Quest every time the group enters a new area.  You can also make one plotline the \emph{primary} quest.  The primary quest should have a conclusion that the characters are interested in, and it should shift from area to area, which will ensure that it takes longer to complete than the other quests.

However you run them, players should each receive 5 XP for completing a Side Quest for each part it contained.  A 2 part Side Quest grants 10 XP, while a 4 part Side Quest grants 20 XP.

For a few examples, see page \pageref{forest_encounters}.  Side Quests listed with ``\ding{106}'' increase the local nura rating by 1.\footnote{See page \pageref{nura} for more on the nura.}

Most Side Quests come with some `box text', suggesting how the encounter might play out.  This is only for illustrative purposes, and won't necessarily work for your group.

\section{Random Starting Conditions}

For players who want some inspiration for their character's background, or just more randomness, and for GMs looking for a little inspiration for their NPCs, there is a table of Random Starting Conditions.  If you want to offer this to your players, tell them to roll 1D100 and read the result to the player.  Remind players that if they are going to roll then they should be ready to accept both good and bad results; this system is here to provide an interesting result, not a beneficial one.

If the player rolls twice and achieves two rerolls, they may find their parents died twice -- this is fine.  Grandparents, god parents and adopted parents all count as parents; other examples may have to be modified to retain consistency.  No more than three starting conditions should be applied, no matter how many times the player achieves a reroll.

	If two players obtain the same result, they share the same background, as per the Story, `It'll be just like the old days'.  If their parents were both killed in the same event then they are from the same village or perhaps siblings.  They might both have worked on a similar job or share a magical item which they found at the same time.

\setcounter{list}{0}
\begin{list}{\addtocounter{list}{1}\textbf{\arabic{list}:}}{\raggedleft}

	\item{ You have murdered a travelling alchemist in cold blood to steal his magical item but do not know its command word.  It is an alchemical wand, encrusted with expensive rubies, which holds 3 MP at maximum and regenerates 2 MP per scene.  Once the command word is uttered it emits a fireball up to 6 squares away, dealing $1D6+1$ Damage across 3 squares.   This costs 2 MP from the wand's store.  The GM secretly decides on a word in English which activates the wand.  If you ever state the word while within `earshot' of the wand, it activates.  Alternatively, you may find out the command word by simply trawling through all the words you know -- this requires an Intelligence  Academics task, TN 10, which can only be attempted once per downtime. }

	\item{ A prophecy from a high priest of Ohta decreed that you were fated for great things.  You might not follow this god -- you may not even approve of Her.  It matters not a bit to the church how you feel, all they demand is that you carry their sacred weapon.  The sword can cast any spell as per the fourth level of the Conjuration sphere.  If it is ever taken from you, the sword teleports back to your hand as soon as it has a sufficient amount of mana.  It holds 4 MP at maximum and regenerates 2 per scene.  It counts as having an Intelligence bonus of 4.  This sword also grants you a +4 bonus to all social tasks involving those who follow Ohta as it is famous among her scriptures.  You are required to give the sword a name, just as those who used it to fight against the encroaching nura during the nura wars each gave it a name.}

	\item{You are exceptionally well travelled and have a million tales to tell and friends you have met.  You gain an additional 3 Story points but may only spend 2 at a time. }

	\item{ You are wanted for stealing from the mages' guild.  You snuck into the forbidden parts of \gls{college} and took an item.  The item is long gone, as is most of the money you might have collected for fencing it, but there is a permanent illusion of you sitting in the middle of the mages' \gls{college} so everyone who has ever been there knows your face.  The reward is currently up to 200 gp. }

	\item{ Your identical twin sibling trained as a magic-user of some sort and opened a portal from somewhere in Fenestra.  Of course this is a criminal offence -- there are portraits of your face all over the guard stations in every city in the kingdom.  If you ever spend Story points, you will be recognised at some point during that adventure. }

	\item{ Your cousin is the head of the Invocation School in \gls{college}. Some mages may hate you for this, others will respect you, or at least pretend to in order to garner the good graces of the \gls{college}'s higher ups. }

	\item{ You were once sold a sword by a gnome.  He says that it is also a key to a grand treasure, and his story, while vague, seemed very convincing.  The sword has several odd grooves and cuts around the base, but otherwise functions as a normal short sword. }

	\item{The elves of the Quennome forest count you as a friend after you saved one of their number from a woodspy.  They even have a song about you. You receive a +2 Bonus when interacting socially with the elves of that region.}

	\item{Your hometown had a bad-tempered, authoritarian captain of the guards. He killed your brother over an insult, but by the time you arrived back in town he had left as he was re-stationed.  The GM has 80 XP to create this character.  Additionally, his base FP is always equal to yours plus 5.  If you ever take your revenge, you receive 10 XP in addition to any for the assault.  You must land the killing blow in order to take this XP.}
	\item{After offending a priest of Laiqu\"{e}, the forest watches you, and waits to attack.  The next time you enter a wooded area, the GM rolls twice on the encounter table.}

	\item{Your mother died of plague shortly after your father was murdered by bandits.  Your family was split up, some being put into forced labour, others imprisoned and still others just went to roam the land to find what they could.  Your mission is to reunite your family.  Once every two adventures the GM rolls $1D6$ -- on the roll of a 1, you find out a sibling has been killed, on the roll of 2-4 you hear of a member of your family who must have your help.  On the roll of 5-6 you receive an additional Story point and must immediately introduce a character who is a long, lost sibling.  In total, you have $2D6$ family members to find.  Each lost sibling you find grants you 3 XP.  If you find and successfully reunite all of those still living, you gain an additional 10 XP.}

	\item{You once fell into an ethereal portal to another land by accident, and wandered through strange lands for years.  You gain 2 Story Points that can only be spent in the ethereal lands.  Additionally, if you ever spend a Story Point on finding a mana lake, that mana lake contains a hidden portal to one of the ethereal lands.}

	\item{You have come a long way from your little home on the little islands around the Pebbles.  You start with the first level of the Seafaring Skill.}

	\item{You are a halfbreed.  Roll $1D6$ to determine your race randomly.  You gain the racial abilities of the first race listed.  Halfbreeds are infertile, like mules.}

		\setcounter{enc}{1}
		\begin{list}{\arabic{enc}\addtocounter{enc}{1}-\arabic{enc}\addtocounter{enc}{1}:}{}
			\item{Human/ Elf: You are a half-elf.  You have no racial penalties or modifiers.}

			\item{Dwarf/ Gnome: You are half gnome, half dwarf.  You gain 1 Intelligence but -1 Strength.}

			\item{Gnoll/ Human: You are half gnoll, half human.  You gain 1 Strength but -1 Charisma.}
		\end{list}

\item{As above, but gain the racial abilities of the second race, not the first.}

\item{You come from a cold, harsh mountainous region in the North of Whiteland.  You start with the first level of the Survival Skill for free, with a specialisation in Snow Tundra.}

\item{Your auntie was a planeswalker -- she travelled to distant, strange lands and brought back tales so fantastic that most thought she was mad, but you listened to each and every one of those tales of the strange lands outside of this world.  You gain the first level of the Ether Lore Skill for free.  With a successful Intelligence roll at a variable (but high) TN, the GM must give you basic information about any plane of existence you find yourself in.}

\item{You are born cursed with the taint of the nura.  Perhaps you were involved in a nura battle as a youth.  Perhaps your mother briefly changed into nura and then saved by a priest while she was pregnant with you.  You begin with the first level of the Saurecanta school for free and are on the Path of the Nura.  You may gain additional magic schools as normal.  You do not know what the different levels of Saurecanta do -- you can only find out by increasing this school with XP.  You can increase and know about any other magic sphere on the Path of the Nura as normal.  You may not gain XP for following a god at any point unless you make penance for your inherent sin and forsake all casting of nura magic for a year.  This addition does not come with MP, so you cannot cast spells unless you buy base MP or your Intelligence is already high enough to grant you free MP.}

\item{You grew up without parents, raised by your older sister on the streets of the city of Lakesend in Whiteland.  You gain the first level of the Larceny Skill for free.}

\item{Your uncle was a famous Academic, and travelled the land before returning to your people to tell them what strange lands lay about them.  You start with the first level of the Academics Skill for free.}

\item{You once, in jest, insulted a dwarven woman who was travelling through your homeland.  She could not break the law by ordering your execution on the spot, but has hired a dwarven assassin to find and kill you.  If that assassin dies, another will replace him.  This assassin is a standard dwarvish warrior with Stealth 2.  They will not stop coming until you are dead or you make a very outlandish apology to the dwarven lady in question.\footnote{See page \pageref{dwarven_soldier} for stats.}}

\item{You have some small speck of draconic blood in your veins as an ancestor mated with a dragon, polymorphed into a member of your race.  You begin play with the first level of the Invocation sphere and are initiated automatically onto the Path of Blood.  If you do not have any mana, you cannot use this spell.  You may instinctively learn any other sphere from the Path of Blood by spending the necessary XP.  Additionally, you gain DR 4 against all forms of fire.}

\item{You proudly crafted your own equipment before beginning the adventure.  You begin with the first level of the Craft Skill for free and select either a weapon, projectile or anything else worth 10 gp or less to begin with, for free.}

\item{You were mauled by a beast when younger and have never forgotten it. You have nightmares about the creature each night which leave you Fatigued.  Roll an Intelligence based task against TN 7 each day -- if you fail you begin with 4 Fatigue points.  You can stop the nightmares only by finding and killing the creature from your dreams in real life.  Roll randomly:}

	\setcounter{enc}{6}

	\begin{list}{\arabic{enc}\addtocounter{enc}{-1}}{}
		\item Dragon

		\item Basilisk

		\item Archmage

		\item Umber Hulk

		\item Chitincrawler

		\item Griffin
	\end{list}

\item{One of your parents was lost in the Etherial realms beyond while exploring there. If ever you find them and bring them safely back home, you gain 15 XP.}

\item{Your home village was constantly assaulted by nura beasts.  Eventually it was entirely destroyed -- you managed to flee, one of the last survivors. During your time there, you were forced to learn how to shoot a bow to defend the place -- you begin with the first level of the Projectiles Skill for free.}

\item{You were born on Qualme's holy day.  The god of death will not touch you until your time.  Gain a 2 bonus to any Vitality Checks you make.}

\item{You were born during the height of Alassea, and that goddess has blessed you.  You begin with the first level in the Empathy Skill for free, reflecting your natural knack for charming people.}

\item{You were born at the height of C\'{a}lea and have always been one for picking up little factums here and there.  You begin with the first level of the Academics Skill for free.}

\item{Born during the holy day of V\'{e}re, you are known to be true to your word and are known by all the priesthood of V\'{e}re, who are always only too happy to help out where they can.  You can spend 1 Story Point every second story to introduce a priest of V\'{e}re, as per the story `There is a man whom they call \ldots'}

\item{Born during the festival of Laique, you have always had a natural affinity with animals.  You begin with the first level of Beast Ken for free.}

\item{You were born during the festival of Ohta, while your entire village was outside in the public square having a series of brawls to settle old disputes.  You begin with the Brawler Knack.}

\item{You are among the last of the Whiteland nobles who were hunted almost to a man after threatening the king's castle wing which sits there.  You begin with the first level of the Deceit Skill as you are used to hiding your heritage.  There is a 40gp bounty on your head, so you had better keep your identity secret -- this will be easy enough around common people, but you might be hard pressed to tell a convincing lie to nobles, many of whom have a good working knowledge of accents and the precise customs of each land and the family and underlings of the landmasters there.  There are nobles in Whiteland among each race, so if you are not human, simply assume that you were the chief family in a village in Whiteland which was destroyed.}

\item{You served in a war some years ago against the South Kingdom.  You fought mostly at sea, but did not do terribly well.  You lost one hundred men due to your poor military decisions and were unceremoniously thrown out of the military.  You begin play with the first level of the Tactics Skill.}

\item{You grew up in a travelling circus -- perhaps as part of a freakshow, perhaps as a magical act.  You start with the first level of the Athletics Skill for free.}

\item{Your father was one of the most famed warriors in the land during the recent war with the Southern Kingdom.  You begin with the Knack: Adrenaline Surge which gives you 1 Damage once per scene whenever you remind yourself of your legacy.}

\item{Your home village was destroyed by goblins when you were younger.  You begin play with the Knack: Chosen Enemy (Nura).}

\item{You grew up on the borderlands between gnolls and humans.  If you are a gnoll, you have the Knack: Chosen Enemy (Humans).  If you are human, gnolls are your chosen enemy.  If you are neither, choose one.}

\item{You were found as a small child and taken in by a kindly couple -- you have no idea if your parents were killed or you were simply abandoned. Roll on the racial chart to determine your parents' race.  Their race determines your background culture.}

\item{You were once captured and used as slave labour in the Southern Kingdom.  You have escaped but now have the Knack: Tireless.}

\item{Your parents were sentenced to death after a legal dispute with a member of another race -- roll randomly on the race chart.  You now have the Knack: Chosen Enemy (race).  If you roll your own race, roll again.}

\item{You were once on the run for a petty theft for over a month.  Almost everyone has now forgotten about it, but the memories of that time, and what you had to do to survive still haunt you.  You begin play with the Knack: Perfect Sneak Attack.}

\item{Everyone gets teased during childhood but yours was beyond severe.  The constant fights have taught you a thing or two however -- you begin play with the Knack: Fox Hop.}

\item{When you were only four years old, you got in a fight with the family dog over a scrap of meat.  You took a blade and killed the dog without a thought, then prized the meat from its still-warm jaws.  You do not recall this event but your mother recounted it every time you misbehaved, reminding you of how evil you are.  You begin play with the Knack: Finishing Blow.}

\item{The military unit you served with have all died -- the undead took most of them, the diseased wounds took others, and madness finished off the rest.  You alone remain, and you miss the days of rounding up simple criminals, giving them a little beating and then throwing them in the cells.  You begin play with the Knack: Expert Flanker.}

\item{You have been assaulted and robbed before.  Some people have had it worse, but you have never been able to let go of the incident.  Perhaps it's because of the scar across your face.  You begin play with the first level of the Vigilance Skill.}

\item{You were kidnapped by a necromancer when you were just a baby.  Your father managed to get you back but you have never been the same since.  Gain the Knack: Consume Soul, even if you have no MP to use it.}

\item{You have a mysterious Deck of Cards which you know to be magical.  Each time you draw one a spell flies out.  See page \pageref{deckofmanythings} for the item.}

\item{You are in love with someone completely inappropriate -- perhaps they are above your station, perhaps from a family who detests you.  If ever you manage you successfully marry this person, gain 20 XP.  The GM is obliged to make this challenging.  You can write letters to this person and if you ever find yourself in your homeland, you will automatically encounter this person.  You must flesh them out a little but the GM will create the exact Traits.}

\item{You begin as a blank slate -- total amnesia.  Assign your Skills as normal. Each time you spend a Story point, a memory resurfaces or someone from your past who knows you arrives.  You must slowly piece your past together from these fragments.  You can allow the GM to decide on a past and give the pieces to you if unsure of what it should be.  Leave 30 XP from your starting XP unspent.  You may spend this XP at any point to gain anything but an Attribute Bonus.  At the time you spend this XP, you remember an additional piece of your past.  You begin play with two pieces of equipment worth up to 40gp; only one of these can be weapons or armour.}

\item{As above, but the GM secretly rolls again on this chart and determines a second background.  This background is unknown to you unless it is obvious.}

\item{You are old.  If you are an elf, you are 100 years old.  Apply the normal Attribute adjustments.  Otherwise, your maximum Body Attributes are decreased by 1 each but you begin with an additional 20XP.}

\item{You share a background with the player to your left.  If they own a magical item, you own it too (or there is some argument).  If they would be the last member of their people, you pair are the last of your people.  You can share the cost of any Stories you tell.}

\item{Your parents were wiped out when an arrogant alchemist strolled into town and refused to pay the inflated prices for nobles.  He incinerated them without a thought and then bribed the local guards to avoid punishment.  You begin play with the Knack: Chosen Enemy (Alchemists) and receive 10 XP if you ever track him down.}

\item{Runes have always fascinated you.  You begin play with the first level of the Academics Skill and a specialisation in Runelore.  You are on the Path of Runes and can purchase any sphere from it.}

\item{You used to work for the temple of V\'{e}re and were turned into a human mana-conduit.  You now store 2MP, available only to priests of the temple of V\'{e}re.}

\item{You were once part of a travelling band of bards, hired as an actor.  You had to put costumes together on the fly.  You begin with the first level of the Deceit Skill.}

\item{You once worked for a crazed mage named Adrian the Red who was banished from the mages' guild -- he never declared why.  During this time he gathered a terrifying army of nura.  He was eventually wiped out, and luckily for you, by this point you had already left.  Every second story you tell using your Story Points must involve either Nura, the undead or mad mages studying forbidden magics.  If you know a blacksmith, he might be a hermetical nura hobgoblin living alone in the forest and coming out to trade wares.  If you know a cut-throat, this might be an ogre who has managed to survive by convincing a village to constantly deliver meat to him.  If you know a powerful mage, this might be an undead priest who sits alone in a chapel, praying to Qualme each day to release him from his torment.}

\item{You originally come from the Southern Kingdom.  You may not spend your original 5 Story Points while in Fenestra, only while in that other land.}

\item{You travelled round the South Kingdom for some time -- you start with 4 Story points to be spent in the North Kingdom and 3 Story points which can be spent in the South Kingdom.}

\item{Your father has been put in prison for murdering a nobleman, but you are sure he is innocent.  You receive 20 XP for getting him out, one way or another.}

\item{You were once travelling with a band of merchants when a horrifying forest creature attacked -- the GM rolls $1D6+12$ on the Forest Encounter chart to determine the exact creature; you now have a crippling phobia of this creature (if the result is a mana lake, reroll).  If you encounter this creature, you must make a Wits check, TN 8, each round; failure indicates you must flee immediately and swiftly.  If you ever deal the killing blow to such a creature, the TN to overcome your phobia reduces by 1.}

\item{You once saved a priest of V\'{e}re from being attacked by bandits.  He blessed you with the safety of the forest from that day onwards.  You store 1 MP like a mana stone and regenerate 1 MP per scene.  If you spend this MP, you can attempt to send a forest creature to sleep by saying a very short prayer to V\'{e}re.  The TN is 4 plus the creature's Wits  Aggression.}

\item{Your little sister once stole a ring of invisibility from a priest of Alass\"{e}.  You thought it was quite funny until he took his vengeance, turning her into a cat.  He got his ring back and walked away, leaving you in shock.  The cat follows you everywhere.  It has Str -5, Spd 4, Dex 3, Int 1, Wts 1, Cha 1.  If you ever find a miracle worker who can undo the spell (TN 10) you gain 10 XP.  Until that time, you have a constant cat companion.}

\item{You were once married, settled, and happy.  Happy until a couple of years into your marriage it transpired that you are infertile.  Your partner left you and now you must carve out some kind of life as an adventurer.  This is not what you envisioned for your life at all!}

\item{You are in possession of a book you found on the mutilated body of a man on the road.  The bandits took all his goods but left the book out of disinterest.  It details extremely advanced alchemical instructions on how to summon creatures from other worlds.  This book is worth up to 200 gp to the mages' guild.  It can be used to learn the 5th level of the conjuration sphere through the Path of alchemy.}

\item{You were hired for a battle with a powerful priestess of Qualme but were so terrified of her undead host that you fled before the battle.  While fleeing you accidentally encountered the undead host on the road and had to hide in the surrounding forest.  Then the host stopped, and the priestess left her wagon in a forest clearing before continuing towards the battle. After you were sure they were gone, you crept up to the wagon and stole the first thing you could see -- a painted skull, sitting on the table.  Since then you learnt two things -- that all of your companions are dead (and you have been presumed dead with them) and that the skull you stole comes from an old priest of Qualme.  His spirit remains in the skull.  He cannot speak but can summon the dead back to life.  He has a base MP of 6, Intelligence 3, Wits 1, Necromancy level 5, Invocation level 3, Fate 2 and Aldaron 1.  He can affect someone with Necromancy level 4, allowing them Death Sight, at which point the target can see and hear the necromancer.  His current wish is to return to a priest of Qualme.  He is not a cooperative magical item for others' use but a creature in his own right.  He has only a limited fear of death as returning to this land would be difficult but not impossible.}

\item{You once spent a month in cells being tortured before the guards realise they had the wrong person.  Begin with the Knack: Hardened.}

\item{There is a prophecy that you will somehow interfere with the will of Qualme -- you have no idea what this is about but some cult or other out there are not happy with you.  Someone has sent an undead knight to kill you.  It instinctively knows roughly where you are.  It follows you on foot. You keep moving, on horseback (begin play with a horse) whenever you can, every day.  You are safe only in populated areas, where it usually does not approach.  See page \pageref{ghast}.}

\item{You serve a minor deity from your home village and wish to spread word of their glory throughout the land.  You must select their name and personality.  You will work with the GM to create a list of pious endeavours which grant you XP.  If you prefer, this can be a unique Code.}

\item{Do not read this aloud -- pass a note to the player saying that they were secretly involved in the back story of the player to the right.  Perhaps the powerful magical item they have was going to be used for a nefarious plan and should be destroyed.  Perhaps they have some connection to an enemy of the character's.  Perhaps they know some terrible secret about the other character's history.}

\item{You once destroyed a powerful artefact from some temple -- perhaps a sacred tree in a temple of Laique, or a doll from a temple to Alasse. Whether this was on purpose or by accident, that god now has it in for you.  Every time you are outside of a city and obtain a result of `no encounter' the GM rerolls once in order to try and obtain an encounter. This continues until you somehow make up to the god (via the church) for your terrible transgression.}

\item{You are addicted to alcohol.  Any time you have a strong drink -- ale does not count -- you must make a Wits roll, TN 8; failure indicates you must have another drink.  Each drink increases your Fatigue and increase the TN by 1 and you must roll again for another, and another, until you prize yourself away or are too drunk to drink any more or you are forced to stop drinking by some other means.}

\item{You started your adventuring career with a heartfelt gift from your father -- the family warhammer.  Apparently it has been in the family for three generations, and now four.  You said you were not sure if the weapon fitted your exact adventuring style, but your father only told you that you would get better in time.}

\item{Your family are rich, and very fond of doting on you.  Begin with 50 gp and 100 gp worth of equipment.  This can include an entourage who will join you for the first adventure, and possibly a few more -- perhaps jesters, cooks, a few soldiers and maybe the services of an alchemist for a single session.}

\item{Your grandmother, high priestess of Laique, died recently and made you promise to look after the family cat.  For those who follow the Path of Divinity (Laique) it counts as a 5 point mana stone and can Polymorph people as per the 3rd level spell, Polymorph Other.  It has a range of 6 squares and uses your late grandmother's Intelligence bonus of 2.  The cat has Str -5, Spd 4, Dex 2, Int 0, Wts 2, Cha 3.  It cannot speak and the GM makes all decisions about when it will use this power according to the interests of the cat, not the interests of your character.}

\item{You have been captured for a crime you did not commit and were shoved naked into a cell.  You have managed to escape and become a stowaway on a travelling wagon to get where the adventure starts but you start with nothing -- not a sword, piece of bred, spellbook or loincloth.  You begin with an appropriate number of Fatigue Points if you start in a cold region.}

\item{ You are very young.  You begin play with only 20XP.  No Trait may be higher than 1, although FP and MP can be at any level.  You may raise no Trait to above 1 during play.  During your first downtime you enter a growth spurt and gain 40XP and can spend any saved XP. }

\item{ Your mother had to give birth unexpectedly while travelling.  The camp did not know it at the time but they had stopped by a mana lake.  Since then your eyes have had an unnatural shine to them, begin with 2 MP base -- this is a free level but the next level will cost 10 XP as usual. }

\item{ Do not read this aloud.  One of the player's pieces of starting equipment is cursed -- it summons aggressive animals as per the second level of the conjuration sphere at inopportune moments.  It casts with 3 MP and an effective Int 1 and Wits 0. }

\item{ You own a lucky coin.  It looks like a gold coin from the Dogland region, but Gregor's face is twisted with an unnatural smile and you happen to know it is not real gold.  You can flip it to request the answer to a yes/ no question and the GM must give you an answer. This is a magical item, dedicated to Alasse, and must advise the player of her will, not necessarily the best course of action.  It works once per scene and the rest of the time gives random answers }

\item{ You are cursed.  Random members of your family have died, previous adventuring companions have died.  All of your friends have died and you have sauntered through life taking all their good luck.  If you ever take a Vitality Check, you can request a 4 bonus before the roll.  If you do so, the next companion who takes a Vitality Check takes a -4 penalty.  All Vitality Checks are rolled by the GM in secret.  You must indicate whether you are taking the 4 bonus by handing a secret note to the GM. }

\item{ You have been captured by a dwarvish necromancer so that your dead body could be used as the vessel for an ancient warrior-wizard who is destined to return and destroy the King of the North while reclaiming all the land around the Bearded Mountains for the dwarves.  Unfortunately for them, two things went wrong -- the first is that you were not actually dead when the spell started (just close to death), the second is that other dwarves came to interrupt the illegal ritual and kill the necromancer.  They set you free, the spell completed but with a strange twist.  If ever you die, that powerful spirit shall return and inhabit your body with full access to powerful spells and begin amassing an army and increasing his own personal power by drinking the souls of the dead.  The dwarves explained to you that they thought this was a likely result but let you go because you were not guilty of any crime. }

\item{ You once served under your grandmother as a village healer, and later went to serve in the militia to fight in the Whiteland wars of humans and gnolls.  You begin play with the first level of the Medicine Skill. }

\item{ You once fought as a gladiator in the arena of a city.  You were quickly thrown out for your irritating ability to defeat opponents without drawing blood.  Begin with the Knack: Disarm. }

\item{ You once fought in a gladiatorial arena but managed to escape.  If ever you use a Story point or journey to a large city, you are recognised, though not everyone will have an interest in telling local guards where you are - technically, escape is not illegal.  However, if your captors ever want to get you back they have the paperwork to show that kidnapping you would also not be illegal.  Begin with the first level of the Combat Skill. }

\item{ You once worked as a court jester -- you wandered different noble courts, even entertaining the king on one occasion.  People took you for an idiot, but during that time you learnt a lot.  You gain 1 Charisma and have a minimum Charisma score of 1.  This cannot take you above your racial maximum.  You also gain an additional Story you can tell -- if ever you meet a noble, you can spend one Story point and say `Didn't I see you in his majesty's court that time when \ldots}

\item{You enlisted to fight with one of the most famous heroes of the land -- Elric Redthirst, paladin of Ohta, to fight against encroaching nura horrors.  You met with the opposing forces on a hilltop, chastised and slaughtered many of them, then chased the remaining host back into their cavernous lair and destroyed each of them to a man.  The forces were so equally matched that everyone died -- everyone except you. You picked up Elric's sword -- a beautiful yet simple thing made from the finest steel, enchanted to destroy any creature.  It is known as The God Ruiner.  If the church of Ohta find out you have taken the sword, they will hunt you to the ends of the earth for it.  At present, nobody knows you well enough to be looking for you -- they probably assume that some nura survived and stole the artefact.  However, almost all followers of Ohta and many others know exactly what the sword looks like -- that distinctive purple colouring along the handle and the words of vengeance carved into the hilt are difficult not to remember.  The sword stores 5 MP.  If it ever inflicts 12 or more Damage on an opponent (before DR is applied), it can errupt with a level five invocation spell, inflicting an additional $2D6+3$ Damage which bypasses all DR, including Perfect DR.  It otherwise functions as a normal great sword.}

\item{You have been hired by priests of C\'{a}le to chronicle your adventures in the land.  You are largely expected to make your own way and earn your own money, but any time you are in a large city you may report to a temple of C\'{a}le and receive 2 gp per month of adventuring time, or 6 gp per season.  They want all the information on the various secrets of the land that you can obtain for them -- hidden grottos, conspiracies, uncharted mana lakes, anything new and interesting.}

\item{You once wished upon a shooting star during the cold festival at the height of Alassea. You wished for all your dreams to come true.  Since then they have -- sort of.  Any time you sleep the GM secretly rolls 2D6, on the result of 10 or more an illusion is created out of your dreams.  This uses any level of the illusion sphere, up to 5th, and counts as if it were cast by someone with Intelligence and Wits 5.  The GM crafts a dream out of elements and characters from the current adventure and from your background stories -- if you have already specified you have known a blacksmith from childhood and you are currently running away from nura creatures, that blacksmith might arrive after having been transformed into an ogre.}

\item{Your grandmother was a village healer and taught you all you know about the craft. You later served in Whiteland during the war between gnolls and humans.  Begin play with the first level of the Medicine Skill.}

\item{You have been enchanted to serve as the personal body guard of a high priestess of V\'{e}re.  However, you managed, in a drunken stupor, to leave long enough for the enchantment to dim a little.  Your mistress is still looking for you, and the more legend spreads of your deeds the more intent she will be on finding you.  Any time you spend a Story point, your identity will be known by someone during that adventure.  If you ever meet your mistress again, you will not be able to refuse her -- you must obey her every command.  Mechanically, she has used the 5th level of the Enchantment sphere on you.}

\item{You were a high ranking member of a guild before deciding to pack it all in and adventure.  If you are not human then you have some equivalent expertise in your own culture.  You begin play with an additional $2D6\times4$ gp.  The guild type is determined by one of your highest Skills.}

\item{Your father gave you a map to some burried treasure, but wouldn't tell you how he got it, only that you are not the only one with a copy of the map.  Each time you spend a Story Point, roll a die.  On the roll of a `6' you run into someone chasing the same treasure as you.  At this point you can make an Intelligence + Survival check to understand how the map relates to the local landscape.  The GM will fill in the details.}

\item{If you roll \arabic{list} or above, reroll twice}

\end{list}
