\chapter{Tools \& Toys}

\epigraph{As far as the laws of Mathematics refer to reality, they are not certain; and as far as they are certain, they do not refer to reality.}{Albert Einstein}

\section{Mana Lakes}\label{mana_lake}\index{Mana Lakes}

\begin{multicols}{2}

Throughout Fenestra there are little nodes where mana wells up from the ground like a wellspring.
In these locations, some number of MP is created each round which then raises the current MP of mages or nearby magical items.
Whoever has the most empty mana slots - whether an item or a person - gets that mana point.
It pours into the biggest vacuum like water into a hole.
So if a gnome with a current maximum of 6 MP spent 4 MP, that would mean they had a deficit of 4.
Another mage who had only spent 1 MP could only get a point after that gnome.\footnote{If two are tied for having the biggest deficit, ties are broken by the win going to the highest current maximum MP holder, then the highest Wits Attribute holder, then the highest Intelligence Bonus holder.}

Mana lakes also extend the range of all spells cast within to the entire mana lake at least as spells travel easily from one end to the other.  Each has a level, denoting the number of MP it doles out each round, so a mana lake of level 3 would give out 3 MP per round.

Magical items regenerate mana just like people do, so many miracle workers create magical items especially to sit in mana lakes, and recharge quickly.

A selection of such mana lakes with magical items (or some other permanent magical effect) are given below.
If your PCs encounter a mana lake on the Random Encounter tables, or you want one for one of your own stories, just pop one of the following into the map.

\subsection{King's Hand Valley}

There is a valley which was once filled with bandits who would pillage local villages.
To put a stop to it, a priest of Qualm\"{e} cursed it with a mummified hand of former King Dalius Quennome.\footnote{See page \pageref{h_dalius}.}
Nobody has any idea how he obtained the hand.

\textit{Level: 2}

\begin{boxtext}

	Your arrow hits the bandit leader in the face, killing him instantly.
	The other duck down behind the hill, then screaming starts and all of them jump over the top, running towards you, despite your drawn bows.

	As you loose more arrows, you see the leader you shot stand back up with an arrow still protruding from his eyesocket and blood smeared from his mouth downwards. Another stands up, his throat torn out, and both move quickly towards the bandits running towards you.

\end{boxtext}

If, at any time, a dead humanoid body is in the area, it rises from the dead to eat the souls of the living.  Most of the time such dead creatures like to stay and feast off the ambient magics in the area.  They hunker down under snow, or red, squelching leaves, or simply lie and let the moss grow over them.   At first impatient traders would attempt to sneak across the valley as a short cut.  Later, only a few brave warriors would come to test their strength.  Now none come, and the dead are being eaten away by the moss.  But they will not leave.  They simply remain, and wait to hear human speech again, and feel human warmth.  It has been a long time since their last meal and they have gathered a decade's worth of hunger.

The hand hangs from a tree, and looks very much like a hanging vine or a part of the tree it hangs from.
It holds a total of 7 MP and spends 2 to raise anyone in the area as a ghoul as per the second level of the necromancy sphere Calling the Dead.
It automatically activates instantly for as long as it has MP to use.

\subsection{The Petrified Forest}

\textit{Level: 1}

One day during the season of C\'{a}lea, a great flood came and washed a marsh clean out.  Mud trickled out for the entire season and when all was done, a petrified forest was uncovered.  Stone cylinders from one to six feet tall were uncovered throughout this mile long valley.  The valley is currently covered in grass and fungi but is otherwise uninhabited by any vegetation.

Around the centre of the valley, anyone mentioning something will soon see an illusion of that thing.  The illusions each have a total of 6 points to divide between Speed and Dexterity Bonuses, but have no other Traits.  People moving through might imagine what the forest was once like might see the petrified trees turning into real trees.  Those worrying about monsters will see an illusion of a monster.

The illusion holds 15 MP in total and can spend 3 at any point in order to create an illusion of up to 5 squares size.  Basic illusions are TN 11 to identify as illusions while larger illusions are only TN 9.

\subsection{The Myriad Web Forest}

\textit{Level: 1}


\begin{boxtext}

	You continue towards the sound of the crying deer, and brushing foliage aside you can finally see the great tree, a stone's throw away.
	Half way up the gargantuan tree, a giant, fat spider looks down, clasped to various branches.
	Beneath it, another giant spider is bound to the tree by a great web with two leggs sticking out the side.
	Beside it, bears, deer, and one chitincrawler are trapped in more webbing, all around the tree.

	The bushes start to wrustle, and man-sized arachnids pour out from every direction.

\end{boxtext}

In the middle of a great forest sits a massive and majestic shiva tree.
A guardian spell was placed on it long ago by a priestess of Laiqu\"{e} such that if anyone enters the area, marauding animals are summoned to the area.
This is no problem to an initiate of Laiqu\"{e}, as they can typically charm such animals -- but anyone who was not a friend of the forest could be in serious trouble.

Since that time, a massive nura spider has taken up residence in the area.  This spider, grown to gargantuan size and twisted by strange magics, has made its home here.  It perpetually traps summoned animals in its web then eats them.  The summoned animals do not leave until they have been killed or leave the area.  However, the nura spider does not simply kill them but bind them with its web and slowly liquefies them with its poison.  This blurring of the boundaries of when the animal is dead or alive, of where the web ends and the creature begins, means that the animals so summoned remain after the spell has ended.

At any given time, its web is full of up to a half dozen assorted creatures -- griffins, aurochs, a bear and such.
The web must be laid slowly, but once on the ground or stretched between branches and tree trunks it is extremely strong.

\npc{\N}{Gargantuan Nura Spider}

\animal{6}% STRENGTH
{2}% DEXTERITY 
{3}% SPEED
{0}% WITS
{4}% DR
{3}% AGGRESSION
{Athletics 2, Survival 1, Tactics 2}% SKILLS
{\web}% ABILITIES
{}

During her time here, the spider has become incredibly large and fat and has given birth to many children.  Some move away to terrorise nearby farms.  Others remain and usually starve to death after failing to fight her for food.  She has one mate -- one male spider with whom she mates.  When hunger overcomes her she sometimes eats another section of his legs.  Currently, he has two and a half legs left.  He cannot move properly as she she has bound him in webbing and left him on branch near the top of the shiva tree, hanging like her prey.  Every so often she approaches the top to mate with him, or to stop one of her children attempting to eat him.

\npc{\G\N}{8 Nura Spider Children}

\nuraspider

The shiva tree holds 11 MP in total and can conjure animals as per the spell Forest's Call.

\subsection{The Wishing Wellspring}

\textit{Level: 1}

Long ago gnomes who lived near this area thought it would be fun to create a functioning wishing well for anyone who needed it.
They found a natural wellspring deep underground and created a layer of polished stone on top, complete with a gazebo.

In the centre of this stone floor placed in the middle of a large plain is the well.
The metal bucket and ropes were reinforced with steel wiring and function to this day, though they could do with being replaced.

\begin{boxtext}

	The wishing spring lies just ahead.
	The roof has deteriorated badly, but the stones laid all around still shine brightly.
	All the rocks are smooth, and as many are gemstones the entire platform glistens in the sunlight.
	Then you notice a foot sticking out the side.
	It belongs to a fat man whose face has been completely burnt off.

\end{boxtext}

Any time someone makes an audible wish, puts the well's bucket down and drops a coin into the well, a summoning spell activates and creates the thing they wished for inside the bucket.
When it is drawn back up, the wisher can grab the item.

Some time after the well's creation, a priest of Qualm\"{e} came and noted how some people were disrespecting the well and were using it too commonly.
Their wishes threatened to break the magic which allowed the well to function.
He decided to protect the well with a spell of his own.
He carved a message in elvish -- the tongue of all academics -- into the side of the well.
It proclaims ``Mine m'{e}re ilyain er'' - ``Only one wish for everyone''.
The skull of an older priest of Qualm\"{e} was placed at the bottom of the well to guard against misuse.
If anyone attempts a second wish a storm of lightning and fire erupts from the well and spreads across a 5 square diameter, dealing $2D6+2$ Damage to everyone there.

The well holds a total of 16 MP and an use any of the first four levels of the Conjuration sphere.
If the well spends mana to summon an item, the mana must, of course, remain spent, so the well will be depleted by 3 MP for as long as the item remains in use.
If the well runs low on mana because too many people are making wishes, it simply cancels the oldest spell it has cast and the wished-for item disappears, whether it is by the well or a hundred leagues away.

\subsection{The Dwarvish Prophet}

\textit{Level: 3}

\noindent Deep underground there are many natural streams, flowing quickly into the depths.  Some of these run alongside passages which the dwarves carve out between the various realms they inhabit or realms they inhabited.  It is not uncommon for underground travellers to wander along natural caverns only to later find a dwarvish road to take them part of the journey.  On one of these roads is a carving of the face of a great dwarvish prophet.


\begin{boxtext}

	The massive face has no beard, leading almost everyone to believe she was female.
	A waterfall flows down her face and has worn down everything but the nose, leaving a single, perfect nose and part of the lips while the rest of the face looks more and more like bare rock with each passing decade.
	There are various possibilities as to who the face might represent -- there have been many dwarvish prophets and almost half were female, but the exact identity has never been discovered.

	As your guide passes through, runes around the statue's head glow.

	``Exactly which happens is not well understood.'', he explains.

	``Sometimes she blesses you, sometimes she curses you, and sometimes she does nothing at all.''

	``Some say that spells themselves can become insane after long centuries sitting along in the dark.

	The guide wanders past, content with his fate.
	It's time for you to move past too.

\end{boxtext}

Those following the Codes of the Tribe or Experience receive $2D6$ FP (costing 2 MP).
Those following Qualm\"{e} or the Code of Acquisition lose $1D6+4$ FP.
Others are unaffected.
The statue head holds 12 MP in total.

\subsection{The Resting Grove}

\textit{Level: 2}

\begin{boxtext}

	Large, ripe fruits litter the forest clearing ahead, and short tufts of grass lay about and look perfect for sitting.
	Looking up, you suddenly notice various tree houses, with men looking down at you.

\end{boxtext}

This group of trees and bushes looks like a lucky find.  There are apples, blueberries, raspberries and tiny strawberries growing all around.  Even better, when one piece of fruit it picked, more grows in its place.  If it rains, the trees naturally knit their branches together  to form an overhead canopy thick enough to keep the rain out.  If people nestle up to their trunks, they can warp and open up, allowing people to sleep inside the newly created hollow.  The entire grove is a luxurious place which caters for the every need of anyone within, including creating simple items like tables or goblets out of wood.

Once people leave the area, any items they have remain as they were, but the trees return to normal.  Any items stored in those temporary shelters is absorbed by the tree, and further mutating magics will not necessarily bring the items back because the tree might not open out in the same way the second time it opens.

While this place was once a welcome resting stop for many a traveller, or at least for those brave enough to go near such strange magics, it has since been taken over by bandits.  They use the trees as a continuous supply of arrows for their bows.  They sit in treehouses at the top and shoot down to any attempting to get them out of the area.  Since the grove is not far from a major road, the local town master has made a high price on the bandits' heads.  None have claimed the bounty yet.

\npc{12 Bandits}

\humanarcher

The grove contains 13 MP in total and spends 4 in order to make any alteration or 1 MP to calm any wild animal in the area.
All spells cast count as having an Intelligence Bonus of +3.

\end{multicols}

\section{Magical Items}\index{Magical Items}

Magical items are fairly formulaic since each path of magic has a limited number of `mana stones' which they can make, and a limited number of spells.
Nevertheless a few examples sit here for quick reference.

\begin{multicols}{2}

\subsection{Alchemical Items}

\magicitem{Bag of Tricks}{Conjuration 4}{Quick}{Item}{4}{0}{4}

\noindent This gem-studded bag has a tough, silver chord around the top, and won't open more than the width of a man's wrist.  By simply naming an object, people can pull out anything they desire.  Unfortunately, the bag only speaks the gnomish language.

It stores 4 MP, and spends one produce any small, valueless item or substance, with the full use of the fourth level of Conjuration.

\magicitem{Bag of Holding}{Conjuration 5}{Quick}{Item}{3}{1}{0/5}

This looks identical to the bag of tricks, but instead of summoning items, it leads to a closet.
The bag is in fact a magical portal, held permanently open, and allows the owner to deposit and remove items at will.

Exactly how much space is available depends upon how long the user's arm is.  Inside is completely dark, and the folds of the bag (or perhaps further magic) do not allow anyone to peer into the bag -- only to fondle the various shelves resting inside.

Anyone putting an item into the bag in a hurry must make a Dexterity plus Crafts check at TN 6 or risk dropping the item instead of placing it on one of the dark shelves within.  Alchemists have theorized about whether the bag leads to a bottomless pit in some demonic realm, or simply a larder.  Either way, nobody's arm's long enough to reach the bottom.

\magicitem{The Townmaster's Hat}{Illusion}{Instant}{Item}{2}{-1}{1}

\noindent Townmaster James of Whiteland was especially scared of assassination during the time of King Dalius, so he commissioned four hats with an illusion, making anyone who put them on look exactly like him.
He has since died, and the hats have been used by various thieves to disguise their identity.

\subsection{Divine Items}

\magicitem{Arthur the Sleeping Cat}{Sleep}{Normal}{Item}{3}{0}{5}

A priest of Laiqu\"{e} once left all her worldly mana to a cat, imbueing it with a single spell.
The cat can instinctively make animals of any kind fall asleep.
He sometimes uses this to hunt mice, and sometimes to shut people up who are too loud.

\magicitem{Deck of Cards}{Various}{Instant}{Pocket Spell}{3}{2}{X}\label{deckofmanythings}

This deck, presumably fashioned by various priests of Alass\"{e}, holds a number of cards.  Once grasbed between \emph{three} or more fingers, the card activates.  If it's held, the target is the caster.  If thrown, the target is whoever the caster throws it at (the caster doesn't need to be a good thrower for the magic to work).

Each card looks much like a normal playing card, with one non-obvious flaw, such as a joker who's wet himself, or a king who's lost his crown.  The spells are each cast with an effective Intelligence and Wits bonus of +4.  Only a few cards remain.  Once one is pulled, the effects activate immediately and lasts for a full scene.

The Deck of Cards resists all magic, including gaining information about it through Fate.

\vfill\null

\begin{tcolorbox}
	\begin{enumerate}
		\item{Illusion of an ogre appears}
		\item{Nearby animals gather to the caster -- roll an encounter}
		\item{Nearest Target becomes insisible}
		\item{Nearest Target turns into a chicken}
		\item{Mist fills the area}
		\item{Nearest Target turns into a bear}
		\item{Blinding light}
		\item{Darkness fill the entire area}
		\item{Nearest Target turns into mist}
		\item{Illusion of a tree growing}
		\item{Nearest Target turns into a gnome}
		\item{Illusion of a portal to the Realm of Darkness and Fire}
		\item{Nearest Target sees a vision of the next encounter}
		\item{All nearby vegetation begins growing}
		\item{Nearest recently deceased target gets to reroll a Vitality Check at +5.  If none, nearest target gains $2D6+4$ FP and can store an additional 10 FP}
	\end{enumerate}
\end{tcolorbox}

\magicitem{Spider Arrows}{Transient Web}{Instant}{Pocket Spell}{2}{1}{0}{1}\label{spiderarrows}

Back when followers of Ohta still regularly went to wars, they would create these Spider Arrows to make sure their quarry couldn't flee.
Wherever the arrow lands, it casts a Web spell on the nearest target by turning the surrounding air into a sticky substance.
The target resists with Strength + Athletics, TN 10 or remains caught in the web.

\magicitem{The Raven Tree}{Realistic, Trans Species, Empathic Polymorph}{Instant}{Item}{2}{1}{22}

Anyone taking an apple from this tree instantly transforms into a bird of some kind, unless they are strong enough to overcome the spell.  The magic can rearrange up to 7 Attribute points in order to reduce their Strength to -5, and can add up to 7 points to their Speed to bring them high enough to fly.  Once the spell has completed, the target remains as a bird until sundown.

The tree regenerates 3 per scene.  Apples taken from the tree keep their magic for up to a week if kept in good condition.

\magicitem{Ring of Asphyxiation}{Torpor}{Instant}{Item}{2}{0}{1}\label{ring_asphyxiation}

This simple ring, inlaid with bone, act as per the first level of the Necromancy sphere to stop the user feeling the effects of cold or hunger while the ring is worn.

\magicitem{Venomous Seeds}{Plantform, Wind Blast}{Instant, Continuous}{Pocket Spell}{2}{0}{6}

This bag of seeds forms a one-use item.
Once opened the bag produces the effects of Wind Blast, moving everyone around back by 4 squares minus their Strength Bonus.
Effected creatures lose 4 initiative.

Once the seeds have laid, if they find any kind of purchase -- whether mud, soil or marsh -- they grow into thorny plans, full of venom, over the course of 4 rounds.
Anyone touching them sickens, gaining 6 Fatigue Points.

\subsection{Runic Items}

\magicitem{The Eternal Warrior's Armour}{Call the Dead}{Instant}{Pocket Spell}{1}{1}{1}

This plate armour, inscribed with necromantic runes, raises the wearer from the dead as a ghoul.  They're typically used by dwarvish suicide squads and worn only after making a solem pledge that the wearer will never stop fighting, even past death.

The breastplate contains two spells -- a second level spell to ensure the wearer rises from the dead, and a first level spell, ensuring the wearer never attacks another dwarf.

\subsection{Songs}

\magicitem{The Chocking Song}{Reading the Ripples}{Normal}{Item}{2}{4}{3}
The song of growth is sung by village elders in the Bearded Mountains regions.
The elders sing only on their own birthday, thanking Laiqu\"{e} for their village's prosperity.

The song is special as it sounds most pleasing when sung by someone devoted to Laiqu\"{e}, or a follower of the Code of Experience.
Conversely, those following the Code of Acquisition, or who follow Qualm\"{e} or Ohta, reliably choke and cough while attempting the song.

The song holds a total of 17 MP, and needs spend only 1 when sung.

\end{multicols}

\section{Random Starting Conditions}

\begin{multicols}{2}

For players who want some inspiration for their character's background, or just more randomness, and for GMs looking for a little inspiration for their NPCs, there is a table of Random Starting Conditions.  If you want to offer this to your players, tell them to roll 1D100 and read the result to the player.  Remind players that if they are going to roll then they should be ready to accept both good and bad results; this system is here to provide an interesting result, not a beneficial one.

If the player rolls twice and achieves two rerolls, they may find their parents died twice -- this is fine.  Grandparents, god parents and adopted parents all count as parents; other examples may have to be modified to retain consistency.  No more than three starting conditions should be applied, no matter how many times the player achieves a reroll.

	If two players obtain the same result, they share the same background, as per the Story, `It'll be just like the old days'.  If their parents were both killed in the same event then they are from the same village or perhaps siblings.  They might both have worked on a similar job or share a magical item which they found at the same time.

\setcounter{list}{0}
\begin{list}{\addtocounter{list}{1}\textbf{\arabic{list}:}}{\raggedleft}

	\item{ You have murdered a travelling alchemist in cold blood to steal his magical item but do not know its command word.  It is an alchemical wand, encrusted with expensive rubies, which holds 3 MP at maximum and regenerates 2 MP per scene.  Once the command word is uttered it emits a fireball up to 6 squares away, dealing $1D6+1$ Damage across 3 squares.   This costs 2 MP from the wand's store.  The GM secretly decides on a word in English which activates the wand.  If you ever state the word while within `earshot' of the wand, it activates.  Alternatively, you may find out the command word by simply trawling through all the words you know -- this requires an Intelligence  Academics task, TN 10, which can only be attempted once per downtime. }

	\item{ A prophecy from a high priest of Ohta decreed that you were fated for great things.  You might not follow this god -- you may not even approve of Her.  It matters not a bit to the church how you feel, all they demand is that you carry their sacred weapon.  The sword can cast any spell as per the fourth level of the Conjuration sphere.  If it is ever taken from you, the sword teleports back to your hand as soon as it has a sufficient amount of mana.  It holds 4 MP at maximum and regenerates 2 per scene.  It counts as having an Intelligence bonus of 4.  This sword also grants you a +4 bonus to all social tasks involving those who follow Ohta as it is famous among her scriptures.  You are required to give the sword a name, just as those who used it to fight against the encroaching nura during the nura wars each gave it a name.}

	\item{You are exceptionally well travelled and have a million tales to tell and friends you have met.  You gain an additional 3 Story Points but may only spend 2 at a time. }

	\item{ You are wanted for stealing from the mages' guild.  You snuck into the forbidden parts of \gls{college} and took an item.  The item is long gone, as is most of the money you might have collected for fencing it, but there is a permanent illusion of you sitting in the middle of the mages' \gls{college} so everyone who has ever been there knows your face.  The reward is currently up to 200 gp. }

	\item{Your identical twin sibling trained as a magic-user of some sort and opened a portal from somewhere in Fenestra.
Of course this is a criminal offence -- there are portraits of your face all over the guard stations in every city in the kingdom.
If you ever spend Story Points, you will be recognised at some point during that adventure.}

	\item{A childhood friend recently met you and had some stolen items to sell -- a quiver of five Spider Arrows (see page \pageref{spiderarrows}).
These holy items are worth a lot of gold, and she expects at least 100gp back from you.}

	\item{Your cousin is the head of the Invocation School in \gls{college}. Some mages may hate you for this, others will respect you, or at least pretend to in order to garner the good graces of the \gls{college}'s higher ups.}

	\item{You were once sold a sword by a gnome.  He says that it is also a key to a grand treasure, and his story, while vague, seemed very convincing.  The sword has several odd grooves and cuts around the base, but otherwise functions as a normal short sword. }

	\item{The elves of the Quennome forest count you as a friend after you saved one of their number from a woodspy.  They even have a song about you. You receive a +2 Bonus when interacting socially with the elves of that region.}

	\item{Your hometown had a bad-tempered, authoritarian captain of the guards. He killed your brother over an insult, but by the time you arrived back in town he had left as he was re-stationed.  The GM has 80 XP to create this character.  Additionally, his base FP is always equal to yours plus 5.  If you ever take your revenge, you receive 10 XP in addition to any for the assault.  You must land the killing blow in order to take this XP.}
	\item{After offending a priest of Laiqu\"{e}, the forest watches you, and waits to attack.  The next time you enter a wooded area, the GM rolls twice on the encounter table.}

	\item{Your mother died of plague shortly after your father was murdered by bandits.  Your family was split up, some being put into forced labour, others imprisoned and still others just went to roam the land to find what they could.  Your mission is to reunite your family.  Once every two adventures the GM rolls $1D6$ -- on the roll of a 1, you find out a sibling has been killed, on the roll of 2-4 you hear of a member of your family who must have your help.  On the roll of 5-6 you receive an additional Story Point and must immediately introduce a character who is a long, lost sibling.  In total, you have $2D6$ family members to find.  Each lost sibling you find grants you 3 XP.  If you find and successfully reunite all of those still living, you gain an additional 10 XP.}

	\item{You once fell into an ethereal portal to another land by accident, and wandered through strange lands for years.  You gain 2 Story Points that can only be spent in the ethereal lands.  Additionally, if you ever spend a Story Point on finding a mana lake, that mana lake contains a hidden portal to one of the ethereal lands.}

	\item{You have come a long way from your little home on the little islands around the Pebbles.  You start with the first level of the Seafaring Skill.}

	\item{You are a halfbreed.  Roll $1D6$ to determine your race randomly.  You gain the racial abilities of the first race listed.  Halfbreeds are infertile, like mules.}

		\setcounter{enc}{1}
		\begin{list}{\arabic{enc}\addtocounter{enc}{1}-\arabic{enc}\addtocounter{enc}{1}:}{}
			\item{Human/ Elf: You are a half-elf.  You have no racial penalties or modifiers.}

			\item{Dwarf/ Gnome: You are half gnome, half dwarf.  You gain 1 Intelligence but -1 Strength.}

			\item{Gnoll/ Human: You are half gnoll, half human.  You gain 1 Strength but -1 Charisma.}
		\end{list}

\item{As above, but gain the racial abilities of the second race, not the first.}

\item{You come from a cold, harsh mountainous region in the North of Whiteland.  You start with the first level of the Survival Skill for free, with a specialisation in Snow Tundra.}

\item{Your auntie was a planeswalker -- she travelled to distant, strange lands and brought back tales so fantastic that most thought she was mad, but you listened to each and every one of those tales of the strange lands outside of this world.  You gain the first level of the Ether Lore Skill for free.  With a successful Intelligence roll at a variable (but high) TN, the GM must give you basic information about any plane of existence you find yourself in.}

\item{You are born cursed with the taint of the nura.  Perhaps you were involved in a nura battle as a youth.  Perhaps your mother briefly changed into nura and then saved by a priest while she was pregnant with you.  You begin with the first level of the Saurecanta school for free and are on the Path of the Nura.  You may gain additional magic schools as normal.  You do not know what the different levels of Saurecanta do -- you can only find out by increasing this school with XP.  You can increase and know about any other magic sphere on the Path of the Nura as normal.  You may not gain XP for following a god at any point unless you make penance for your inherent sin and forsake all casting of nura magic for a year.  This addition does not come with MP, so you cannot cast spells unless you buy base MP or your Intelligence is already high enough to grant you free MP.}

\item{You grew up without parents, raised by your older sister on the streets of the city of Lakesend in Whiteland.  You gain the first level of the Larceny Skill for free.}

\item{Your uncle was a famous Academic, and travelled the land before returning to your people to tell them what strange lands lay about them.  You start with the first level of the Academics Skill for free.}

\item{You once, in jest, insulted a dwarven woman who was travelling through your homeland.  She could not break the law by ordering your execution on the spot, but has hired a dwarven assassin to find and kill you.  If that assassin dies, another will replace him.  This assassin is a standard dwarvish warrior with Stealth 2.  They will not stop coming until you are dead or you make a very outlandish apology to the dwarven lady in question.\footnote{See page \pageref{dwarven_soldier} for stats.}}

\item{You have some small speck of draconic blood in your veins as an ancestor mated with a dragon, polymorphed into a member of your race.  You begin play with the first level of the Invocation sphere and are initiated automatically onto the Path of Blood.  If you do not have any mana, you cannot use this spell.  You may instinctively learn any other sphere from the Path of Blood by spending the necessary XP.  Additionally, you gain DR 4 against all forms of fire.}

\item{You proudly crafted your own equipment before beginning the adventure.  You begin with the first level of the Craft Skill for free and select either a weapon, projectile or anything else worth 10 gp or less to begin with, for free.}

\item{You were mauled by a beast when younger and have never forgotten it. You have nightmares about the creature each night which leave you Fatigued.  Roll Charisma + Deceit against TN 7 each morning -- if you fail you begin with 4 Fatigue points.  You can stop the nightmares only by finding and killing the creature from your dreams in real life.  Roll randomly:}

	\setcounter{enc}{6}

	\begin{list}{\arabic{enc}\addtocounter{enc}{-1}}{}

		\item Dragon

		\item Basilisk

		\item Woodspy

		\item Umber Hulk

		\item Chitincrawler

		\item Griffin

	\end{list}

\item{One of your parents was lost in the Etherial realms beyond while exploring there. If ever you find them and bring them safely back home, you gain 15 XP.}

\item{Your home village was constantly assaulted by nura beasts.  Eventually it was entirely destroyed -- you managed to flee, one of the last survivors. During your time there, you were forced to learn how to shoot a bow to defend the place -- you begin with the first level of the Projectiles Skill for free.}

\item{You were born on Qualm\"{e}'s holy day.  The god of death will not touch you until your time.  Gain a +2 bonus to any Vitality Checks you make.}

\item{You were born during the height of Alassea, and that goddess has blessed you.  You begin with the first level in the Empathy Skill for free, reflecting your natural knack for charming people.}

\item{You were born at the height of C\'{a}lea and have always been one for picking up little factums here and there.  You begin with the first level of the Academics Skill for free.}

\item{Born during the holy day of V\'{e}r\"{e}, you are known to be true to your word and are known by all the priesthood of V\'{e}r\"{e}, who are always only too happy to help out where they can.  You can spend 1 Story Point every second story to introduce a priest of V\'{e}r\"{e}, as per the story `There is a man whom they call \ldots'.}

\item{Born during the festival of Laiqu\"{e}, you have always had a natural affinity with animals.  You begin with the first level of Beast Ken for free.}

\item{You were born during the festival of Ohta, while your entire village was outside in the public square having a series of brawls to settle old disputes.  You begin with the Brawler Knack.}

\item{You are among the last of the Whiteland nobles who were hunted almost to a man after threatening the king's castle wing which sits there.  You begin with the first level of the Deceit Skill as you are used to hiding your heritage.  There is a 40gp bounty on your head, so you had better keep your identity secret -- this will be easy enough around common people, but you might be hard pressed to tell a convincing lie to nobles, many of whom have a good working knowledge of accents and the precise customs of each land and the family and underlings of the landmasters there.  There are nobles in Whiteland among each race, so if you are not human, simply assume that you were the chief family in a village in Whiteland which was destroyed.}

\item{You served in a war some years ago against the South Kingdom.  You fought mostly at sea, but did not do terribly well.  You lost one hundred men due to your poor military decisions and were unceremoniously thrown out of the military.  You begin play with the first level of the Tactics Skill.}

\item{You grew up in a travelling circus -- perhaps as part of a freakshow, perhaps as a magical act.  You start with the first level of the Athletics Skill for free.}

\item{Your father was one of the most famed warriors in the land during the recent war with the Southern Kingdom.  You begin with the Knack: Adrenaline Surge which gives you 1 Damage whenever you remind yourself of your legacy.}

\item{Your home village was destroyed by goblins when you were younger.
You begin play with the Knack: Chosen Enemy (Goblins).}

\item{You grew up on the borderlands between gnolls and humans.  If you are a gnoll, you have the Knack: Chosen Enemy (Humans).  If you are human, gnolls are your chosen enemy.  If you are neither, choose one.}

\item{You were found as a small child and taken in by a kindly couple -- you have no idea if your parents were killed or you were simply abandoned. Roll on the racial chart to determine your parents' race.  Their race determines your background culture.}

\item{You were once captured and used as slave labour in the Southern Kingdom.  You have escaped but now have the Knack: Tireless.}

\item{Your parents were sentenced to death after a legal dispute with a member of another race -- roll randomly on the race chart.  You now have the Knack: Chosen Enemy (race).  If you roll your own race, roll again.}

\item{You were once on the run for a petty theft for over a month.  Almost everyone has now forgotten about it, but the memories of that time, and what you had to do to survive still haunt you.  You begin play with the Knack: Perfect Sneak Attack.}

\item{Everyone gets teased during childhood but yours was beyond severe.  The constant fights have taught you a thing or two however -- you begin play with the Knack: Fox Hop.}

\item{When you were only four years old, you got in a fight with the family dog over a scrap of meat.  You took a blade and killed the dog without a thought, then prized the meat from its still-warm jaws.  You do not recall this event but your mother recounted it every time you misbehaved, reminding you of how evil you are.  You begin play with the Knack: Finishing Blow.}

\item{The military unit you served with have all died -- the undead took most of them, the diseased wounds took others, and madness finished off the rest.  You alone remain, and you miss the days of rounding up simple criminals, giving them a little beating and then throwing them in the cells.  You begin play with the Knack: Expert Flanker.}

\item{You have been assaulted and robbed before.  Some people have had it worse, but you have never been able to let go of the incident.  Perhaps it's because of the scar across your face.  You begin play with the first level of the Vigilance Skill.}

\item{You were kidnapped by a necromancer when you were just a baby.  Your father managed to get you back but you have never been the same since.  Gain the Knack: Consume Soul, even if you have no MP to use it.}

\item{You have a mysterious Deck of Cards which you know to be magical.  Each time you draw one a spell flies out.  See page \pageref{deckofmanythings} for the item.}

\item{You are in love with someone completely inappropriate -- perhaps they are above your station, perhaps from a family who detests you.  If ever you manage you successfully marry this person, gain 20 XP.  The GM is obliged to make this challenging.  You can write letters to this person and if you ever find yourself in your homeland, you will automatically encounter this person.  You must flesh them out a little but the GM will create the exact Traits.}

\item{You begin as a blank slate -- total amnesia.
Assign your Skills as normal.
Each time you spend a Story Point, a memory resurfaces or someone from your past who knows you arrives.
You must slowly piece your past together from these fragments.
You can allow the GM to decide on a past and give the pieces to you if unsure of what it should be.
Leave 30 XP from your starting XP unspent.
You may spend this XP at any point to gain anything but an Attribute Bonus.
Any time you spend this XP, you remember an additional piece of your past.
You begin play with two pieces of equipment worth up to 40gp; only one of these can be weapons or armour.}

\item{You are old.  If you are an elf, you are 100 years old.  Apply the normal Attribute adjustments.  Otherwise, your maximum Body Attributes are decreased by 1 each but you begin with an additional 20XP.}

\item{You share a background with the player to your left.  If they own a magical item, you own it too (or there is some argument).  If they would be the last member of their people, you pair are the last of your people.  You can share the cost of any Stories you tell.}

\item{Your parents were wiped out when an arrogant alchemist strolled into town and refused to pay the inflated prices for nobles.  He incinerated them without a thought and then bribed the local guards to avoid punishment.  You begin play with the Knack: Chosen Enemy (Alchemists) and receive 10 XP if you ever track him down.}

\item{Runes have always fascinated you.  You begin play with the first level of the Academics Skill and a specialisation in Runelore.  You are on the Path of Runes and can purchase any sphere from it.}

\item{You used to work for the temple of V\'{e}r\"{e} and were turned into a human mana-conduit.  You now store 2MP, available only to priests of the temple of V\'{e}r\"{e}.}

\item{You were once part of a travelling band of bards, hired as an actor.  You had to put costumes together on the fly.  You begin with the first level of the Deceit Skill.}

\item{You once worked for a crazed mage named Adrian the Red who was banished from the mages' guild -- he never declared why.  During this time he gathered a terrifying army of nura.  He was eventually wiped out, and luckily for you, by this point you had already left.  Every second story you tell using your Story Points must involve either Nura, the undead or mad mages studying forbidden magics.  If you know a blacksmith, he might be a hermetical nura hobgoblin living alone in the forest and coming out to trade wares.  If you know a cut-throat, this might be an ogre who has managed to survive by convincing a village to constantly deliver meat to him.  If you know a powerful mage, this might be an undead priest who sits alone in a chapel, praying to Qualm\"{e} each day to release him from his torment.}

\item{You originally come from the Southern Kingdom.  You may not spend your original 5 Story Points while in Fenestra, only while in that other land.}

\item{You travelled round the South Kingdom for some time -- you start with 4 Story Points to be spent in the North Kingdom and 3 Story Points which can be spent in the South Kingdom.}

\item{Your father has been put in prison for murdering a nobleman, but you are sure he is innocent.  You receive 20 XP for getting him out, one way or another.}

\item{You were once travelling with a band of merchants when a horrifying forest creature attacked -- the GM rolls $1D6+11$ on the Forest Encounter chart of your local area to determine the exact creature\footnote{If the local area has no forest encounters, substitute the encounter table with the most results.}; you now have a crippling phobia of this creature (if the result is a mana lake, reroll).  If you encounter this creature, you must make a Wits check, TN 8, each round; failure indicates you must flee immediately and swiftly.  If you ever deal the killing blow to such a creature, the TN to overcome your phobia reduces by 1.}

\item{You once saved a priest of V\'{e}r\"{e} from being attacked by bandits.  He blessed you with the safety of the forest from that day onwards.  You store 1 MP like a mana stone and regenerate 1 MP per scene.  If you spend this MP, you can attempt to send a forest creature to sleep by saying a very short prayer to V\'{e}r\"{e}.  The TN is 4 plus the creature's Wits  Aggression.}

\item{Your little sister once stole a ring of invisibility from a priest of Alass\"{e}.  You thought it was quite funny until he took his vengeance, turning her into a cat.  He got his ring back and walked away, leaving you in shock.  The cat follows you everywhere.  It has Str -5, Spd 4, Dex 3, Int 1, Wts 1, Cha 1.  If you ever find a miracle worker who can undo the spell (TN 10) you gain 10 XP.  Until that time, you have a constant cat companion.}

\item{You were once married, settled, and happy.  Happy until a couple of years into your marriage it transpired that you are infertile.  Your partner left you and now you must carve out some kind of life as an adventurer.  This is not what you envisioned for your life at all!}

\item{You are in possession of a book you found on the mutilated body of a man on the road.  The bandits took all his goods but left the book out of disinterest.  It details extremely advanced alchemical instructions on how to summon creatures from other worlds.  This book is worth up to 200 gp to the mages' guild.  It can be used to learn the 5th level of the conjuration sphere through the Path of alchemy.}

\item{You were hired for a battle with a powerful priestess of Qualm\"{e} but were so terrified of her undead host that you fled before the battle.  While fleeing you accidentally encountered the undead host on the road and had to hide in the surrounding forest.  Then the host stopped, and the priestess left her wagon in a forest clearing before continuing towards the battle. After you were sure they were gone, you crept up to the wagon and stole the first thing you could see -- a painted skull, sitting on the table.  Since then you learnt two things -- that all of your companions are dead (and you have been presumed dead with them) and that the skull you stole comes from an old priest of Qualm\"{e}.  His spirit remains in the skull.  He cannot speak but can summon the dead back to life.  He has a base MP of 6, Intelligence 3, Wits 1, Necromancy level 5, Invocation level 3, Fate 2 and Aldaron 1.  He can affect someone with Necromancy level 4, allowing them Death Sight, at which point the target can see and hear the necromancer.  His current wish is to return to a priest of Qualm\"{e}.  He is not a cooperative magical item for others' use but a creature in his own right.  He has only a limited fear of death as returning to this land would be difficult but not impossible.}

\item{You once spent a month in cells being tortured before the guards realise they had the wrong person.  Begin with the Knack: Hardened.}

\item{There is a prophecy that you will somehow interfere with the will of Qualm\"{e} -- you have no idea what this is about but some cult or other out there are not happy with you.  Someone has sent a ghast to kill you.  It instinctively knows roughly where you are.  It follows you on foot. You keep moving, on horseback (begin play with a horse) whenever you can, every day.  You are safe only in populated areas, where it usually does not approach.  See page \pageref{ghast}.}

\item{You serve a minor deity from your home village and wish to spread word of their glory throughout the land.  You must select their name and personality.  You will work with the GM to create a list of pious endeavours which grant you XP.  If you prefer, this can be a unique Code.}

\item{You once destroyed a powerful artefact from some temple -- perhaps a sacred tree in a temple of Laiqu\"{e}, or a doll from a temple to Alasse. Whether this was on purpose or by accident, that god now has it in for you.  Every time you are outside of a city and obtain a result of `no encounter' the GM rerolls once in order to try and obtain an encounter. This continues until you somehow make up to the god (via the church) for your terrible transgression.}

\item{You are addicted to alcohol.  Any time you have a strong drink -- ale does not count -- you must make a Wits roll, TN 8; failure indicates you must have another drink.  Each drink increases your Fatigue and increase the TN by 1 and you must roll again for another, and another, until you prize yourself away or are too drunk to drink any more or you are forced to stop drinking by some other means.}

\item{You started your adventuring career with a heartfelt gift from your father -- the family warhammer.  Apparently it has been in the family for three generations, and now four.  You said you were not sure if the weapon fitted your exact adventuring style, but your father only told you that you would get better in time.}

\item{Your family are rich, and very fond of doting on you.  Begin with 50 gp and 100 gp worth of equipment.  This can include an entourage who will join you for the first adventure, and possibly a few more -- perhaps jesters, cooks, a few soldiers and maybe the services of an alchemist for a single session.}

\item{Your grandmother, high priestess of Laiqu\"{e}, died recently and made you promise to look after the family cat.  For those who follow the Path of Divinity (Laiqu\"{e}) it counts as a 5 point mana stone and can Polymorph people as per the 3rd level spell, Polymorph Other.  It has a range of 6 squares and uses your late grandmother's Intelligence bonus of 2.  The cat has Str -5, Spd 4, Dex 2, Int 0, Wts 2, Cha 3.  It cannot speak and the GM makes all decisions about when it will use this power according to the interests of the cat, not the interests of your character.}

\item{You have been captured for a crime you did not commit and were shoved naked into a cell.  You have managed to escape and become a stowaway on a travelling wagon to get where the adventure starts but you start with nothing -- not a sword, piece of bred, spellbook or loincloth.  You begin with an appropriate number of Fatigue Points if you start in a cold region.}

\item{ You are very young.  You begin play with only 20XP.  No Trait may be higher than 1, although FP and MP can be at any level.  You may raise no Trait to above 1 during play.  During your first downtime you enter a growth spurt and gain 40XP and can spend any saved XP. }

\item{ Your mother had to give birth unexpectedly while travelling.  The camp did not know it at the time but they had stopped by a mana lake.  Since then your eyes have had an unnatural shine to them, begin with 2 MP base -- this is a free level but the next level will cost 10 XP as usual. }

\item{ Do not read this aloud.  One of the player's pieces of starting equipment is cursed.  Random members of your family have died, previous adventuring companions have died.  All of your friends have died and you have sauntered through life taking all their good luck.  If you ever take a Vitality Check, you can request a 4 bonus before the roll.  If you do so, the next companion who takes a Vitality Check takes a -4 penalty.  All Vitality Checks are rolled by the GM in secret.  You must indicate whether you are taking the 4 bonus by handing a secret note to the GM. }

\item{ You have been captured by a dwarvish necromancer so that your dead body could be used as the vessel for an ancient warrior-wizard who is destined to return and destroy the King of the North while reclaiming all the land around the Bearded Mountains for the dwarves.  Unfortunately for them, two things went wrong -- the first is that you were not actually dead when the spell started (just close to death), the second is that other dwarves came to interrupt the illegal ritual and kill the necromancer.  They set you free, the spell completed but with a strange twist.  If ever you die, that powerful spirit shall return and inhabit your body with full access to powerful spells and begin amassing an army and increasing his own personal power by drinking the souls of the dead.  The dwarves explained to you that they thought this was a likely result but let you go because you were not guilty of any crime. }

\item{ You once served under your grandmother as a village healer, and later went to serve in the militia to fight in the Whiteland wars of humans and gnolls.  You begin play with the first level of the Medicine Skill. }

\item{ You once fought as a gladiator in the arena of a city.  You were quickly thrown out for your irritating ability to defeat opponents without drawing blood.  Begin with the Knack: Disarm. }

\item{ You once fought in a gladiatorial arena but managed to escape.  If ever you use a Story Point or journey to a large city, you are recognised, though not everyone will have an interest in telling local guards where you are - technically, escape is not illegal.  However, if your captors ever want to get you back they have the paperwork to show that kidnapping you would also not be illegal.  Begin with the first level of the Combat Skill. }

\item{ You once worked as a court jester -- you wandered different noble courts, even entertaining the king on one occasion.  People took you for an idiot, but during that time you learnt a lot.  You gain 1 Charisma and have a minimum Charisma score of 1.  This cannot take you above your racial maximum.  You also gain an additional Story you can tell -- if ever you meet a noble, you can spend one Story Point and say `Didn't I see you in his majesty's court that time when \ldots}

\item{You enlisted to fight with one of the most famous heroes of the land -- Elric Redthirst, paladin of Ohta, to fight against encroaching nura horrors.
You met with the opposing forces on a hilltop, chastised and slaughtered many of them, then chased the remaining host back into their cavernous lair and destroyed each of them to a man.
The forces were so equally matched that everyone died -- everyone except you.
You picked up Elric's sword -- a beautiful yet simple thing made from the finest steel, enchanted to destroy any creature.
It is known as The God Ruiner.
If the church of Ohta find out you have taken the sword, they will hunt you to the ends of the earth for it.
At present, nobody knows you well enough to be looking for you -- they probably assume that some nura survived and stole the artefact.
However, almost all followers of Ohta and many others know exactly what the sword looks like -- that distinctive purple colouring along the handle and the words of vengeance carved into the hilt are difficult not to remember.
The sword stores 5 MP.
If it ever inflicts 12 or more Damage on an opponent (before DR is applied), it can errupt with an invocation spell, inflicting an additional $2D6$ Damage which bypasses all DR, including Perfect DR.
It otherwise functions as a normal great sword.

\magicitem{God Ruiner}{Raging, Internal, Lightning Bolt}{Instant}{Artefact}{2}{0}{5}

}

\item{You have been hired by priests of C\'{a}l\"{e} to chronicle your adventures in the land.
You are largely expected to make your own way and earn your own money, but any time you are in a large city you may report to a temple of C\'{a}l\"{e} and receive 2 gp per month of adventuring time, or 6 gp per season.
They want all the information on the various secrets of the land that you can obtain for them -- hidden grottos, conspiracies, uncharted mana lakes, anything new and interesting.
}

\item{You once wished upon a shooting star during the cold festival at the height of Alassea.
You wished for all your dreams to come true.
Since then they have -- sort of.
Any time you sleep the GM secretly rolls $2D6$, on the result of 10 or more an illusion is created out of your dreams.
This uses any level of the illusion sphere, up to 5th, and counts as if it were cast by someone with Intelligence and Wits 5.
The GM crafts a dream out of elements and characters from the current adventure and from your background stories -- if you have already specified you have known a blacksmith from childhood and you are currently running away from nura creatures, that blacksmith might arrive after having been transformed into an ogre.
}

\item{Your grandmother was a village healer and taught you all you know about the craft. You later served in Whiteland during the war between gnolls and humans.  Begin play with the first level of the Medicine Skill.}

\item{You have been enchanted to serve as the personal body guard of a high priestess of V\'{e}r\"{e}.  However, you managed, in a drunken stupor, to leave long enough for the enchantment to dim a little.  Your mistress is still looking for you, and the more legend spreads of your deeds the more intent she will be on finding you.  Any time you spend a Story Point, your identity will be known by someone during that adventure.  If you ever meet your mistress again, you will not be able to refuse her -- you must obey her every command.  Mechanically, she has used the 5th level of the Enchantment sphere on you.}

\item{You were a high ranking member of a guild before deciding to pack it all in and adventure.  If you are not human then you have some equivalent expertise in your own culture.  You begin play with an additional $2D6\times4$ gp.  The guild type is determined by one of your highest Skills.}

\item{You played a fantastic game of dice, and won a magical scroll.
You just need to read the Quenya written and you can find out what it does.

\magicitem{Portal Scroll}{Unrestrained, Open Teleport}{Normal}{Pocket Spell}{3}{2}{5}

The scroll takes 5 rounds to speak properly, and opens a doorway to the Realm of Shifting Corridors.\footnote{See page \pageref{shiftingcorridors}.}
The portal remains open for 2 scenes.

}

\item{Your father gave you a map to some burried treasure, but wouldn't tell you how he got it, only that you are not the only one with a copy of the map.  Each time you spend a Story Point, roll a die.  On the roll of a `6' you run into someone chasing the same treasure as you.  At this point you can make an Intelligence + Survival check to understand how the map relates to the local landscape.  The GM will fill in the details.}

\item{If you roll \arabic{list} or above, reroll twice}

\end{list}

\end{multicols}

