\documentclass[10pt,twoside]{book}

\usepackage{config/bind}
\usepackage{config/booklet}
\setcounter{bookLevel}{2}
\input{glossary.tex}

\declareRegions{broch,village,lonelyRoad,lonelyTavern}

\begin{document}

\miniCover{Drops}{}

\bookletThreads{broch}

\bookletThreads{lonelyRoad}

\bookletThreads{village}

\chitincrawler

\segment[\gls{morning}]{lonelyRoad}% AREA
{Patience}% NAME
{\Pgls{crawler} observes patiently}% SUMMARY

The air is calm; the troupe receive only 2~\glspl{mp} this day.
\Pgls{crawler} observes the \glspl{pc} from a tall tree, over 200~\glspl{step} away.
If confronted, it flees.

\segment[\gls{night}]{village}% AREA
{Payment}% NAME
{\Pgls{crawler} attacks at night}% SUMMARY

Rain beats the ground, and the \gls{crawler} from earlier comes to visit the \gls{village}.
If the \glspl{pc} make any noise (or have any light) it attacks them.
Otherwise, it enters the \gls{village}, climbing the wall, and then begins to feel around the houses for an opening.
The \glspl{pc} may hear the sound of crying children over the rain, but they won't hear the \gls{crawler}.

\thread{What He Needs}

\chitincrawler
\setcounter{gold}{\value{xp}}

\segment[\gls{morning}]{broch}% AREA
{Carrying the Fallen}% NAME
{\Gls{ranger} Cainboak needs urgent medical care}% SUMMARY

The piper drones as the Sun rises slowly.
The muddy-orange light shines down on a crawling shape the size of a wagon, approaching the \gls{broch} at a gallop.

The \gls{broch} has \glspl{bow} for everyone's use, so the \glspl{pc} can aim from the balcony.
Once the \gls{crawler} approaches, the \glspl{pc} can aim at \tn[12], but need to hit \tn[17] for \pgls{vitalShot}.

If the \glspl{pc} kill the \gls{crawler}, \gls{jotter} Carndug values the corpse at \arabic{gold}~\glspl{sp}, takes
\roundUp{gold}
\arabic{gold}~\glspl{sp} for the \gls{templeOfBeasts}, divides
\set{gold}{xp}\divide\value{gold} by 2%
\addtocounter{gold}{0}%
\arabic{gold}~\glspl{sp} among the \glspl{pc}, then takes back the \glsentrylongpl{sp} to repay their debt.

Finally, he hands them a `town-pass', stating they may enter town for a day.
Their mission is to approach a nearby \gls{bothy}, where \gls{ranger} Cainboak lies on the ground, badly wounded.
They must take him back to town, unharmed.

The \gls{bothy} lies four miles away, and a cart waits for them outside (they will have to push it by hand).
Inside, Cainboak is clutches his sides.
\Pgls{basilisk} trampled over him, breaking an arm, a leg, and most of his ribs.
He won't be able to say much, or do anything.

\segment{lonelyRoad}% AREA
{Unfriendly Faces}% NAME
{\Glspl{guard} ahead carry a dangerous prisoner}% SUMMARY

\begin{exampletext}
  Yesterday, six \glspl{guard} attacked a bandit camp.
  Three survivors fled with their captured leader --- \pgls{witch} called Guilter.
\end{exampletext}

As the troupe trundle slowly onwards, another wagon joins them from another road.
The back has a man, shackled to the ground, his arms held painfully outstretched.
He stares at the \glspl{pc} with pain and malice.

The \glspl{pc} may wish to overtake, the wagon in front, but they cannot move fast without damagine Cainboak.
Or they may wish to hang back, but this will make them very late to get to town.

\humanalchemist[\NPC{\M\Hu}{Guilter}%
  {eyes burning with hatred}% DESCRIPTION
  {always mouthing words, mostly silent}% MANNERISM
  {freedom}% WANTS
  \npcQuote{Jesters, all of you}]

\showStdSpells

If the \glspl{pc} do nothing about the \gls{witch} staring at them, Guilter looks for the worst possible moments to cast spells on them.

\thread{What You Said}

\segment[\gls{night}]{broch}% AREA
{The Voiceless Roc}% NAME
{The \glsfmttext{jotter} needs someone to kill a local \glsfmttext{griffin}}% SUMMARY

Thunder threatens, as \glspl{guard} talk business in Rankvore \gls{broch}.
``\textit{The farmers say the \gls{griffin}'s as big as a house.
They say it picked up a sheep and just flew away.''}

After the pipes at dusk, \gls{jotter} \composeHumanName\ gives the \glspl{pc} the privilege of going to Tallbow \gls{village} to kill that \gls{griffin}, and bring it back, for a 50\% cut of the value of the corpse.

\griffin[\npc{\A}{Massive \Glsfmttext{griffin}}\setcounter{r2}{3}]

\segment{lonelyRoad}% AREA
{What Lies in the Forest}% NAME
{\Pgls{griffin} mimics a dog's bark from the forest}% SUMMARY

Passing the forest, the troupe hear endless barking noises, each one a perfect repetition of the last.
The noise comes from \pgls{griffin}, imitating the sound that it associates with sheep.

The \gls{griffin} soon goes silent, but the troupe can find the tree it stands on with a \roll{Intelligence}{Survival} roll at \tn[8].
If they talk or make noises as they approach, it imitates them perfectly, then jumps from the treetop and flies somewhere nearby, and repeats.

\segment[\gls{vlg}]{village}% AREA
{Competition}% NAME
{Other \glsfmtplural{guard} hack at the bushes, but wish they could kill the \gls{griffin}}% SUMMARY

As the skies clear, \glspl{pc} arrive in Tallbow \gls{village}, which has two giant \glspl{crossbow} along its massive wooden wall.
The giant \glspl{crossbow} have a -2~Penalty to aim, but inflict \dmg{4}~Damage.

Three \glspl{guard} \glspl{gDigger} --- \composeHumanName, \composeHumanName, and \composeHumanName\ --- have the unhappy duty of cutting down the bushes and trees around the farmland outside the walls, this gives the local archers more space to see danger coming towards people in the fields.
They hate the menial chore, and accuse the \glspl{pc} of nepotism.

\begin{itemize}\it
  \item
  Hey \gls{jotter}'s pet!
  How much you giving him to get the profitable missions?
  \item
  What's your rank?
  I'm \pgls{gDigger}, why can't I get this job?
\end{itemize}

If the \glspl{pc} agree to split the profits from the \gls{griffin}'s corpse, the \glspl{gDigger} help happily.
The farmers have sided with the \glspl{gDigger}, and ask the \glspl{pc} to sleep outside the walls.
However, they will help \gls{harvesting} the \gls{griffin}'s corpse once it's dead.

\humansoldier[\npc{\T[3]\F\M\M\Hu}{\Glsfmtplural{gDigger}}]

\segment[\gls{evening}]{village}% AREA
{Echoes}% NAME
{The \glsfmttext{griffin} mimics \pgls{pc}}% SUMMARY
The \gls{griffin} mimics \pgls{pc}, but won't attack until someone stands away from the rest (and outside the \gls{village}).
One they leave, the \gls{griffin} swoop-attacks, attempting to grab any character with \pgls{weight} under 6.
If it receives Damage, it's forced to land; once down it cannot fly without climbing a tree.

After one swoop the \gls{griffin} leaves for $1D6$ days.
The farmers of Tallbow suggest the \glspl{pc} go into the forest and hunt for it immediately.

The \glspl{gDigger} finish their work after three days and leave.


\end{document}
