\documentclass[10pt,twoside]{book}

\usepackage{config/bind}
\usepackage{config/booklet}
\setcounter{bookLevel}{2}
\input{glossary.tex}
\externalReferent{core}
\externalReferent{judgement}

\declareRegions{broch,bothy,lonelyRoad}

\assignHumanName[\fjot]

\begin{document}

\miniCover{A Short Guide}{}

\randomize
\Person{%
  \NPC{\N\F}% Symbol
    {Goblin}% Name
    {lanky, with mottled-marble skin}% Description
    {licks lips}% Mannerism
    {meaty food}% Wants
  }%
  {{-1}{1}{1}}% BODY
  {{-1}{2}{-3}}% MIND
  {%
    \set{Projectiles}{2}
    \set{Caving}{2}
    \set{Stealth}{1}
  }% SKILLS
  {}% KNACKS
  {}% EQUIPMENT
  {\ifodd\value{r3}\else\mutation{r4}\fi}% ABILITIES

\randomize
\Person{%
  \NPC{\N\M}% Symbol
    {Goblin}% Name
    {black veins on milk-white body}% Description
    {clicks tongue}% Mannerism
    {sugary food}% Wants
  }%
  {{0}{1}{0}}% BODY
  {{0}{-1}{-2}}% MIND
  {%
    \set{Melee}{1}
    \set{Brawl}{2}
    \set{Deceit}{1}
    \set{Stealth}{2}
  }% SKILLS
  {}% KNACKS
  {}% EQUIPMENT
  {\ifodd\value{r3}\else\mutation{r4}\fi}% ABILITIES

\vspace{-1.6em}
\noindent
\underline{Men}

{
  \setcounter{track}{2}
  \noindent
  \setlength\tabcolsep{1.5pt}
  \begin{tabular}{cc|cccc|cccc|cccc|}
    \showInterval{\thetrack}
    \Repeat{13}{%
      & \ifnum\value{track}>2\setcounter{track}{0}\else\stepcounter{track}\fi%
      \showInterval{\thetrack}%
    }
    \\
    \Square
    \Repeat{13}{& \Square}
    \\
    \Repeat{12}{ & }
    \tiny A & \\
  \end{tabular}
}

\noindent
\underline{Miles}

\noindent
\boxStat{20}

\vfill
{
  \footnotesize\sffamily
  \noindent
  With thanks to Dyson Logos for the maps \vpagerefrange{Dyson_Logos/winding_overview}{Dyson_Logos/winding_cave} (\texttt{CC-BY}).
}

\clearpage

\pagestyle{minizine}%

\begin{exampletext}
\noindent
Goblins have captured a dozen men.
\Pgls{ranger} has captured two goblins.
Now the troupe must take one as a guide, and rush across twenty miles, before the goblin horde eat all their prisoners.
\end{exampletext}

Check the rules on \gls{foraging} and \gls{navigation} before starting.
Mark off miles travelled and men eaten on the left.

\subsubsection{The \Glsfmttext{broch}}
stands on a tall hill at the \gls{edge}.
It keeps the local \glspl{village} safe from wandering \glspl{monster} by making noise to attract them, so  archers can shoot them down.
But today the \glspl{ranger} have all left, along with the last piper.
So \Gls{jotter}~\fjot\ gives the \gls{broch}'s pipes to the \glspl{pc}, but fails to mention that someone else should stand at the top with \pgls{bow}.

The roll is \roll{Strength}{Performance} at \tn[12] (to shout) or \tn[9] (if they can play the pipes).
On success, begin the next \gls{monster} encounter.
On a tie, the \gls{monster} arrives an hour after Sunset.

Whether or not the pipes succeed, anyone on the top can see a figure before dusk:

\begin{boxtext}
  On the high hill, on top of a tower, feels like being on top of a boat, watching a  sea of green in every direction.
  Small openings let you see patches of the \gls{lonelyRoad}, heading towards open farmland.
  The farmland huddles around a distant town.
  
  Not far below, a bailey's high walls stand defiantly against the \gls{edge}.
  And a man in black approaches the bailey, with two child-sized figures on a leash, both pale and naked.
\end{boxtext}

\paragraph{If the \glspl{pc} report this to the \gls{jotter},}
he sends them out immediately to investigate.
\fjot\ allows each \gls{pc} to take \rations, or a halberd with them, if they request either.

\paragraph{Otherwise,}
everyone in the \gls{village} stands on their rooves, waving their arms at the \gls{broch}.
If the \glspl{pc} don't respond to that, five men with spears approach the \gls{broch} to inform \Gls{jotter}~\fjot\ of the situation, and \fjot\ sends the \glspl{pc}.

\subsubsection{At the \Glsfmttext{village}}
\gls{ranger} Spittlespite lays groaning in bed, festering wound in his leg, loaded crossbow at the ready, and a couple of scared farmers with pitchforks watching over two goblins in the corner.

\begin{speechtext}
  A bloodied servant of the \gls{warden}'s son told us that their caravan has been ambushed by goblins, so naturally, we had to move immediately, or the warden would have our heads if he hears we hesitated\ldots
  Goblins ambushed the Warden's son and his entourage, and took them captive, then ate all their horses.
  We tracked them down, and after a bloody fight, I got away with two goblins on a leash.
  The rest died, or were taken.
  Dead either way\ldots

  But we killed a lot of those \gls{sylf}'s sucklings, too.
  No more than four goblins survived, and ushered bound prisoners in the direction of, what I believe, must be their lair.
  But those will grow hungry soon, and I fear you have no more than a couple of days before all those men are goblin feed.
\end{speechtext}

Pointing towards two goblins, he continues:

\begin{speechtext}
  Take one with you; two are too much trouble, as you can see by my leg.
  They seem to only understand blade and food, but that should be enough for them to take you to where others are.
\end{speechtext}

\subsubsection{Through the Woods}
the goblin's lair is 20 miles away, in a part of the woods seldom traversed.
Only the goblin knows where exactly, but cannot speak.

\needspace{5\baselineskip}%
Each \gls{interval} a number of things happen:

\begin{itemize}
  \item
  The troupe can march 4~miles through the dense woods.
  Marching another mile inflicts \pgls{ep} on everyone (including the goblin), and carrying the goblin inflicts \pgls{ep}.
  \item
  The goblin wants to sleep for 2 hours, which reduces travel by 1 mile.
  Failure to sleep inflicts 3~\glspl{ep}, as usual.
  Gently cradling the goblin while it sleeps requires a \roll{Wits}{Empathy} roll at \tn[8].
  Failure means it has not slept properly, and must try again.
  \item
  The goblin wants food.
  It must eat 3~\glspl{ration} per day or suffers 3~\glspl{ep} per missing meal.
  \item
  When the goblin sleeps and eats, it heals 3~\glspl{ep}.
  A full day of healthy-living means it heals a quarter of its missing \glspl{hp}.
  \item
  The goblin horde eat another prisoner.
\end{itemize}

The horde began eating last night, so go mark some men as dead on the opening page, then give each goblin 3~\glspl{ep}.

The goblin doesn't speak the same language as the \glspl{pc}, but it understands the offers of food, and threats of violence.
It will help the troupe with foraging, or spotting any danger, but it's only interested in food it can eat, and danger that threatens it directly.

It is malicious, cunning, and shameless.

\paragraph{If the goblin's hands are tied,}
it will be visibly upset, and will try to subtly lead the troupe in a wrong direction, moving one mile away from the lair each \gls{interval}.
\roll{Intelligence}{Survival} at \tn[10] to notice at the start of each \gls{interval}, or \glspl{pc} realize after two \glspl{interval}.
It might also occasionally trip on purpose, and then struggle to get up, making the troupe move one less mile each \gls{interval}.

\paragraph{If left unsupervised,}
it attempts to eat as much food as possible, and destroy any equipment.

\subsubsection{While the goblin is well fed}
(no \glspl{ep} Penalty, and has eaten enough)
then it leads the troupe to its home, where the captives lie.
But it also enacts mischief:

\paragraph{\Square~Attempts to steal from \pgls{pc}.}
It will prioritize whatever useful thing it can destroy, or secretly drop in a bush, or anything sharp or easily flammable it can use to cause harm.

\paragraph{\Square~Spots loose stitching}
on one of the \glspl{pc}' bags, then bites into it, tearing a hole.
All the items in it spill to the ground, and the bag becomes unusable.
The \gls{pc} without a bag can carry items only in its hands.

\paragraph{\Square~Throws a handful of its fresh shit on the campfire,}
then wipes its hand on the closest \gls{pc}.
If the campfire is not extinguished, the stench will prevent anyone nearby (except the goblin) from sleeping well that night, resulting in one extra \gls{ep} in the morning.

\paragraph{\Square~Leads the troupe through mud,}
then tries to trip the clumsiest \gls{pc}, who resists with \roll{Dexterity}{Athletics} against goblin's \roll{Dexterity}{Brawl}.
If it succeeds, it cackles at its victim, as they struggle in deep mud.

%\begin{dlist}
%  \item
%  Lazily drags its feet, and will fall asleep if left undisturbed.
%  The troupe moves two miles less this \gls{interval}, unless someone keeps hurrying the goblin, in which case, the penalty reduces to one mile.
%  \item
%  Needs to take a dump.
%  A big one.
%  No, it doesn't need to go to a bush, nor does it care if anyone's watching, it'll do it right here, right now.
%  In fact, there is a grin across its face as the most horrid smell violates the troupe's nostrils.
%  Everybody who stays nearby during the suspiciously lengthy act will stink for \pgls{interval}.
%  \item
%  Squeals while pointing at the \glspl{pc}~\glspl{ration} and its stomach.
%  Throws a tantrum if not fed.
%\end{dlist}

\boar

\paragraph{\Square~Finding boar tracks,}
it points towards them, and then towards troupe's \glspl{weapon}.
Going after the boar will send the troupe one mile away from their goal, and successfully tracking it down will require a \roll{Wits}{Survival} at \gls{tn} 12.

\paragraph{\Square~Falls down,}
and pretends it sprained an ankle, crying in pain.
If \pgls{pc} tries to examine it up close, it leaps and tries to bite them.
If the wound isn't treated with \roll{Intelligence}{Medicine} at \gls{tn}~10, an infection develops, and the \gls{pc} gains one \gls{ep} in the following \gls{interval} (and the treatment \gls{tn} rises by 2), and loses \pgls{hp} one \gls{interval} after that.

\paragraph{\Square~Shrugs comically,}
while looking about in pantomime confusion.
It points to a tree, and pretends to climb and then spot something.
Through these mimes, it tries to indicate that it has lost its way, and must climb a tree to get a better view.

If the \glspl{pc} let it climb a tree, it stands on the top, admiring the view, for the rest of the \gls{interval}, then tries to escape.
If \pgls{pc} climbs up after it, then the goblin shits on their face.

\subsubsection{While the goblin remains unsated}
(either gaining \glspl{ep} or has \pgls{ep} Penalty)
its antics become progressively nastier.

\paragraph{\Square~Feels spiteful,}
and leads the troupe through \pgls{thornyThicket}.
Being almost naked, it wants to be carried through the spiky growths, but if the \gls{pc} carrying it has any food, it will try to steal and immediately eat some.
It has +2 to this roll due to surprise and terrain.

% \begin{dlist}
%   \item
%   Picks up a fistful of dirt, and throws it in the face of the closest \gls{pc}, blinding them on a hit.
%   Then it tries to escape.
%   \item
%   Screeches loudly enough to attract an encounter.
% \end{dlist}

\paragraph{\Square~Drools,}
and hungrily stares at the weakest \gls{pc}, or the troupe's pack animal.

\paragraph{\Square~Jumps on a nearby \gls{pc}'s backpack,}
trying to steal and immediately eat food.
If the \glspl{pc} manage to tear it off the backpack before it succeeds, it will grab a random item instead, preferably something small, and fling it into woods.
\roll{Wits}{Vigilance} to find the item fast (\gls{tn}~10 with daylight, 12 with \pgls{torch}, and 15 with nothing), or the troupe will have to spend more time looking for it, moving one less mile this \gls{interval}.

\paragraph{\Square~When the troupe are about to eat,}
it spits on their food, making sure it's seen.
A big, nasty ball of spit, with a glob of something foul in it, that the goblin picked up and chewed along the way.
Is that earwax?
The goblin hopes no one will want to eat this now, so it can have it all.
Anyone eating this food has to roll \roll{Strength}{Survival} at \gls{tn}~12, or they develop a random disease%
\exRef{judgement}{Judgement}{diseases}
on the start of the next day.

\paragraph{\Square~Attacks \pgls{pc},}
but falls to the ground, tired, after \pgls{round}.

\paragraph{\Square~Goes delirious,}
bites off one of its fingers, and eats it and loses \pgls{hp}.

\paragraph{\Square~Delirium progresses,}
as it bites into its own tongue, and loses another \gls{hp}.

%\begin{dlist}
%  \item
%  Leads the troupe into a cave, but there is only a stash of dried nuts (worth one \gls{ration}) hidden under a rock there.
%  This moves the troupe one mile further from the lair.
%  \item
%  Hungrily screams during night, waking everyone up.
%  All \glspl{pc} will have one more \glspl{ep} in the morning, due to a sleepless night.
%  \item
%  It forages as it goes, and eats everything the moment it picks it up.
%  It rolls a \roll{Wits}{Survival} at \gls{tn}
%  \ifcase\thetemperature%
%    14%
%  \or%
%    12%
%  \else%
%    10%
%  \fi, and on success, will only find enough for one \gls{ration} during the whole \gls{interval}.
%  Stopping, and slightly deviating from the path all the time means that the troupe travels two miles less this \gls{interval}.
%\end{dlist}

%\begin{dlist}
%  \item
%  It leads the troupe to a griffin eating a dead deer.
%  The goblin will slobber and gesticulate towards the carcass, and if \glspl{pc} try to move on, it will raise a lot of noise, provoking the beast to charge at the troupe, protecting its catch.
%  This detour leads the troupe one mile away from the lair, and the remains of the deer provide two~\glspl{ration}.
%  \item
%  Becomes disoriented, leading the troupe two miles away from the lair.
%  \glspl{pc} can realize the direction makes no sense with \roll{Intelligence}{Survival} at \gls{tn}~10, but this only stops them from moving further away, and they make no progress forward this \gls{interval}.
%  \item
%  Faints, losing consciousness for \pgls{interval}.
%  \roll{Intelligence}{Medicine} at \gls{tn}~10 to wake it up, but unless fed, it still takes \pgls{interval} for it to be able to properly lead the troupe.
%  \item
%  Keeps stopping to check if anything is edible.
%  Digs through the dirt, rummages through bushes, chews grass, but finds no food.
%  The troupe moves one less mile this \gls{interval}.
%\end{dlist}

\clearpage
\subsection{At the Goblin Lair}
The entrance smells of burnt meat, and is covered with a tangle of stacked branches and leaves, masking it from the outside.
\sideMap[4]{Dyson_Logos/winding_overview}%
  {%
    \ref{lair:tunnel}/77/74,
    \ref{lair:right}/86/44,
    \ref{lair:left}/11/67,
    \ref{lair:lower}/32/15,
  }%
{Cavern side view.\par\null}

If the barrier is disturbed, a human skull, loosely hanging on the inside of it, rolls down the cave mouth, alerting the lone goblin in the lower right cave.
Carefully examining the branches reveals the skull, otherwise it's \roll{Wits}{Vigilance} at \tn[8] to notice, for the \gls{pc} who moves them.
Any loud noises will echo through the caves and alert all the goblins.

\mapPic[\large]{b}{Dyson_Logos/winding_cave}{
  Goblins!/54/58,
  \ref{lair:tunnel}/91/67,
  \ref{lair:right}/94/33,
  \ref{lair:left}/28/80,
  \ref{lair:lower}/04/12,
}

\mapentry[lair:tunnel]{Slipdown}
The muddy slope descends fast.
\roll{Dexterity}{Caving} at \tn[8] to avoid a fast slide, and falling 8~\glspl{step} down to \gls{area}~\ref{lair:right}.%
\exRef{core}{Core Rules}{falling}

\ifnum\thepage<14
  \goblin[\npc{\T[2]\N}{\arabic{noAppearing}~Goblins}]
\fi

\hobgoblin[\npc{\M\N}{Hobgoblin}]

\mapentry[lair:left]{Crack and Smoke}
A small fire dimly lights the floor littered with bones, one goblin eating badly cooked human remains, and another goblin, and a hobgoblin, asleep.
If the goblins hear the \glspl{pc} coming, they throw two dried \gls{griffin} wings on the fire.
Within 3~\glspl{round} they make a rancid smoke, which rises up the tunnel, inflicting 4~\glspl{ep} on anyone who breathes it in.%
\exRef{core}{Core Rules}{holdingBreath}

\mapentry[lair:right]{Gnawing Dark}
Four goblins gnawing on a leg bone, guarding the captives.
Climbing up the ceiling-hole to \gls{area}~\ref{lair:tunnel} requires a \roll{Speed}{Athletics} roll at \tn[12].

\goblin[\npc{\T[2]\N}{\arabic{noAppearing}~Goblins}]

\mapentry[lair:lower]{Gentle Weeping}
Human captives, if any remain at this time.
Their arms and legs are bound with their own torn clothes, and each has 2~\glspl{ep} for every day since they were abducted.

\ifnum\thepage<14
  \begin{speechtext}
    Help!
    My daddy is very rich, you must help!
  \end{speechtext}
\fi

Artix (the \gls{warden}'s son) has manipulated and sacrificed his entourage until the last man --- \composeHumanName.
But \currentName\ does not want to die.
So if people come to save Artix, then to live he must \emph{become} Artix.
So if only \currentName\ remains alive, he claims to be Artix; and if Artix also remains alive, both of them claim to be Artix, and \currentName\ mimicks every statement.

If the \glspl{pc} rescue \currentName, he will flee just before reaching town.

\end{document}
