\documentclass[10pt,twoside]{book}
\usepackage{config/bind}
\usepackage{config/booklet}
\setcounter{bookLevel}{2}
\input{glossary.tex}
\externalReferent{judgement}
\newcommand\carnBothy{}
\assignHumanName[\carnBothy]
\appto{\carnBothy}{~\glsentrytext{bothy}}

\declareRegions{broch,bothy,lonelyRoad}

\begin{document}

\miniCover{Granny's Soup}{\vfill\pic{Unknown/cottage}}

\bookletThreads{broch}

\bookletThreads{bothy}

\bookletThreads{lonelyRoad}

\begin{speechtext}

  Old Carnpox was old when I was young, and she's still old.
  \Pgls{witch} is one thing, but she's something else.
  She doesn't buy or sell.
  No need for civilization.
  She's dangerous.

  I sent \pgls{ranger} to deal with her, then two more.
  None of them returned.
  She must know that I sent them.
  She would never let them die without knowing, so she knows.

  So now I'm deep in debt.
  I hope this box works.
  It cost so much that it has to\ldots

\end{speechtext}

\segment[\gls{morning}]{broch}% AREA
{The Parcel}% NAME
{The troupe must deliver a box, unopened, to \pgls{hag}}% SUMMARY

\pagestyle{minizine}%

\Gls{susjot} figures that Granny Carnpox can detect murderous intent, so he finds the most clueless and unassuming people in the \gls{broch} to deliver the box.

\begin{speechtext}
  Got a special parcel to deliver.
  Simple mission, and the temple's happy to pay two \glsentrylongpl{sp} each.
  Just take this wooden box to old Carnpox, unopened.
  She lives just past \carnBothy.
  You understand about keeping the box shut, and delivering the box to her, still shut, and \emph{not} opened.
  Am I making sense?
\end{speechtext}

\Gls{susjot} quizzes the troupe about the mission, to make sure they understand how to not open a box.
If a \underline{player} says they will open the box, \gls{susjot} hears the \gls{pc}, and declares they're no longer on the mission
(the player must select another \gls{pc}).

\segment{bothy}% AREA
{Halfway House}% NAME
{The \glsfmttext{bothy} has no firewood}% SUMMARY

\carnBothy\ stands eight miles down the \gls{lonelyRoad}.
It has no firewood, just an axe.
Nobody else arrives, so the \glspl{pc} have a simple decision --- go cut wood (and make noise) in the night, or don't.
If they cut wood, they can restock the \gls{bothy}.
\ifnum\value{temperature}=0
  Without wood, they sleep in the cold, and take \pgls{ep} in the morning.
\fi

\needspace{4\baselineskip}%
\segment{lonelyRoad}% AREA
{Mission Report}% NAME
{Two \glspl{ranger} ask the troupe how their mission with the box went}% SUMMARY

\humansoldier[%
  \npc{\T[3]\Hu\M\F}{\composeHumanName, \composeHumanName, \& \composeHumanName}%
  \Mount{Horses}%
  {3}% Str
  {0}% Dex
  {1}% Spd
  {2}% Wts
  {1}% Athletics
  {Vigilance 2}
]

After the troupe leave Granny Carnpox's place, they meet three \glspl{ranger} on the road, riding horses.
The \glspl{ranger} ask them how the mission went.
If the troupe have opened the box, the \glspl{ranger} sigh as they deduce \gls{susjot}'s plan.
The \glspl{ranger} should --- by law --- take them to town for prosecution.
%\exRef{judgement}{Judgement}{pitOfJustice}
However, if the \glspl{pc} agree to return to \carnBothy\ with Carnpox' head, the \glspl{ranger} will wait for a day.

If the troupe attack, the \glspl{ranger} turn their horses and flee.
The horses spend \pgls{ap} to turn, and one more to move.

\subsection{The Way to Granny's House}

\vspace{-1em}
\begin{boxtext}
  A mile from the \gls{bothy}, a narrow foot-path leads through the forest.
  A man's croaky voice comes from the shadows of the forest, asking \textit{`what's in the box?'}.
\end{boxtext}

\Person{%
  \NPC{\footnotesize\T[2](\E)\M\Hu\tiny\glsentrysymbol{sylf}}% Symbols
  {Rentfold \& Baxter}% Name
  {chad jawline, glorious moustache, limbless}% DESCRIPTION
  {raised eyebrow}% MANNERISM
  {to find a sane home}% WANTS
  \npcQuote{Baxter, stand!}%
  }%
  {{3}{2}{1}}% BODY
  {{0}{-1}{0}}% MIND
  {%
    \set{Brawl}{2}
    \set{Survival}{2}
    \set{Academics}{1}
    \set{Stealth}{5}
  }% SKILLS
  {}% KNACKS
  {}% EQUIPMENT
  {\tentacles}% ABILITIES

The voice is Rentfold,%
\footnote{Rentfold worked as \pgls{ranger}, but when he tried to kill Granny Carnpox, chef Gorflake cut his legs open, then cut them off, then the arms.}
who appears to be alone.
However, Rentfold has a symbiotic relationship with a beakless \gls{woodspy} called `Baxter'.
Granny cut out Baxter's beak to enhance her \glspl{spell}, but kept the \gls{woodspy} alive.
Now it goes everywhere with Rentfold, lifting him up as they travel.
Baxter lies nearby, camouflaged perfectly into the forest.

\begin{speechtext}
  How did you get here?

  \textbf{Magic.
  What's in the box?
  Are you here for the soup?}

  What soup?
  Who did your moustache?

  \textbf{My elf-girlfriend.
  Who sent you?}
\end{speechtext}

\togglefalse{examplecharacter}

\Person{%
  \NPC{\F\Hu}{Coinfilch\tiny\glsentrysymbol{sylf}}
  {body-hair covers everything but scalp and mouth}% DESCRIPTION
  {looks about for mushrooms}% MANNERISM
  {to find enough mushrooms to complete the soup}% WANTS
  \npcQuote{The key to soup is mushrooms, and soup is the key to everything}}%
  {{1}{1}{1}}% BODY
  {{-4}{0}{1}}% MIND
  {%
    \set{Melee}{2}
    \set{Cultivation}{1}
    \set{Projectiles}{1}
    \set{Survival}{1}
    \set{Academics}{2}
    \set{Stealth}{1}
    \set{Larceny}{1}
    \rapier
    \renewcommand\weaponName{rapier (hidden in fur)}
    \set{xpbonus}{3}
  }% SKILLS
  {}% KNACKS
  {satchel of mushrooms}% EQUIPMENT
  {}% ABILITIES

\paragraph{Three miles from the cottage}
the troupe meet a hairy creature, prowling for mushrooms.%
\footnote{Coinfilch and her crew came to kill Granny in her sleep.
Granny inverted their hair, then left them to catch fire while moving through her cottage's cramped kitchen.}
Her naked, perfectly human mouth speaks normally, but the rest of the face has hair all over it.
She has no clothes -- just thick, brown and white hair covering her whole body, except where humans normally have hair.
Coinfilch asks the troupe if they have come for the evening soup.

\paragraph{Two miles from the cottage}
a man approaches with bone-quills instead of his hands.
They drip `ink', which comes from \pgls{crawler} spinneret, embedded in his abdomen.

\Person{%
  \NPC{\M\Hu\tiny\glsentrysymbol{sylf}}%
  {Nettlepike}% Name
  {oily beard covered in red and blue ink-stains, dark-blue ink dripping from quill-bones}% DESCRIPTION
  {wipes pen clean}% MANNERISM
  {to be useful}% WANTS
  \npcQuote{the pen is mightier than the sword, you know?
  So I'm actually stronger than when I had hands}}%
  {{1}{0}{0}}% BODY
  {{-4}{-1}{-2}}% MIND
  {%
    \set{Brawl}{2}
    \set{Cultivation}{1}
    \set{Projectiles}{1}
    \set{Survival}{1}
    \set{Academics}{2}
    \set{Stealth}{1}
    \set{Larceny}{1}
  }% SKILLS
  {}% KNACKS
  {}% EQUIPMENT
  {\addAbility{%
    Quill-bones for hand, which produce translucent ink}%
    \stepcounter{weaponDamage}%
  }% ABILITIES

\begin{speechtext}
  What's in the box?
  A delivery?
  I'm the official letter writer here, so I'm basically in charge of Granny's post.
  Yea, that looks like postage.
  Open it up, let me see!
\end{speechtext}

Nettlepike happily explains that he's a bad man because he tried to stab Granny.
He can tell them what happened to Coinfilch and Rentfold, if the \glspl{pc} ask.
However, he can't think very clearly, as he relates every sentence to letters he once wrote, and has a lot of trouble speaking about anything not concerned with writing.

\paragraph{One mile from the cottage}
the forest opens, revealing the cottage, and Carnpox' garden.

\begin{boxtext}
  The forest ends suddenly in
  \ifnum\thetemperature=0%
    a grand circle of flat snow.%
  \else%
    dead, grassless, earth.
    You stand in the outer ring, full of berries, carrots, and a peach tree which literally drips with sweet juice as the fat peaches have split open.%
  \fi%
  \footnote{Granny grows plants with \gls{witchcraft}, which decreases the soil's fertility until it becomes completely barren.
  Her garden forms an expanding ring around the infertile centre.}

  One mile in the distance, a wattle cottage with a smoking chimney.
\end{boxtext}

\paragraph{Half a mile from the cottage}
Granny Carnpox spies the crew, and casts \textit{Hospitable Inquiry}, which detects if someone has murderous intent.
And if they do, she rallies the troupes, secretively, and quietly, while leaving through the back door.
Otherwise, she tells Gorflake to prepare for guests.

\paragraph{At the cottage}
Rentfold sits on the roof (his symbiotic \gls{woodspy} lies camouflaged into the wattle beside him).
He points (by pursing his lips) to the weapons rack by the gnome-sized door; it holds a single, rusted, greatsword.
He tells the \glspl{pc} to place their weapons upon it.
He does not reveal how he got on the roof.

\subsubsection{Inside the cottage}
chef Gorflake cooks for granny.%
\footnote{Gorflake once tried to poison Granny, so she's `teaching him how to cook'.}
He has a special procedure to turn around without his fat stomach throwing everything off the table, and another procedure to get to the shelves on the other side of the room (it begins with putting his stomach on the table, and pulling it inwards, then shuffling round the side of the room).
Gorflake cannot leave the cottage through the front door.

\Person{%
  \NPC{\M\Hu\glsentrysymbol{abderian}}% Symbols
  {Gorflake}% Name
  {sweating and flustered}% DESCRIPTION
  {wipes running nose on sleeve}% MANNERISM
  {to make everyone eat}% WANTS
  \npcQuote{could you move aside!?  I need the big spoon!}}%
  {{3}{2}{-2}}% BODY
  {{-4}{-3}{-2}}% MIND
  {%
    \set{Brawl}{2}
    \set{Cultivation}{3}
    \set{Survival}{1}
    \set{Academics}{2}
    \set{Larceny}{1}
  }% SKILLS
  {}% KNACKS
  {}% EQUIPMENT
  {}% ABILITIES

The shelves above chef Gorflake hold flour, oils, rice, and beer.
The shelves at the other side of the room hold two clay pots of Water \glspl{boon}, three jars of dried \gls{basilisk} skin, and ink-dye suppositories for Nettlepike's pen-arms.

\Person{%
  \NPC{\M\Hu\tiny\glsentrysymbol{yonder}}% Symbols
  {The Whittler}% Name
  {gangly man with rust-coloured beard}% DESCRIPTION
  {chip-chip-chip-chip}% MANNERISM
  {to whittle, and remember his name}% WANTS
  \npcQuote{}}%
  {{1}{3}{0}}% BODY
  {{-4}{-3}{-2}}% MIND
  {%
    \set{Brawl}{1}
    \set{Melee}{1}
    \set{Crafts}{3}
    \set{Cultivation}{1}
    \set{Deceit}{1}
    \Dagger
  }% SKILLS
  {\specialist{whittling bowls}, \weaponmaster}% KNACKS
  {sandpaper}% EQUIPMENT
  {}% ABILITIES

By the fire, `the Whittler' whittles.
He can't remember his own name, or what he did wrong, he just knows he has to make a complete set of bowls and cutlery for Granny Carnpox.
If he completes this task, the enchantment wears off, and he will leave.
But every time someone has soup, they throw the bowl and spoon into the fire.

Over the fire, a massive pot simmers with eternal soup.
\Gls{chef}~Carnpox adds ingredients, people drink, he adds some more.
It changes a little bit every day.

Movement inside the room demands a \roll{Dexterity}{Athletics} check with \pgls{tn} equal to the character's \glspl{hp}.
Every movement deserves a claustrophobic description, and every second person entering the room adds +1 to the \gls{tn}.
Failure will knock something off a shelf, inflicting \dmg{-2}~Damage.

The next room has straw for sleeping, and a door to Granny Carnpox' room, where she sleeps with her cats.
This end-room has a barred exit-door.

\Person{%
  \NPC{\F\Hu}{Granny Carnpox}
  {blue headscarf, sunken eyes}% DESCRIPTION
  {touches people's cheeks}% MANNERISM
  {to care for the cottage}% WANTS
  \npcQuote{Death is such an awful waste}}%
  {{-1}{1}{-2}}% BODY
  {{2}{0}{2}}% MIND
  {%
    \set{Brawl}{1}
    \set{Academics}{2}
    \set{Crafts}{1}
    \set{Cultivation}{3}
    \set{Deceit}{1}
    \set{Empathy}{2}
    \set{Seafaring}{2}
    \set{Fate}{2}
    \set{Water}{2}
    \set{Earth}{2}
    \Dagger
  }% SKILLS
  {\specialist{growing peaches}}% KNACKS
  {Water \glsentrytext{boon}}% EQUIPMENT
  {}% ABILITIES

\showStdSpells[%
  \stepcounter{Water}
]

\subsubsection{At Sundown}
everyone arrives to eat the soup.
Coinfilch leaves hair everywhere.
Nettlepike complains that the wrong food will make his ink-gland produce weird colours.
Rentfold chews up chunks of meat, and spits them into Baxter's beakless maw.
The Whittler eats, then cries quietly as everyone burns their bowls after the meal.
Two cats arrive and meow loudly outside, but Granny won't let them eat until everyone else has finished.
These are Granny's `special little boys', modified with \glspl{spell} to grow big and strong.

\morphcat[\npc{\T[2]\A\R}{Cats: Gorgeous \& Rainfall}]

\act[Characters working together in \pgls{bandAct}]{finish the soup}{Strength}{Athletics}{10}
Once done, Coinfilch and Gorflake are free of their oaths; their Mind \gls{attribute} Penalties vanish and they may help the troupe.

\ifnum\thepage<14
  \subsubsection{Defeated Characters}
  become \glspl{npc}.
  \Pgls{cycle} later, the new \glspl{pc} may hear \pgls{doula} telling children the chilling tale of `the curious \gls{guard}', or `the man who kept his hat on, while accepting hospitality'.
\fi

\subsection{Inside the Box}
\label{insideTheBox}

\act{open the box}{Intelligence}{Seafaring}{10}
Inside, \pgls{crossbow} waits, loaded.
The steel bow stays rigid, dwarven beard makes an unbreakable string.
The white arrowheads gleam, all three made from quartz which shatters on impact.

The arrows inflicts \dmg{6}~Damage (helmets have a $\frac{2}{\dicef{6}}$ chance to negate all Damage).
Once the bolts fly, a scroll unfurls across the box's open lid:

\begin{speechtext}
  This device was manufactured by Nanpa.
  By operating this device, you agree that Nanpa remains \gls{temple}-neutral, and is not responsible for the use, misuse, or consequences of the device.
  If you would like to purchase a similar device for the return address, or any other address, send a detailed letter to Nanpa via the \gls{paperGuild} in \gls{ftown}.
\end{speechtext}

\act{reset the box}{Intelligence}{Crafts}{12}

\end{document}
