\documentclass[10pt,twoside]{book}

\usepackage{config/bind}
\usepackage{config/booklet}
\setcounter{bookLevel}{2}
\input{glossary.tex}

\declareRegions{broch,village,lonelyRoad,lonelyTavern}

\begin{document}

\miniCover{Spy in the Tavern}{}

Maps by Dyson Logos (CC-BY-SA 4.0)

\bookletThreads{lonelyRoad}

\bookletThreads{lonelyTavern}

\clearpage
\pagestyle{minizine}%

\thread{Where You Last Saw It}

\segment[\gls{morning}]{lonelyRoad}% AREA
{Weaver's Warnings}% NAME
{\Pgls{woetide} travels to confront Greetfall about his debt}% SUMMARY

\Gls{woetide} travels with a crew of traders
%(selling ropes and \glspl{ration})
to settle a dispute between the \Gls{tender}~Greetfall and \gls{proctor}~Shrewlame.

\begin{speechtext}
  Greetfall thinks that if he's far from town, he's far from the law.
  He thinks he can lie about anything he wants, and now he's contesting borrowing 100~\glsentrylongpl{gp}.
\end{speechtext}

\segment[\gls{afternoon}]{lonelyTavern}% AREA
{Ripe Tales}% NAME
{\Gls{tender} Greetfall is missing, his faceless corpse is lodged down the chimney}% SUMMARY

Stoopcraw and Gaulbloat are a pair of little shits, barely seventeen years old.
They love archery, but hate all other work.
So when they shot \pgls{woodspy}, they sold it as the nearest \gls{lonelyTavern} for 12~\glspl{sp} and started drinking.

Soon after, the tavern's \gls{tender} died.
It happened like this:

\mapPic[\large]{t}{Dyson_Logos/lonely_tavern_0}{
  Vegetable Garden/24/98,
  Four horses/75/17,
  $\leftarrow$To \ref{lt:guard}/58/53,
  Latrines/45/42,
  Well/88/45,
  \rotatebox{90}{Coaches}/88/86,
  Pantry/50/64,
  \ref{lt:fire}/48/77,
  Kitchen/57/72,
}

\begin{enumerate}\itemsep0em
  \item
  Greetfall hid his illegal \glspl{ingredient} above the stables, where traders can pick them up easily, and without him technically handing them anything.
  \label{lt:ingStore}
  If anyone had asked why he didn't pay the \gls{templeOfBeasts} 50\% of the sale as tax, he would have explained that he had in fact been robbed\ldots by someone who also left a ridiculously generous tip.
\end{enumerate}

\sideMap[5]{Dyson_Logos/lonely_tavern_stable_1}%
  {%
    \rotatebox{90}{Horse feed}/85/50,%
    \ref{lt:ingStore}/15/75,
  }%
{Above the stables.\par}
Currently, the room has two \gls{crawler} spinnerets and three \gls{woodspy} beaks encased in small barrels of alegar.
The door's complex string-mechanism can be opened without a trace with an \roll{Intelligence}{Larceny} roll at \tn[9].
Or someone can just cut the mechanism with a thin knife.

\begin{enumerate}\itemsep0em\stepcounter{enumi}
  \item
  Greetfall has managed to save up \lootBig, which he keeps hidden under the floorboards in his room (\roll{Intelligence}{Empathy} at \tn[10] to locate).
  \label{lt:tenderRoom}
  \item
  \Gls{server}~Gobvore sleeps next to the \gls{harvesting} room, but insists on knowing nothing about it (room \vref{lt:cutting}).
  \label{lt:server}
  Even thinking about illegal activity may induce a panic-attack.

  The room has a dozen teas and fancy pillows.
  One has \lootMedium\ hidden underneath.
  \item
  \label{lt:cutting}
  After Stoopcraw and Gaulbloat sold the \gls{woodspy} corpse to Greetfall, they helped him carry it up to the \gls{harvesting} room.
  They left to drink as prepared candles and knives for dissection.
  \setcounter{diceNo}{\value{enumi}}
\end{enumerate}

\sideMap[4]{Dyson_Logos/lonely_tavern_2_b}%
{%
  \ref{lt:chimney}/20/50,
  \ref{lt:guestFire}/38/50,
  \ref{lt:cutting}/38/82,
  \ref{lt:server}/76/82,
  \ref{lt:tenderRoom}/42/22,
  \ref{lt:singer}/07/05,
}%
{}
%{Creaking stairs and smell of copper.}

However, the \gls{woodspy} was not dead, merely wounded, then dazed.
Once Stoopcraw and Gaulbloat ran downstairs to drink, the \gls{woodspy} latched onto Greetfall's head, and chewed through the front of his skull.

The entire floor is covered in blood.
Two lanterns still burn, but the corpse has gone.

\mapPic[\large]{t}{Dyson_Logos/lonely_tavern_1}{
  To ground/60/05,
  \ref{lt:proctor}/70/30,
  \ref{lt:woodspy}/86/80,
  \ref{lt:farmer}/70/80,
  \ref{lt:mixer}/83/30,
  \ref{lt:parlour}/49/52,
  \ref{lt:chimney}/60/52,
  \ref{lt:guard}/20/30,
}%{Parlour and \glsentrytext{guard} rooms}

\begin{enumerate}\itemsep0em\setcounter{enumi}{\value{diceNo}}
  \item
  Stoopcraw sings on the roof, happy to have made some money, selling \pgls{woodspy} body.
  \label{lt:singer}
  \item
  The \gls{woodspy} found a guest room with a chimney but the grate at the top of the chimney blocked the \gls{woodspy}'s exit, so it went down.
  \label{lt:guestFire}
  \item
  But the fire at the bottom forced it out the chute before reaching the bottom.

  The \gls{woodspy} pulled its food down with it, but Greetfall was too fat to fit, leaving him stuck between floors.
  \label{lt:chimney}
  \item
  Sliding across the parlour, covered in soot and cinders, it fled into the first open door it saw\ldots
  \label{lt:parlour}
  \item
  \ldots and hid inside a chest, pulling a blanket over itself before camouflaging its skin to match the texture and colour of the linens.
  \label{lt:woodspy}
  \item
  \Gls{proctor} Shrewlame has arrived to settle how much Greetfall owes to the \gls{templeOfHate}.
  \label{lt:proctor}
  He and three other \glspl{armourer} took a single room, and wait for \gls{woetide} to arrive.
  \setcounter{diceNo}{\value{enumi}}
\end{enumerate}

\begin{boxtext}
  A tuneless song assaults your ears, theorizing about how \glspl{woodspy} mate.
  The singer stands on \pgls{lonelyTavern}'s roof.
  As the rest of the building comes into view, it shows more wall, but not a single window in either building.
\end{boxtext}

Once the troupe arrive, they find an overworked \gls{server}, but no proprietor.
Everyone starts shouting at \gls{server}~Gobvore to get some some ale, but she can't keep up while working alone.
\Gls{tender}~Greetfall told her that he is not to be disturbed under any circumstances (he wanted solitude and secrecy while \gls{harvesting} the \gls{woodspy} corpse) so she will try to stop the \glspl{pc} from contacting him, and won't go up.

\begin{speechtext}
  %\Gls{tender} Greetfall?
  \Gls{tender} Greetfall is not to be disturbed.
  I'll get you an ale, just a moment.
  Two moments.
  Maybe five\ldots
\end{speechtext}

\humanmaid[\npc{\F\Hu}{Gobvore}\toggletrue{allyCharacter}]

\begin{enumerate}\itemsep0em\setcounter{enumi}{\value{diceNo}}
  \item
  The fire has a cauldron overhead, but it has nasty chunks of black soot inside, dislodged from when the \gls{woodspy} clamoured down the chimney.
  \label{lt:fire}
  \item
  \Gls{mixer}~\composeHumanName\ has come to purchase \glspl{ingredient} for the \gls{templeOfSickness}.
  \label{lt:mixer}
  \item
  \composeHumanName\ and his two sons have come from Tallbow \gls{village} to sell their vegetables in town (they own a horse in the stables)
  \label{lt:farmer}
  \item
  An elf named Quetill\"e is trying to convince everyone that money is a `human construct' so she shouldn't have to pay any money just to sleep on the floor.
  \label{lt:elf}
  \item
  Once the \glspl{pc} arrive, Gobvore will show them to the \gls{guard} quarters, up the stairs, and separate from the other guests.
  \label{lt:guard}
\end{enumerate}

Gobvore can lead the \glspl{pc} to their quarters (room~\ref{lt:guard} \vpageref{Dyson_Logos/lonely_tavern_2_a} and \ref{Dyson_Logos/lonely_tavern_1}), before she finds the soup is burning in the cauldron.
The patrons complain bitterly about the black lumps in the soup (see area~\vref{lt:fire}).

\sideMap[4]{Dyson_Logos/lonely_tavern_2_a}%
  {%
    \ref{lt:guard}/25/70,
    \ref{lt:guard}/77/70,
  }%
  {}

Gaulbloat tells everyone in the tavern how much he's drunk, but he keeps losing count half-way through the `story'.
He has no intention of giving the \gls{templeOfBeasts} their 50\% cut of the money from selling the \gls{woodspy} corpse.
If \pgls{pc} pushes the matter, he argues, tells them to `take it up with \Gls{tender} Greetfall', threatens to kill them, vomits on them, then falls over.

Upstairs the parlour slowly fills with smoke, as Greetfall's corpse has blocked the chimney (\vref{lt:parlour}).
The \glspl{pc} may find the problem and pull him down, or if they enter room~\vref{lt:guestFire} and notice his feet sticking out of the fireplace.
Or they may notice nothing, since every room in the \gls{lonelyTavern} lies in shadow, and the little tallow candles people wander with give just enough light to see your own feet.

Noticing the charcoal marks around the parlour, or the feint blood-spatters leading into room~\vref{lt:guestFire}, should not be assumed.
The \gls{gm} must provide a clear image of the darkness, and the few things the \glspl{pc} \emph{can} see, and the players should ask for details.

If players want to abstract this with a roll, they can roll \roll{Intelligence}{Vigilance} to plan a proper search.
\Pgls{bandAct} requires splitting up.

\woetide

At \tn[6] they find the pool of blood in room~\vref{lt:cutting}.
At \tn[10] they find the corpse in the chimney.
And at \tn[12], the \gls{woodspy} in room~\vref{lt:woodspy}, inside an empty chest of old linens.

\paragraph{If Greetfall's corpse falls out,}
\Gls{woetide} accuses Shrewlame of murder, since Greetfall denied owing the money to the \gls{templeOfHate}.

Shrewlame says that Greetfall has killed someone, removed their face to confuse everyone, and run away into the forest.
He demands that the troupe go and search around the forest immediately.

Gobvore wants everyone to check on the horses but won't go outside after dark.

\humansoldier[\NPC{\glsentrysymbol{wrecan}\M\Hu}{\Glsentrytext{proctor}~Shrewlame}%
  {beard like a badger's arse}% DESCRIPTION
  {shoulders up to ears}% MANNERISM
  {get his loan back from Greetfall}% WANTS
  ]% QUOTE

The \gls{woodspy} remains, resting quietly, until night, when everyone hides in their own rooms.%
\footnote{At this point players may slow down to talk about the events.  Don't let them!  Go round the table clockwise and hand a `situation' to each player.  \Gls{woetide} speaks to one, another notices the room getting smokey.  On the second turn, the other patrons retire, and leave the \glspl{pc} to sleep.}

\segment[\gls{night}]{lonelyTavern}% AREA
{Creaking Stairs}% NAME
{The first is a human, the next \pgls{monster}}% SUMMARY

Once the tavern dies down, and everyone sleeps, \gls{woetide} goes down the creaking stairs to the latrines outside.
Any \gls{pc} on \pgls{vigil} will hear her, otherwise they can roll \roll{Wits}{Vigilance} at \tn[12].

\setcounter{wounds}{3}
\woodspy

Soon after, the \gls{woodspy} comes down, and climbs up the stairs then \emph{up} onto the roof (where Gaulbloat sang).
If nobody woke Gaulbloat, the \gls{woodspy} eats him, leaving a pool of blood, then enters the well (noted \vpageref{Dyson_Logos/lonely_tavern_0}).

Tracking the \gls{woodspy} requires a \roll{Wits}{Survival} roll at \tn[10].
Meanwhile, the \gls{lonelyTavern} fills with accusations and theories.
\Gls{woetide} wants to get statements from everyone, and tells Shrewlame he can't take the money he's owed from a dead man.
Shrewlame tries to search for any valuables he can soon after, to ``take back what's owed him''.

\segment[\gls{morning}]{lonelyTavern}% AREA
{Well Again}% NAME
{After a little time healing, the \gls{woodspy} has returned}% SUMMARY

The \gls{woodspy} crawls a little up the well, grabs Gobvore, and retreats back down, leaving no clue except the fact that the bucket is missing.

Soon after, the \gls{lonelyTavern} gets a new \gls{tender}, who buys a lot of guard dogs.
The \gls{woodspy} eats all of them.
This cycle continues until someone finds and kills the \gls{woodspy}.

\null

\thread{What You Want}

\segment[\gls{morning}]{lonelyRoad}% AREA
{Great Deals}% NAME
{Traders arrive with \glspl{ration} for sale}% SUMMARY

A caravan of three wagons trundles through the misty morning.
Two \glspl{soldier} --- \composeHumanName\ and \composeHumanName\ --- remain alive.
\Pgls{woodspy} killed the third when he took a shit without checking the area.
They tell the \glspl{pc} `\textit{Don't be like \composeHumanName.  Look before relief!}'.

\begin{description}
  \item[\composeHumanName]
  has \glspl{torch} (\mkPrice[cp]{7}), and speaks with endless positivity.
  \item[\composeHumanName]
  has \rations, \rations, and \rations\ -- \mkPrice[cp]{10} each.
  He disapproves of everything, and everyone.
  \item[Cindercough]
  adjusts her hat until the \glspl{soldier} are distracted and talks to \pgls{pc} alone.
  She wants to know if the troupe have found anything `\emph{interesting}', that they haven't shared with the \glspl{jotter}.
  She can get a good price\ldots
\end{description}

\segment[\gls{afternoon}]{lonelyRoad}% AREA
{Greater Deals}% NAME
{Cindercough returns}% SUMMARY

% A lightning storm builds.
% By evening, it will break, and each \gls{round} has a 1 in 6 chance of lightning providing light for the \glspl{pc} to see by.
% The energy provides 4~\glspl{mp} throughout the \gls{interval}.

A caravan approaches, with stone (for cenotaphs); barrels of milled flour, elderflower cordial, alegar (from the \gls{templeOfPoison}, and cheese; purple dyes (made of \gls{woodspy}), and brass pots.
And of course, some items the \glspl{pc} may want:

\begin{description}
  \item[\composeHumanName]
  has eight longswords for \mkPrice{10} each.
  \item[\composeHumanName]
  has \rations, \rations, and \rations\ -- \mkPrice[cp]{10} each.
  \item[Cindercough]
  Cindercough returns to ask if the troupe have anything to sell.
  ``\textit{Maybe the spinneret from \pgls{crawler}?  Or some \gls{griffin} wings?}''.

  She can pay \mkPrice{8} for any useable \glspl{ingredient}.
\end{description}

Four \glspl{gDigger} from the \gls{templeOfBeasts} stand watch with \glspl{bow}.

\end{document}
