\documentclass[10pt,twoside]{book}

\usepackage{config/bind}
\usepackage{config/booklet}
\setcounter{bookLevel}{1}

\input{glossary.tex}

\begin{document}

\miniCover{Bino}%
  {
    \vfill
    A Choose-Your-Own-Adventure Zine
  }

{
  \toggletrue{allyCharacter}
  \small
  \shortsword
  \noindent
  Grab a pencil, and a fistful of dice.
  Your \gls{pc} sits above.

  If the \gls{pc} takes Damage, mark off that many \glspl{fp}.
  If you run out of \glspl{fp}, mark off \glspl{hp}.
  When you run out of \glspl{hp}, the character dies.

  The \weaponName\ grants a \absNum{weaponBonus}~Bonus to hit, and \absNum{weaponDamage}~Damage.

  \randomCYOAcharacter{\gls{bow}, quarrel}

  \begin{description}
    \item[\Glspl{fp}]
    \fpStat
    \item[\Glspl{hp}]
    \hpStat
    \item[Cause of Death I:]
    \bigLine
    \item[Cause of Death II:]
    \bigLine
    \item[Cause of Death III:]
    \bigLine
  \end{description}

}

\pagebreak
\pagestyle{minizine}

\mapentry[dayOne]{A pregnant woman leans on her spear}
while rubbing her own lower back.
A man stands by her side, with a massive hood covering his head, and a wooden staff in his hands.
Behind them a funny-looking young thing slouches; last night everyone asked if they were human or elf, but they just kept sneering `\texit{what does it matter?}'.
Now finally, you see the \gls{jotter} climb the ladder to give everyone the first mission.

\begin{speechtext}
  Welcome to the \gls{templeOfBeasts}, where everyone dies a hero!
  Today's glorious (and profitable) quest involves \pgls{basilisk} whom we call `Bino', because she's white, and because `Al' would make a ridiculous name for \pgls{basilisk}.

  Bino has been spotted lying by the crossroads at the other side of the bog, where she likes to eat traders and their horses.
  You should be able to cross the bog before nightfall if you hurry, so get moving!
  Just remember to avoid that \gls{witch} --- she's trouble.
\end{speechtext}

Once out the door, the hooded man says he thinks everyone should take the road round the bog, while the pregnant woman insists on going `the proper way'.

\begin{selectPath}
  \willYe{Do as you're told,}%
  {go through the bog.}%
  {toBog}

  \willYe{Take the main road.}%
  {}%
  {toRoad}
\end{selectPath}
