\section{Town Encounters}

\epigraph{A criminal is a person with predatory instincts who has not sufficient capital to form a corporation. Most government is by the rich for the rich.
Government comprises a large part of the organized injustice in any society, ancient or modern.
Civil government, insofar as it is instituted for the security of property, is in reality instituted for the defence of the rich against the poor, and for the defence of those who have property against those who have none.}%
{Adam Smith}

\index{Encounters!Town}

\renewcommand{\sqarea}{Town}

\setcounter{encnum}{1}

\begin{multicols}{2}

\sidequest{The Trouble with Ale}
\label{troubleAle}

\startcontents[sq]

\sqminitoc

\noindent
The Ale Guild of the Shale takes dwarvish ale to \gls{town}, then waters it down, and makes an absolute killing in profits.
The local Ale Guild aren't happy, and shenanigans ensue.

\Gls{alemaster}, dresses his men as Knights\footnote{The term ``Knight'', is used by townsfolk for anyone in armour.
Guilds who travel need a lot of protection, as the roads are dangerous, and they general count as `knights'.}
from the faraway Ale Guild of the Shale, and has them cause trouble in order to give his rivals a bad name.
In retaliation, \gls{beardedalemaster}, dignitary to the Shale Ale Guild, destroys much of the local ale stores.

When not getting into trouble, \Gls{beardedalemaster} walks with the Wolf Heads adventuring party.
See page \pageref{beardedalemaster} for more on them.

\sqpart{Town}% AREA
{The Drunken Knight}% NAME
{A knight of the local Ale Guild starts fights while pretending to be from the Ale Guild of the Shale}% SUMMARY

This encounter can easily slip into the background -- the knight has no interest in assaulting the players, especially if they look well-armed.

\begin{boxtext}

	From the street ahead a drunken man in full plate armour falls against a wall, before getting back up and shouting at passers by that they're pathetic for drinking elvish nonsense, and need a proper dwarvish drink.

\end{boxtext}

\ldots except, he's not a knight.
He's a member of the local Ale Guild, and a patron of the temple of Alass\"{e}.
He's acting drunk and showing a symbol of the Shale in order to give the rival sellers a bad reputation.

Players may notice he's pretending to be drunk with a Wits + Medicine Group Roll, or notice that his coat or arms isn't painted onto his breastplate quite right with a Wits + Academics Roll.
Both have TN 10.

\NPC{\M}{Chris, Town Ale Guild Member}{Angry}{Pointing at people}{Alass\"e}

\person{2}% STRENGTH
	{0}% DEXTERITY
	{1}% SPEED
	{{0}% INTELLIGENCE
	{-1}% WITS
	{+1}}% CHARISMA
	{0}% DR
	{1}% COMBAT
	{Athletics~1, Deceit~1}% SKILLS
	{\longsword, \completeplate}% EQUIPMENT
	{\addtocounter{xpbonus}{-3}}

If questioned, Chris says this is ``just a little joke''.

\sqpart{Town}% AREA
{Ale Explosion}% NAME
{Barrels of Ale explode down the street}% SUMMARY

Play this encounter at the same time as the next on the list.

\begin{boxtext}

	Suddenly the house to your right explodes, and water rushes down the street, knocking aside stalls, women, two stray dogs, and sweeping the lot down the road.  It foams and froths like mad, and then you notice -- this isn't water, it's ale!

\end{boxtext}

The party can avoid being swept down the street by the storm with a Wits + Athletics roll, TN 8.

\paragraph{Background:} The local Ale Guild found that the more ale is brewed in a batch, the cheaper it becomes.
They bought a few cheap houses around the city, and build massive barrels to house the ale in.
However, \gls{beardedalemaster} found where they're stored, and paid a few street-rat children to come in and mess the place up, pulling corks out of the barrels, and such.
Unfortunately, with just a little too much jostling and jumping on the barrels, the unstable equipment exploded.

\paragraph{If the PCs investigate,}
an Intelligence + Vigilance Team Roll, TN 10, will allow the characters to find the bodies of the children who started the problem, along with the fact that they were recently speaking with \gls{beardedalemaster}, who currently drinks in \gls{pig}.%
\iftoggle{core}%
{\footnote{See the core rules, page \pageref{teamwork}, for Team Rolls.}}%
{}%

\sqpart{Town}% AREA
{You Can't Drink Here}% NAME
{People jeer at envoys from the Shale Ale Guild and a fight breaks out}% SUMMARY

\begin{boxtext}

	A crowd is vomiting out of a pub, pushing six men out with them\ldots or rather three men and three dwarves, all heavily armed.
	The crowd shout for them to go back to the mountains and that they're not welcome in the area, but a dwarf, covered in metal from head to toe, and wearing the emblem of the Ale Guild from the Shale, starts argues back at them, listing a litany of laws concerning public houses and the rights of foreign Guild Captains, with deafening volume.

	The crowd just jeer, making a large vein in the dwarf's head bulge while his face turns redder by the moment.

\end{boxtext}

A party of three men and three dwarves from the Shale are tired from a long day's walk, and need a rest and drink, but the townsfolk think of the Ale Guild from the Shale as nothing but troublemakers.

Laxen, the angry dwarvish Guild Captain, has had enough of the bickering and feels tough in his full plate armour.
He has no idea that \gls{beardedalemaster} has been pulling stunts, such as destroying the stores of the local Ale Guild.

\paragraph{If the characters do nothing,}
the crowd at one point pelt Laxen with some old soup, and local ale.
He swings his axe, and a fight breaks out instantly.
The group then flee the city, and all ale trade with the Shale stops.

\paragraph{If the characters try to de-escalate the problem,}
give them an appropriate roll.
They have time to try multiple tactics, from asking the crowd to calm down, to threatening to call the guard.
However, the crowd are very angry, so the default TN for the encounter is 10.

\dwarvensoldier[\NPC{\M}{Guild Captain Laxen}{Dour}{Strokes beard}{Experience}]

\humansoldier[\npc{\T\M\F}{Joshua, Rachael and Rob}]

These two are heavy drinkers, and heavy fighters, but they will stop as soon as Captain Laxen gives the word.

\dwarventrader[\npc{\T\M}{Mugin \& Thunin}]

\Gls{beardedalemaster} still drinks at \gls{pig}.  

\paragraph{Conclusion:} If the characters can track down \gls{beardedalemaster} and have her talk with \gls{alemaster} they might be able to make peace between the two groups.
In this case, crisis is averted, and the final part can be discarded.

\beardedalemaster

\sqpart{Town}% AREA
{\N Dry}% NAME
{Guild fights have caused all ale supplies to run out}% SUMMARY

\begin{boxtext}

	Later that day, you can find no rest in a tavern.  The first has run out of ale, and the second.  Asking around, people are saying they've all dried up, and now serve nothing but foul water.  The town are becoming agitated.

\end{boxtext}

If the party did not manage to avert total disaster in the previous encounter, play this encounter together with the next Side Quest.

For the next two encounters, all social rolls have a -2 penalty, as everyone around is feeling irritable.

Meanwhile, townsfolk are drinking more and more unpurified water, all of which has been tainted by the sickness of the nura who live in the sewers.  Raise the local nura rating by 1.

There is no possibility to repair this damage once it's done, and no option to repair the situation any time soon.

\stopcontents[sq]



\resumecontents[Villages]
\sidequest{The Lizardite Amulet}\label{lizardite}
\index{Lizardite Amulet}
\stopcontents[Villages]

\startcontents[sq]

\sqminitoc

\noindent
The lizardite amulet summons any item, and is exceptionally powerful as it has 30 MP in total.  It functions as per the third level conjuration sphere.  Unfortunately, it cannot be used unless the user first uses the activation word, and secondly, speaks in Gnomish.\footnote{The amulet was made by a gnomish alchemist, though the college don't like to talk about this fact.}

\magicitem{The Lizardite Amulet}{Any}{Alchemy}{Continuous}{Greater Item}{5}{30}

\par\noindent
The amulet was loaned from \gls{college} to \gls{king} for his entertainment, and now has to return.

The party begin by chasing after it, lose it, hear rumours about it, and eventually find it in a random village.

\sqpart{Town}% AREA
{The Shell Game}% NAME
{The characters must take a fake magical amulet to \glsentrytext{nightguard} so it can be swapped for the real one}% SUMMARY

The local Woodspy Bandits have heard that the amulet is returning to the Shattered Castle, and plan to attack the amulet bearers and steal it for themselves.
\Gls{college} have already heard that news of the item's location has got out, and are aware of bandits coming to take the item.

Chief alchemist Jared, of \gls{town}'s citadel, has arranged for three replicas to be created, and has transported them to the town.%
\footnote{See page \pageref{citadel_alchemist} for Jared.}
The guards plan to pass them off as the real item, and have them delivered to the real amulet bearers, with a message:

\begin{boxtext}

	The enemy will attack.
	Take this fake item.
	Bury the real one, and give up the fake.

	-- Jared

\end{boxtext}

The only problem is that nobody knows which road \gls{nightguard} who carry the Lizardite Amulet will take, so Jared is sending any able-bodied men he can in all directions, each with their own copy of the amulet.

\Gls{captain} is going around giving speeches to quickly recruit people for a mission.

If the characters are part of \gls{nightguard}, they must take the mission.
If not, they will be instantly deputized, and have to take the mission (but can expect financial compensation).

\begin{boxtext}

	You see \gls{captain} talking to a group of \gls{nightguard} nearby.

	\begin{quotation}

		\ldots and at that point, hand the amulet to them, but do not read the note.

		Journey immediately to the crossroads east.
		Once you find them, hand over the amulet and the note, and leave.

	\end{quotation}

	Once the captain has finished his speech, he hands over a green amulet to the group of men wearing chain armour -- apparently members of \gls{nightguard}.
	He then turns, spots you, and approaches.

	\begin{quotation}

		Men, the king has appointed you to carry out an immediate mission.
		This Alchemical Amulet must be delivered immediately to a group of \gls{nightguard} coming up from the South.
		Journey South, find them at the crossroads, and hand the amulet to them, but do not read the note.

		Once you find them, hand over the amulet and the note, and leave.

	\end{quotation}

\end{boxtext}

At this point the characters can accept or refuse the mission.
If they accept, they will be loaned horses.
Whichever they choose, play another encounter in the villages as soon as they leave \gls{town}.

\paragraph{The Amulet Bearers} are five brave members of \gls{nightguard} who have ridden for a long time.
But today, they've decided to go off-road, and rest.
The smoke from their campfire still wanders up into the sky, but they're not moving -- just resting.
They have no idea how to use the Lizardite Amulet.

\humansoldier[\npc{\T}{Five members of \glsentrytext{nightguard}}]

\paragraph{The Bandits} only want to grab the item, and head back to the Green Tower \footnote{See page \pageref{green_tower} for more on the Green Tower.}.
The bandits want to sneak up on the amulet bearers at the crossroad, unaware that the amulet bearing Guards are in fact camping nearby.

Most of the bandits are ex-soldiers.
They get their information from a traitorous member of \gls{nightguard} -- Darren.%
\footnote{See page \pageref{traitor} for \glsentrytext{traitor}.}

\humansoldier[\npc{\T}{Ten of the Woodspy Bandits}]

\woodspyleader

\Gls{woodspyleader} is a massive man, unable to run well, but he rarely needs to.
If things turn sour, he will flee at the first sign of trouble, taking his men with him.

\paragraph{The Crossroads} have a little tree coverage a stone's throw from the road -- just the place to ambush people coming from the South.

\paragraph{How this plays out} depends upon the players.
They might go towards \gls{nightguard}, in which case they have a moment to speak, and then the bandits attack.

Alternatively, they may just wait at the crossroads, in which case the bandits see them, then surround them quietly.
If the party aren't wearing \gls{nightguard} uniform, they bandits assume they are nothing to do with \gls{nightguard} and wander off quietly\ldots assuming the characters don't spot the abortive ambush and make trouble.

\sqpart{Town}% AREA
{The Book of War}% NAME
{\Glsentrytext{townmaster} hires a local thief to steal a book to learn of the amulet's command word}% SUMMARY

\Gls{townmaster} hired Derek, a local thief, to steal the Book of Ancient Wars, detailing various battles about the time of time of Rex Dalius.\footnote{See page \pageref{h_dalius} for more.}
The book comes from the restricted works in the Temple of C\'{a}l\"{e} (i.e. the local library).
Among the various things the book talks about, it mentions the alchemical amulet which conjures things, and describes the word which allows the user to activate its powers.
This won't be obvious as the book is very long, and very tedious.

What \emph{is} obvious, is the sprinting thief.

\begin{boxtext}

	A man sprints towards you at lightning speed, carrying a book, then darts to the side and down the street.

\end{boxtext}

Players who say they want to stop him before he rushes past should make a Wits + Empathy check at TN 9.
As soon as he's gone, three of \gls{town} guard round the corner in pursuit, shouting at the characters to get him.

The thief attempts to run to \gls{pig}, then disappear in the lively crowd, and into Jane's room.\footnote{See page \pageref{priestessjane}.}

\humanthief[\npc{\M}{Derek}]

If questioned at any point, \gls{townmaster} will have plenty of excuses, and in the end be given a library fine of 3cp.

\sqpart{Villages}% AREA
{Rumours of Magic}% NAME
{Local villagers think they have figured out the command word for the amulet}% SUMMARY

Whether or not the book survived, one way or another, the item's magical activation word has been discovered\ldots or so the villagers think.
Specifically, they think the command word ``Saur\"{e}'' will activate the item.
Everyone's talking about what they would do if they found the ``magical wishing amulet''.
Most people don't understand its powers, and think they can just wish to be the Rex, or to be young.

An Intelligence + Academics Group Roll at TN 7 lets the party know that this is nonsense.%
\iftoggle{core}%
{\footnote{See the core rules, page \pageref{grouproll}, for Group Rolls.}}%
{}

Play the next Side Quest from the Villages immediately.

\sqpart{Town}% AREA
{The Mob}% NAME
{A fake amulet is dropped, and the townsfolk all clamour after it}% SUMMARY

\begin{boxtext}

	A trumpet at the city gate sounds off, and \gls{nightguard} can be seen standing upon the great town wall, with three bleeding criminals kneeling at their sides, and three severed heads in their hands.
The town crier shouts out.

\end{boxtext}

The item resurfaces in town, as a guard drops it, and all the townsfolk are ready to trample each other to death for it.

\begin{boxtext}
	\begin{quotation}
		Hear ye, all!

		Our brave, and skilful warriors of \gls{nightguard} have gone out to the depths of the forest, and pulled out some of the local bandits who have so terrorized the poor men and women in these lands, with theft and murder.

		They have lost one of their number, and we mourn the loss of \ldots how the hell do I say this?

		We mourn the loss of `Stanisele', who died fighting for \gls{king}.

		But we also celebrate, as the famous lizardite amulet, property of \gls{college}, was rested from the hands of the bandits.

	\end{quotation}

	One of the night guard holds the amulet aloft in great, gloved hands, but then fumbles, and slips.  The amulet falls many yards smoothly to the ground.  The crowd hear the tiniest `clink', and then a young girl runs forward, shouting that she wants to wish for shoes.  The crowd begin to move, \emph{fast}.

\end{boxtext}

Getting through the crowd requires a Strength + Empathy roll, TN 10.
Town guards nearby also want to use the item, so they pull out their swords and prepare to cut down anyone stealing ``\glsentrytext{king}'s property''.
If the players don't intervene, the scene turns into a bloodbath within a couple of rounds.

Once finally obtained, an Intelligence + Academics Group Roll at TN 10 shows that this is not the real amulet.%
\iftoggle{core}%
{\footnote{See the core rules, page \pageref{grouproll}, for Group Rolls.}}%
{}%

\sqpart{Town}% AREA
{They Took Our Jobs!}% NAME
{Novice cutthroats assault the characters, but they have no idea what they're doing.}% SUMMARY

\begin{speechtext}

	Don't try any funny business.  Just hand over the gold. Or silver.  Or at least definitely those boots -- god I'd kill for proper boots.  I mean, I actually will.  I'll kill you.  Hand them over.

\end{speechtext}

Pick a character who's most likely to go somewhere alone.  If that's not feasible, the party are ambushed when sleeping in a tavern.
However it goes down, they're being robbed by some wannabe-thieves.

A quick Wits + Empathy check, TN 8, reveals the thieves haven't a clue what they're doing.
Once the non-fight has ended, the characters might ask what these three think they were doing.
They reply that they once worked for a Villagemaster, but Clarisa was accused of stealing tax money.
Robert was set to guard the money after the next round of tax collection, but in the morning, it was all gone.
The next week, all the staff were fired, as their replacements had arrived.

They claim that they never stole a penny, and have no idea how the money vanished.

The truth of the matter is that the taxes in the local village were being paid with false coin.
One woman, using the Lizardite Amulet, summoned false coinage, which disappeared soon after.

\humanmaid[\npc{\M\F}{Clarisa and Robert}]

\humanthief[\npc{\M}{Steven}]

\sqpart{Villages}% AREA
{The Old Lady}% NAME
{An old lady has found the amulet, and used it to hand out magical jewellery to local serfs}% SUMMARY

\begin{boxtext}
	A grubby little girl dances down the road with bear feet, occasionally singing, then stopping.

		``Do you like my jewels?'', she asks.

	She has one massive green rock of immense value, hanging from a copper chain, and four bracelets on each arm, each studded with crystals.
\end{boxtext}

The actual Lizardite Amulet taken by a loyal group of \gls{nightguard}, but soon after they were killed by ghouls created by \gls{necromancer}.
The ghoul carrying the amulet wandered away from the group, and since \gls{necromancer} had no idea about the amulet, it was just ignored.
Finally, that ghoul ended up attacking the villages, where Martha killed it and took the item.
After a visit from a gnomish alchemist, she learned the item's command word, and a few gnomish words for things to summon.\footnote
{\textit{Gold coins, ledge, boots, gold-necklace, dry sticks, satchel, fancy bracelet, pestle, morter, horseshoe, bandage, water, frying pan, coat, hat} and \textit{bowl.}}

Today, she summoned jewels to entertain Emily, a little girl.
Emily wants to keep the source of her jewels a secret, and the characters have zero chance of bribing her.
After a day, the jewels vanish into thin air.

\NPC{\F}{Martha the Healer}{Slow}{``Um\ldots''}{Acquisition}

\person{-1}% STRENGTH
	{0}% DEXTERITY
	{-3}% SPEED
	{{1}% INTELLIGENCE
	{0}% WITS
	{1}}% CHARISMA
	{0}% DR
	{0}% COMBAT
	{Academics 1, Empathy 2, Medicine 3}% SKILLS
	{Lizardite amulet, bandages}% EQUIPMENT
	{}

\sqpart{Villages}% AREA
{Martha Returns}% NAME
{The old woman who has the item helps the characters}% SUMMARY

If the characters left Martha to continue her good work in the villages with the magical item then there's one more part.
Combine this with the next Side Quest's encounter, and if the characters get into trouble, Martha comes to their aid, along with her sons, Harry and Oscar.
If not, leave this encounter till the next time the characters are in trouble in the villages.


\humansoldier[\npc{\T\M}{Harry \& Oscar}]

\stopcontents[sq]

\startcontents[Forest]
\resumecontents[Villages]
\sidequest{Wolf Heads}
\stopcontents[Forest]
\stopcontents[Villages]

\startcontents[sq]

\sqminitoc

\noindent
The characters hear of an adventuring party travelling around, and eventually meet them.
This Side Quest is designed to run in the background, and flavour other encounters.

\NPC{\F}{Jubilee}{Cheeky}{Thumbs up}{Experience}
\person{0}% STRENGTH
{2}% DEXTERITY 
{1}% SPEED
{{0}% INTELLIGENCE
{1}% WITS
{1}}% CHARISMA
{0}% DR
{1}% COMBAT
{Projectiles 2, Athletics~1, Deceit 1, Larceny~3, Stealth 1, Vigilance~1\knacks{\precisestrike, \quickshot}}% SKILLS
{\shortsword, \partialleather, dagger, short bow, 12sp.}% EQUIPMENT
{\addtocounter{fp}{5}}

\NPC{\M}{Bertold}{Angry}{Massive Grin}{V\'er\"e}
\person{3}% STRENGTH
{1}% DEXTERITY 
{1}% SPEED
{{0}% INTELLIGENCE
{-1}% WITS
{0}}% CHARISMA
{0}% DR
{2}% COMBAT
{Projectiles 1, Athletics 2, Deceit 1, Wyldcrafting 2\knacks{\adrenalinesurge, \unstoppable, \finishingblow, \charge}}% SKILLS
{\greatsword, \partialchain, dagger, long bow, camping equipment}% EQUIPMENT
{}

\NPC{\M}{Lucretius}{Quiet}{Clicks tongue}{V\'er\"e}
\person{1}% STRENGTH
{0}% DEXTERITY 
{-1}% SPEED
{{2}% INTELLIGENCE
{0}% WITS
{0}}% CHARISMA
{0}% DR
{0}% COMBAT
{Projectiles~1, Academics~2, Crafts~1, Wyldcrafting~1\knacks{\combatcaster}
\Path{Alchemy}{\conjuration 3, \invocation 3, \illusion 1}}% SKILLS
{Mana stone worth 4 MP}% EQUIPMENT
{\lockedmana{2}}

\beardedalemaster
\label{beardedalemaster}

\Gls{beardedalemaster} gets into other shenanigans later, in \nameref{troubleAle}, page \pageref{troubleAle}.

\sqpart{Town}% AREA
{\N Rumours of Glory}% NAME
{The party hear about famous warriors in their area}% SUMMARY

The local gossip relates to how the local heroes, known as the \textit{Wolf Heads}, recently protected a town from a goblin invasion.

\begin{enumerate}

	\item{There were around forty goblins.}
	\item{The Wolf Heads stopped them with grease spells over the walls, so the goblins could not climb properly, leaving the archers time to pick off more.}
	\item{The local village master refused to pay for goblin heads taken, claiming it would be illegal as it would count as him raising his own army.}
	\item{The party refused to gather payment from the local peasants, but have sworn a blood oath to kill Village Master Solus.}
\end{enumerate}

This last rumour isn't entirely accurate -- the Wolf Heads are irritated with Village Master Solus (see page \pageref{solus}), but have not sworn an oath to kill him.

Play this encounter at the same time as the next Side Quest part, and raise the local nura rating by 1.

\sqpart{Villages}% AREA
{Arm Wrestling}% NAME
{The characters meet the Wolf Heads}% SUMMARY

The Wolf Heads have just encountered ten chitincrawlers, and slain the lot.
They used a giant web spell, along with a lot of arrow fire.

They are tired, but unwounded, and are up for fun, so they went to a nearby village to drink.
Wherever the characters are they encounter the Wolf Heads, who then challenge them to arm-wrestling contests, or magical contests.
The Wolf Heads will also speak about encountering goblins, both during the recent assault on a village, and in general.
The games are mostly in good faith, but if the party have any obviously valuable items, Jubilee will attempt to steal it from them.

\sqpart{Town}% AREA
{\N Rumours of Ogres}% NAME
{The party hear more of the Wolf Heads}% SUMMARY

The characters hear more of the wolf heads, specifically about how they are looking for some abandoned temple to Qualm\"e, which could be the source of so many goblins in the area.

Play this encounter at the same time as the next Side Quest on your list.

\sqpart{Forest}% AREA
{Helping Hands}% NAME
{The party encounter the Wolf Heads again, who lend a hand}% SUMMARY

The wolf heads come across the party at the same time as another encounter.
If the party are in trouble, they come to save them.
If the party aren't, they may just walk together with them for safety.

\sqpart{Villages}% AREA
{\N A Cry for Help}% NAME
{The party find the Wolf Heads attacked by ogres}% SUMMARY

A dozen ogres ambushed the Wolf Heads on the road, and the party heard their cries in the distance.
By the time the party arrive (assuming they come towards the cries) Bertold is unconscious (-3 HP), and Jubilee is nearly dead (2 HP left).

Ten ogres remain alive.
Five are focussed on the remaining Wolf Heads, and the other five have decided to wait in the bushes to ambush the party (or if not the bushes, inside nearby houses, or anywhere appropriate).

\paragraph{If the party leave}
then the Wolf Heads die.
If the Wolf Heads survive, they will rest in \gls{town}, and be ready for any call to adventure, such as joining the fight against the nura in the future.

\ogre[\npc{\T\N}{5 Ogres}]

\ogre[\npc{\T\N}{5 Hiding Ogres}]

\stopcontents[sq]

\sidequest{Sewer Bandits}\label{sewerbandits}

\startcontents[sq]

\sqminitoc

\noindent
\Gls{sewerking} of Whiteplains has built an underground lair, and is using it to build a team of bandits.
Unfortunately, the nura influence below starts to spill upwards.

\sqpart{Town}% AREA
{\N Bad Water}% NAME
{The town's main spring smells disgusting}% SUMMARY

Deep underground, all the changes made in the secret sewer lair have unsettled the earth.
As a result, \gls{town}'s water has gone foul.

\begin{boxtext}

	You stop at a nearby fountain, as everyone does in \gls{town}, but the water taste's suddenly foul.
	The rest of the night, others make the same complaints about the rotten taste.

\end{boxtext}

Raise the local Nura Rating by 1.
Play this encounter at the same time as the next Side Quest part.

\sqpart{Town}% AREA
{\N The Nura Child}% NAME
{A street urchin transforms into a nura creature}% SUMMARY

\Gls{sewerthief} was delivering nura-enchanted food to \gls{townmaster}, so that \gls{townmaster} could use it to poison enemies in the future.
En route, a street-child stole a piece of dried meat, and turned into an ogre.

\begin{boxtext}

	Screams erupt nearby and people flee.
	Around the tavern's corner, a monster walks out.
	It stoops to pick up a piece of beef one of the people had dropped, then stares at it with large, innocent eyes.

\end{boxtext}

Characters can make a Wits + Medicine roll, TN 9, to notice that this monster is in fact a kid who's simply been afflicted by dark magic, and turned into a small ogre.

\paragraph{Curing the child won't be easy,}
but sufficient research at a Temple to C\'{a}l\"{e} shows that it can be done through starvation (Intelligence + Academics, TN 8).\footnote{See \autoref{nura} for more.}

\paragraph{If the characters want to ask the child about where he stole the food on the spot,}
they can make an Intelligence + Empathy roll, TN 11, to find which direction \gls{sewerthief} went.

Raise the local Nura Rating by 1.

\deephobgoblin[\npc{\M\N}{James, Street-Child}]

\paragraph{If taken alive,}
James will be able to describe \gls{sewerthief}' face, but couldn't say who he is.
James has only been a nura for a couple of hours, so if the party know to starve him, he can return to being a normal child before long.

\sqpart{Town}% AREA
{\N Streetbrawl}% NAME
{The local alcoholics are on the street, and fights are breaking out}% SUMMARY

Recently, the \gls{pig} has had a problem with the undead nura living underground, and has had to close to deal with the situation.\footnote{See page \pageref{pig_pantry} for details on what happened.}
As a result, all the alcoholics who have been chucked out of every area in the city are out in full swing.
The rest of the night is full of random people brawling.

\begin{boxtext}

A drunken man with uncomprehending eyes stares at you, then shouts.

\begin{quotation}

	Oi! Pointy-eared freak.  The freakshow left last week.  They decide to leave you behind?

\end{quotation}

\end{boxtext}

Trev's not happy, and he's taking it out on the group.
Pete doesn't know the group, but he's decided he's going to speak up for them and kick the crap out of Trev.
However the PCs react, a fight will break out around them.
Once the fight's ended, the characters might think they're out of harm's way, but the streets are rampant with trouble.

\humanfarmer[\npc{\T}{Irate Alcoholics}]

\paragraph{If the characters approach \gls{pig},}
they hear noises inside as of heavy metalwork.
The Whiteplains thieves in the sewers are fighting back a horde of undead who got free.
Three of them are sentient, and quite capable of spotting and attacking humans wearing magical rings.
The other twelve are only taking orders, but the total gang has proved to be too much for the thieves, who've barricaded the trapdoor down, and are trying to poke spears down.

\paragraph{If the PCs come to the rescue and don't know about the situation,}
\gls{pigowner} and the rest tell the PCs that they're as surprised as anyone else that the long, mysterious hallways down there have undead living there.

\sqpart{Town}% AREA
{\N \N Unexpected Ghouls}% NAME
{Hobgoblin ghouls from below emerge and attack the town}% SUMMARY

Twenty of the undead from the sewer have escaped because the bandits who live down there were simply not careful enough.  If the characters confront the dead head-on, they will have a bad time, six of the town guard are only a few streets away.

\begin{boxtext}

	Screams erupt next door.
	Feet move quickly, and you see three men being pulled to the ground by a silent mob of massive, naked men.
	The mob pulls the men inside but makes no sound, and then the screaming stops.

	Another steps closer, and you can see these are not men, but stinking creatures wearing nothing but long beards.
	Their ears are pointed, and their bodies covered in warts.
	Each one has dead, white eyes, or gouged out eye sockets, but they look at you with intense interest, then start to walk towards you.

	In the far distance, the town guard can be heard shouting to keep the noise down.

\end{boxtext}

Any character can summon the nearby guard with a Strength + Empathy roll, TN 7, to shout out loud.
The guards take 5 rounds to arrive, but every Margin on the roll reduces that time by 1 round, to a minimum of 2 rounds.

\paragraph{At the end of the day,}
the ghouls will be put to the sword one way or another, but the question remains; what brought them here?

\undeadhobgoblin[\npc{\T\D\N}{20 Undead Hobgoblins}]

Increase the local nura rating by 1.

\sqpart{Town}% AREA
{\N Underground Assassins}% NAME
{The bandits in the sewer cut \Glsentrytext{captain}'s throat}% SUMMARY

\Gls{captain} recently asked a number of guards about recent nura sightings, and why there was suspicious activity in \gls{town}, such as the strange Whiteplains accent heard in \gls{pig}.
He got too close, and \gls{sewerking} felt he had to die, so four men were requested to pretend to be drunk while walking down the street, and then suddenly stab him in the neck.

\begin{boxtext}

	You hear guards shouting ``After them!'', in the distance, and quickly scurrying feet, as a woman shouts for someone to help.

\end{boxtext}

\paragraph{If the characters stay to help the wounded man,}
they find \gls{captain} with a knife-wound, next to his wife.
The roll is Wits + Medicine, at TN 9 to save his life.

\paragraph{If they run after the thieves,}
the make a Group Roll of Speed + Athletics.%
\iftoggle{core}%
{\footnote{See the core rules, page \pageref{grouproll}, for Group Rolls.}}%
{}%
Remember that whoever's trying to patch up \gls{captain}'s bleeding neck won't be able to join the chase.

\begin{tcolorbox}[tabularx={cX},top=10pt,bottom=10pt]

	Roll & Result \\\hline
	12 & \textit{``Giving chase, you catch up to four men running from the scene of the crime.''} \\
	11 & \textit{``You run round an alley, and find a drain cover clanking. The assassins have jumped underground.''} \\
	9 & \textit{``You run in hot pursuit, but the attackers have disappeared down a street, into thin air.''} \\
	7 & \textit{``The attackers sprint away, leaving you running in the dark.''} \\

\end{tcolorbox}

\sewerthief

\humanthief[\npc{\T}{Four of the Sewer Thieves}]

\paragraph{Next:} If the party follow the assassins underground, they run to the nearest entrance -- perhaps the butchers or \gls{pig}.
Go to page \pageref{sewers}.  Otherwise, this incident will remain a mystery.

If \gls{captain} survives, he has little idea of what's happening, although a little investigation could reveal what he's been asking about recently (Intelligence + Vigilance Teamwork Roll, TN 10).
Raise the local Nura Rating by 1.

\sqpart{Town}% AREA
{\N \N The Nura Rise}% NAME
{The hordes underneath begin to spill above ground}% SUMMARY

\Gls{sewerking} was betrayed.  The hobgoblins sent up to work for him from the hell-realm were instructed to grab him and force the password out of him.  He fled immediately, and a number of them stayed to capture his companions.  All the thieves in the sewer were then turned into ogres.

\begin{speechtext}

	A familiar looking man darts past you.
	Running after him, or perhaps after you, are ten ugly creatures with clubs, just like the dead creatures you saw before, but this time \emph{alive}.

\end{speechtext}

\deephobgoblin[\npc{\T\N}{10 Hobgoblins}]

If the characters capture \gls{sewerking}, he begins shouting something over and over -- the activation phrase for the portal.\footnote{The activation word is simply elvish for `We are open for trade'.}
Once captured, he knows that he's done for, and may as well give the nura the password to open the gate they so desire.
Three ogres are in earshot, and they immediately run to \gls{pig} to go underground and open the gate.

While \gls{sewerking} was sprinting underground, he managed to open one cell with various undead hobgoblins inside, in order to slow the nura horde down.  However, they have now come above ground.  In total, above ground are:

\begin{itemize}

	\item{Two groups of ten hobgoblins.}
	\item{Three ogres.}
	\item{Six undead hobgoblins.}

\end{itemize}

\undeadhobgoblin[\npc{\T\N\D}{6 Undead Hobgoblins}]

\ogre[\npc{\N\T}{3 Ogres}]

\paragraph{If that password ever makes its way underground, \gls{town} is doomed.}
Nura spill from everywhere.
Raise the local nura rating by 2, and roll for an encounter each time the characters enter a new area, even within \gls{town}.
The party may find nura spiders, goblins, or any number of other creatures.

\stopcontents[sq]

\sidequest{Random Meetings}\label{randommeetings}

\startcontents[sq]

\sqminitoc

\sqpart{Town}% AREA
{Pickpocketed}% NAME
{Someone has stolen one of the characters' items}% SUMMARY

\begin{boxtext}

	You feel the side of your leg and suddenly your heart stops.
	You wonder where is\ldots \textit{which item is the last one written on your character sheet?}

\end{boxtext}

The player with the highest Intelligence Bonus makes a Wits + Vigilance roll, TN 9.
Failure indicates a thief has stolen the last item on the player's character sheet, and successfully fled.

It might be possible to track the thief, down with enough time spent investigating, but it'll require an Intelligence + Vigilance roll at TN 11.

Gary, local pickpocket, has no excuse except for his young age, and bad luck in life, though he's 25 and was recently kicked out of an apprenticeship for drinking too heavily on the job.

If the players turn him in, he will be pulled up about previous crimes at the same time, and hanged by the end of the week.
With the right deal, he might agree to accompany them on a dangerous mission.

\humanthief[\NPC{\M}{Gary}{Arrogant}{Picks nose}{Laiqu\"e}]

\sqpart{Town}% AREA
{The Riddle}% NAME
{\Glsentrytext{college} have raised a bounty on a riddle}% SUMMARY

The townsfolk are abuzz with a riddle.
The mage's guild have become tired of only getting students sent by rich parents with no real ability.
They've raised a bounty of 5gp to anyone who can solve the riddle.
Their hope is that the most talented young men of the realm will come forward, and once they do, a long talk proceeds, as the mages tell whoever gets the correct answer why they should join \gls{college}.

\begin{boxtext}
Three gods stand in front of you.  One always says the truth, one always lies, and one speaks randomly.  They refuse to speak your language, but they understand what you are saying.  You have three questions to identify each god.
\end{boxtext}

To answer the riddle correctly, the characters must pose three questions which guarantee they will correctly identify each god -- guessing is not allowed.
\footnote{This question is mostly here for fun.  Most players will have zero chance of answering this riddle.
It's placed here solely as a conversation piece, but be aware that this is technically possible, and the party may soon be very rich.}

\sqpart{Town}% AREA
{The Captain}% NAME
{\Glsentrytext{captain} jumps into another encounter to help the party}% SUMMARY

\captain

\Gls{captain} is heading home after a long night (if this is the afternoon, it's been a \emph{very} long night).
Combine this encounter with the next Side Quest part.
If the party are in trouble, he helps.
If they're known thieves, he tries to follow them, then brings guards.

\sqpart{Town}% AREA
{Illegal Songs}% NAME
{A bard is caught singing about how much better it was before the current ruler, \Glsentrytext{king}}% SUMMARY

\begin{boxtext}
	A little crowd have gathered in the side street, occasionally clapping their hands to a beat.
	Getting closer, you can hear a mandolin and a soft voice, singing of great heroes fighting back nura hordes coming up from the depth.
\end{boxtext}

The singer is \gls{warningbard}, and he's here to sing the praises of any nobles who might pull together an army to push back the nura, like nobles did in the old days.
However, such an army would be illegal, as only \gls{king} can gather armies, or amass large amounts of weaponry, and so far the area does not have enough of \gls{nightguard} to defend itself from rising levels of nura.

\begin{boxtext}

	Five armoured guards arrive, and start shouting ``Move aside!'', to the crowd.  Without word of explanation, they grab the bard and drag him away.  The mandolin sits on the ground alone, and the crowd just sit there, stunned and morose.

\end{boxtext}

The guards understand who their employer is -- they're loyal to \gls{king} and know that it's best to avoid the notion of nobles having their own standing armies, even in songs about a long time ago.

For the rest of the day, all of \gls{town} is full of talk about how in the old days, the nura were not a threat, because individual lords could pull together the armies they needed to quell nura uprisings.

\paragraph{If the party do not intervene,}
\gls{warningbard} will end up in the cells for six months.
\footnote{For the cells under \glsentrytext{nightguard} station, see page \pageref{guardstation}.}

\warningbard

\sqpart{Town}% AREA
{Pickpocketed Again!}% NAME
{Someone else picks a character's pocket}% SUMMARY

The player with the highest Academics Skill rolls Wits + Vigilance, TN 8.

\begin{boxtext}

	You feel your trouser leg, and suddenly think ``Oh, not again!''.  \textit{What's the last item written on your character sheet?}

\end{boxtext}

This time it's a young woman, down on her luck as her family farm was consumed entirely by nura.
The rest have died, and she's now alone.

\paragraph{If caught,}
Judith apologizes profusely, and explains her situation.

\humanthief[\NPC{\F}{Judith}{Fearful}{Messes up her hair}{Acquisition}]


\end{multicols}

\stopcontents[Town]
\stopcontents[sq]
