\chapter{Side Quests}
  \epigraph{He attacked everything in life with a mix of extraordinary genius and na\"ive incompetence, and it was often difficult to tell which was which.}{Douglas Adams}

\begin{multicols}{2}

\noindent
Trouble has arrived in Mt Arthur, as \gls{townmaster} has begun plotting to isolate the area in order to better control it, and two dispossessed nobles from Whiteplains have begun stirring things up.
The characters will feel the echoes of their schemes for a long while before they understand what brews underneath \gls{town}.

In the beginning, the standard chaos that \gls{town} offers will keep them entertained.
Soon after, various hooks will tempt them out to the villages, and eventually into the deep forest.
This area has no reliable maps, even for you -- the Games Master.
While many areas have a set location, others arrive as part of a Side Quest, so they will appear exactly where the PCs wander, and stay put from then on.
The map will morph and grow as the troupe wander across it, and you will have to update and mark the various locations.

As various clues present themselves, inquisitive players will have more opportunities to discover what will happen if they do nothing.

The stories begin during the season of \season\ (take a look at page \pageref{astronomy} for a seasons).

\subsection{Cassus Belli}
\label{expanding_wilderness}

In the words of \gls{townmaster}:

\begin{exampletext}

  Our city has burnt to the ground so long ago that people have forgotten where it is.
  They just call it `\gls{lostcity}'.

  Back before King Dalius, \glsentrylong{town} stretched across every part of the forest, through all the villages.
  Every road and hamlet around \gls{town} once rested safely, surrounded by towering stone walls.
  It had its own little alchemical college, and a great logging industry.
  It had temples to all the gods, and great shipyards.
  The great city held the surrounding farmlands inside its great rings of walls.
  It went so far out that a man could not walk across it in a single day.

  Then the evil elves came.
  Elves cannot lift great rocks to build walls, like men.
  They feel jealous of everything we do, so they destroyed \gls{lostcity}, and with evil magics toppled our walls.

  The common folk returned here a few generations ago, and picked up all the stone along the roads for their houses and little walls, but if you journey just a little out into the darkness, you'll find ancient artefacts still resting, or half-buried.

  I will rebuild our heritage, brick and stone, walls and farms, temples and hearths.

  Of course, I require an army, and of course \gls{king} no longer permits armies.
  His clear negligence has allows beasts and bandits to run loose throughout the land\ldots which gave me a plan.
  What about simply employing the bandits? -- genius!
  After all, I do not have to wander out and speak with them personally.
  Hell, I don't even need to reveal myself.
  I don't need any trust either -- in fact I need nothing from them but obedience.

  My two loyal servants, \gls{traitor} and Jared, take deliveries of food, coin, and sometimes even weapons, into the forest.
  A separate note explains where the reward sits, and what they must do.
  In this way, I have kept all the bandits busy, and the people truly have less to fear.
  People see them seldom, but everyone knows they still exist, so the locals have taken to calling them the `Woodspy Bandits'.
  I rather like the name.

  So far, we have built an outpost and I have ordered it be painted green, for camouflage.
  We cannot let anyone in on the action until some fortifications have been placed.
  
  I will bring up defence towers, rebuild the old walls in the forest, and finally reclaim our land!

\end{exampletext}

In the words of \gls{traitor}:

\begin{exampletext}

  My Lord,
  I have found a map in \gls{town}'s library, with markings around \gls{lostcity} (I have enclosed a copy).
  I cannot see where these locations lie, but I will attempt to find each of them.

  \Gls{captain} does not suspect me of anything wrong.
  He is an idiot, and I have simply told him I am scouting, and whenever I return, I bring him more news of the bandits we ran into, now known as `the immortal bandits'.
  Even if I see nothing, I still return with some news of them.
  Right now, he feels they are everywhere, and even has a map showing them surrounding the city on all sides.
  He has sent men to hunt for them everywhere.

  The \gls{greentower}'s construction is going as planned, and will finish soon.
  We have brought limestone paint from Redfall, and the outside will be painted soon.
  It already looks invisible from far away.

  I remain your faithful servant,

  \gls{traitor}

\end{exampletext}

In the words of \gls{banditking}:

\begin{exampletext}

  We came from something, and we'll return.

  After \gls{king} sent his \gls{nightguard} to kill our parents, take our land, and hand it to bureaucrats, my brother \gls{sewerking} fled from \gls{college}, and I guided him safely to Mount Arthur.
  \footnote{See page \pageref{whiteplainsWar} for the massacre of Whiteplains.}
  We've seen ten cycles\footnote{See \autoref{astronomy}.} pass, and started planning our return to nobility.
  And if \gls{king} can play rough, then so can we.

  We've robbed a lot of people, and I don't feel good about that, but we really have no choice.
  We've done nothing worse than \gls{king}.

  I chose \gls{town} as it was out of the way, and we started to explore the area.
  Just him and me, wandering with our rations and picking up food.
  We were essentially living like bandits even before we became bandits.
  Soon enough, I found our home-to-be, where the dead priest lives.
  He's a shy creature, but we've gained an understanding.

  We came to his temple at night, so I saw an arrow cutting into the earth before I saw the building.
  I could clearly see it was a warning, but didn't understand how anyone could possibly see us walking in the darkness.
  Any idiot could tell that nobody sat there with an army, or we would have heard them, and they could have killed us.
  So we kept back until daylight, and went out again, cloaks waving about, hailing whoever might be there.
  And another arrow went into the ground -- another warning shot.
  When we saw the crumbling temple, alone in the forest, we knew we'd found a new home, despite the arrows.

  It took a long time to make friends with the dead priest.

  We stayed put, camping nearby, and the dead priest came to me in my dreams.
  We spoke for a while -- from a distance.
  In dreams, the dead can fear the living.
  After spending so much time alone, I honestly thing he's gained a fear of the living.
  Like people who hate spiders, but he just can't stand to see anything animate and moving about on its own.

  We don't want to kill people, just get enough money to get by.
  But of course, accidents happen, and at those times, we return with the bodies and give them to \gls{necromancer}.
  He's been building up quite the little army down there.
  It got to be a bit of a problem after a while -- I'm not afraid of the dead wandering about at night, in fact I've become quite used to it.
  \emph{However}, after one attacked me, things had to change.
  \gls{necromancer} eventually gave me a ring which would make me invisible to the dead, and another for my brother.
  We can't wear them for too long, but we can stay a night in the temple without worry, and the dead do not attack our horses.

  And today, things have changed.
  We discovered another team of bandits, who call themselves the Woodspies.
  We found a letter they sent as well, along with a map showing old locations from \gls{lostcity}, and the idiots don't know what they've found.
  Those old symbols show portals to the nura realms.%
  \footnote{See page \pageref{darknessandfire} for \nameref{darknessandfire}.}

  That idiot, \gls{traitor}, will eventually open one of those portals, and pour nura into the land around.
  And this will be exceedingly useful for us.
  Once the road to the outside has closed, the \gls{town} will sit alone, with no way to call for outside help, and no way for \gls{king} to interfere.

  Only one road goes in or out.
  So long as \gls{town} remains divided from the rest of the world, we can simply kill \gls{townmaster}, destroy the Woodspy Bandits, and rule \gls{town} and the surrounding villages.
  If the nura prove too much trouble, \gls{necromancer} can help protect us, and strengthen his army.

  We will regain our birthrights in this new land before the season has ended.

\end{exampletext}

In the words of \gls{sewerking}:

\begin{exampletext}

  I had to return from my studies in \gls{college} to rescue \gls{banditking}, and found him hiding in some seedy bar, as if the \gls{nightguard} had such a high opinion of him that they wouldn't think to look for him there.
  I organized our escape, and found us lodgings with the little coin we still had.

  It started with \gls{pig}.
  \Gls{pigowner} heard our coins a mile away -- people with money don't normally stay in \gls{pig} for more than they need to get the gossip, and don't take their money in with them.
  We stayed in a little area she had tucked away, underneath, and she told us the tale of how there was once an underground library, and temple, and all sorts down there.

  We started cleaning the place out.
  We put on these rings that \gls{necromancer} had gifted us, and found we could hold our breath underwater for a good long time, and set to work investigating what we could in the darkness.
  As luck would have it, we found one passage gave way easily to our shovels, so we expanded it, and pulled out junk and crap stuck in there for centuries, and pushed some more, cleared more, and eventually managed to divert the water pouring downwards.
  We slunk through all the shit of \gls{town} for a solid month, and eventually the waters started to recede.

  We have our own underground palace now.
  It's dirty, and wet, it's dark and dank, but it's ours, and we can store anything we need down there.
  And it has treasures.

  So. Many. Treasures.

  It has an old brazier which infuses any corpse with a malicious spirit.
  It has an alter of skulls which whisper your fortunes and mishaps.
  And it has old scrolls, secreted away, which describe the forbidden magics of the nura.
  
  So I've begun to study that magic, and create magical items.
  And I've asked the skulls about my fortune, and sometimes they tell me I will kill \gls{townmaster}, so now I know I am destined to rule this land.
  And I've begun to steal the bodies to create an army down here.

  We can't have the dead just wandering around, so we lock them up in little cages.
  We have over a dozen cages, just full of the hungry dead, waiting to get out and tear the town apart.

  And we have used the dead to create tunnels.
  We've dug into the catacombs of \gls{town}'s temple of Qualm\"{e}, and started another which will soon reach underneath the citadel.
  Then we can release the dead into the citadel, and take out \gls{townmaster}.

  Tunnels and skulls, forgotten magics in scrolls, and secret passages.
  Everything will come together soon, but we need one more piece before we rule this land\ldots

\end{exampletext}

In the words of \gls{forestpriest}:

\begin{exampletext}

  I walk a lot.
  The last of my friends died of old age ten years ago -- that's the curse of long life inherited from my father.
  But I've started to think of it differently recently -- perhaps I just live many lives, one after another.
  And in this one I walk a lot.

  I quite like being a bird when I'm not walking.
  You can see a lot when you fly.
  It felt dangerous, because sometimes a hawk would try to eat me, but then I learned how to become a hawk.
  I like shouting a lot.
  And recently I've seen a lot of people dropping off military supplies -- food, armour and weapons mostly.
  And when I waited, I saw the Woodspy bandits, picking up the shipments.
  
  I don't usually interfere with this sort of thing -- in fact in the priesthood we swear ourselves to neutrality wherever possible, but that was my last life.
  So I turned the delivery man into a goat.

  I think he'll make a good goat.

  Before the deliveryman starting oinking, he started shouting about how I'm an evil elf.
  That's only half true at best, but it got me wondering if he was correct, and if the elves did destroy the city that spanned this area.
  The elves tell me another story -- that the humans opened portals to the nura realms in order to trade with them.
  So I walk around the ruins, and try to see if I can find something that looks like an alchemical portal.
  And I found those green bandits searching the ruins too, and building things.
  Sometimes questions just get stuck in my head like that, so I'm going to keep searching until I can find the answers.

\end{exampletext}

\subsection{The Party's Premise}

Consider setting up the party as members of \gls{nightguard}.
While not necessary, this will give each of them a reason to adventure, and grant them all a place in the world without any work.

\end{multicols}
